\documentclass[11pt]{article}
\usepackage{amsmath,amssymb}
\usepackage[margin=1in]{geometry}

\title{Blueprint for the Ternary Quartic Rank-4 Proof}
\date{2026-04-09}

\begin{document}
\maketitle

\section*{Status}
This blueprint is not complete yet. The abstract SOCP component, orthogonal
mixing of the four factor coordinates, and all currently identified
representative-level rank-$4$ algebraic branches are now formalized. The
remaining gap is the normalization/classification step that reduces an
arbitrary admissible rank-$4$ factor to those representatives, followed by the
final theorem assembly in \texttt{TernaryQuarticProof.lean}.

\medskip
\noindent
One additional obstruction is now explicit. Orthogonal mixing acts on the
\emph{coefficient-space} preimages of the four coordinates, not directly on the
affine-linear polynomial space. So even when
$W \cap \mathbb{R}[x_1,x_2]_{\le 1}$ contains the lines of $1$, $x_1$, or
$x_2$, orthogonal mixing alone need not place those exact polynomials among the
coordinates. This is why an affine variable-change transport layer is needed in
addition to factor-coordinate mixing.

\section*{Formalized Step}
Let $r(u,p) = \sigma(u) - p$. The following statement is already formalized in
\texttt{TernaryQuarticProof/Certificate.lean}.

\medskip
\noindent
\textbf{Lemma.} Assume:
\begin{enumerate}
\item $u$ is an admissible rank-4 point;
\item $u$ is a SOCP for $p$ with respect to a positive-definite bilinear form
$B$;
\item $r(u,p)$ is orthogonal to $\operatorname{im}(A_{u_{\mathrm{img}}})$ for
some admissible factor $u_{\mathrm{img}}$;
\item there is a decomposition
\[
  p = q_{\mathrm{img}} + \sum_{w \in W} \sigma(w),
\]
where $q_{\mathrm{img}} \in \operatorname{im}(A_{u_{\mathrm{img}}})$ and every
$w \in W$ is an admissible kernel direction for $A_u$.
\end{enumerate}
Then $r(u,p) = 0$.

\medskip
\noindent
\textbf{Proof.}
The image term pairs to zero by the first-order condition. Each kernel-square
term pairs nonnegatively with the residual by the second-order condition,
because $A_u(w)=0$ removes the positive $2\langle A_u(w), A_u(w)\rangle$ part
of the Hessian expression. Summing these inequalities gives
$\langle p, r(u,p)\rangle \ge 0$. On the other hand, the first-order condition
at $v=u$ gives $\langle \sigma(u), r(u,p)\rangle = 0$, hence
\[
  \langle r(u,p), r(u,p)\rangle
  = \langle \sigma(u)-p, r(u,p)\rangle
  = - \langle p, r(u,p)\rangle \le 0.
\]
Positive-definiteness implies $\langle r(u,p), r(u,p)\rangle \ge 0$, so the
objective is zero and therefore $r(u,p)=0$.

\section*{Additional Formalized Reductions}
Two special cases are now formalized.

\medskip
\noindent
\textbf{Image case.} If $p \in \operatorname{im}(A_u)$ with an admissible
preimage, then $r(u,p)=0$ directly from the first-order condition and positive
definiteness.

\medskip
\noindent
\textbf{Constant-relation case.} If there exists a nonzero constant vector
$c \in \mathbb{R}^4$ with $\sum_i c_i u_i = 0$, then $r(u,p)=0$ for every SOS
quartic $p$.

\medskip
\noindent
\textbf{Proof sketch.}
For each quadratic summand $q^2$ in an SOS decomposition of $p$, define the
admissible direction $w_i = c_i q$. Then
\[
  A_u(w) = q \sum_i c_i u_i = 0,
  \qquad
  \sigma(w) = \Bigl(\sum_i c_i^2\Bigr) q^2.
\]
Since $\sum_i c_i^2 > 0$, the SOCP inequality in the direction $w$ implies
$\langle q^2, r(u,p)\rangle \ge 0$. Summing over the SOS decomposition yields
$\langle p, r(u,p)\rangle \ge 0$, and the abstract SOCP argument finishes as
before.

\medskip
\noindent
\textbf{Orthogonal factor-coordinate invariance.} Let $M \in \mathbb{R}^{4
\times 4}$ satisfy $M^\top M = M M^\top = I$, and define
\[
  (\operatorname{mix}_M u)_i = \sum_j M_{ij} u_j.
\]
Then the following are now formalized in
\texttt{TernaryQuarticProof/FactorTransform.lean}:
\begin{enumerate}
\item admissibility is preserved by $\operatorname{mix}_M$;
\item $A_{\operatorname{mix}_M u}(v) = A_u(\operatorname{mix}_{M^\top} v)$;
\item $\sigma(\operatorname{mix}_M u) = \sigma(u)$ and hence
$r(p,\operatorname{mix}_M u)=r(p,u)$;
\item FOCP and SOCP are preserved under orthogonal mixing.
\end{enumerate}
Therefore the eventual factor-space case analysis may be carried out up to
orthogonal recombination of the four coordinates.

\medskip
\noindent
\textbf{Formalized surjective constant branch.} The constant
$\dim(W \cap \mathbb{R}[x_1,x_2]_{\le 1}) = 1$ model
\[
  u_{\mathrm{const}} = (1, x_1^2, x_1x_2, x_2^2)
\]
is now fully formalized in
\texttt{TernaryQuarticProof/RepresentativeSurjective.lean}. The key lemma is
that every quartic monomial of degree at most $4$ lies in
$\operatorname{im}(A_{u_{\mathrm{const}}})$, hence every quartic polynomial does
as well by linearity. Therefore every SOCP at this representative has zero
residual by the already-formalized image case. The same file now also proves a
basis-invariant upgrade: if an admissible factor contains $1$ and admits three
independent homogeneous quadratic relations whose coefficient matrix in the
basis $(x_1^2,x_1x_2,x_2^2)$ has nonzero determinant, then Lean reconstructs
the exact monomial relations and closes the branch abstractly. This
reconstruction is also transported across affine equivalences.

\medskip
\noindent
\textbf{Formalized surjective mixed-affine representative.} The model
\[
  u_{\mathrm{ma},15} = (1, x_1, x_1^2, x_2^2)
\]
is now formalized as well in
\texttt{TernaryQuarticProof/RepresentativeMixedAffine.lean}. Its proof is the
same style as the other surjective branches: every quartic monomial is hit
directly, either through the constant coordinate for total degree at most $2$,
through $x_1^2$ for monomials with $x_1$-degree at least $2$, or through
$x_2^2$ for the remaining quartic monomials. So every SOCP at this
representative has zero residual by the abstract image certificate.

\medskip
\noindent
\textbf{Formalized codimension-one image reduction.} The mixed-affine
rank-$14$ representative
\[
  u_{\mathrm{ma},14} = (1, x_1, x_1x_2, x_2^2)
\]
now has a full representative-level formalization in
\texttt{TernaryQuarticProof/RepresentativeMixedAffine.lean}: if a quartic
polynomial has zero $x_1^4$ coefficient, then it lies in
$\operatorname{im}(A_{u_{\mathrm{ma},14}})$. The proof is again
monomial-by-monomial. Every quartic monomial except $x_1^4$ is represented by
one of the four coordinates, and then support induction upgrades the monomial
statement to all quartics with vanishing $x_1^4$ coefficient. The same file now
also proves that every SOCP at this representative has zero residual. The extra
step uses the six-monomial quadratic normal form to show that the $x_1^4$
coefficient of a quadratic square is exactly the square of its $x_1^2$
coefficient, and then matches that coefficient with the explicit kernel square
\[
  \sigma(0,0,-c x_1x_2, c x_1^2) = c^2(x_1^2x_2^2 + x_1^4).
\]

\medskip
\noindent
\textbf{Formalized codimension-two image reduction.} The mixed-affine
rank-$13$ representative
\[
  u_{\mathrm{ma},13} = (1, x_1, x_1^2, x_1x_2)
\]
is now also fully formalized in the same file. The image theorem proves that a
quartic lies in $\operatorname{im}(A_{u_{\mathrm{ma},13}})$ whenever its
$x_2^3$ and $x_2^4$ coefficients vanish. The key coefficient identities for a
quadratic square are
\[
  [x_2^3](q^2) = 2\,[x_2](q)\,[x_2^2](q), \qquad
  [x_2^4](q^2) = [x_2^2](q)^2,
\]
and the explicit kernel line
\[
  w_{b,c} = (0,0,-x_2(b+c x_2),x_1(b+c x_2))
\]
matches both missing quotient classes simultaneously. Hence every SOCP at this
representative also has zero residual.

\medskip
\noindent
\textbf{Formalized span-three branch.} The representative
\[
  u_{\mathrm{span3}} = (1, x_1, x_2, h),
\]
with arbitrary quadratic homogeneous $h$, is still formalized directly in
\texttt{TernaryQuarticProof/RepresentativeSpanThree.lean}. But the file now
proves a stronger abstract theorem: for any admissible rank-$4$ factor $u$, if
there exist coefficient vectors $c_0,c_1,c_2 \in \mathbb{R}^4$ with
\[
  \sum_i c_{0,i}u_i = 1,\qquad
  \sum_i c_{1,i}u_i = x_1,\qquad
  \sum_i c_{2,i}u_i = x_2,
\]
then every SOCP at $u$ has zero residual. Thus the whole
$\dim(W \cap \mathbb{R}[x_1,x_2]_{\le 1})=3$ branch is now solved without any
exact orthogonal normalization to a rigid representative.

Writing a quadratic factor as
\[
  q = q_{\le 1} + q_{\mathrm{hom}}, \qquad
  q_{\mathrm{hom}} = x_1 a + x_2 b,
\]
with $a,b$ linear, the proof first shows that every cubic lies in
$\operatorname{im}(A_u)$ from the span relations above. It then builds the raw
kernel direction
\[
  w_q^{\mathrm{raw}}
  =
  \operatorname{rel}(-c_0,q_{\mathrm{hom}})
  + \operatorname{rel}(c_1,a)
  + \operatorname{rel}(c_2,b),
\]
which satisfies $A_u(w_q^{\mathrm{raw}})=0$. Its quartic part is
\[
  \sigma(w_q^{\mathrm{raw}})_{\deg 4}
  =
  \Bigl(\sum_i c_{0,i}^2\Bigr) q_{\mathrm{hom}}^2,
\]
so after scaling by
\[
  \alpha = \Bigl(\sqrt{\sum_i c_{0,i}^2}\Bigr)^{-1}
\]
one gets a kernel vector $w_q=\alpha\, w_q^{\mathrm{raw}}$ for which
\[
  q^2 - \sigma(w_q)
\]
has total degree at most $3$. Summing these cubic image terms over an SOS
decomposition of $p$ and using the indexed-family certificate in
\texttt{Certificate.lean} yields $r(u,p)=0$ for every SOCP in this branch.

\section*{Normalization Target}
The main unresolved issue is no longer the representative algebra above. It is
the remaining normalization/classification step for the branches not already
covered by the abstract span-three theorem, namely the independent
constant-containing $\dim = 2$ configurations that should reduce, after affine
variable changes and orthogonal mixing of the four coordinates, to one of the
already-formalized low-affine or mixed-affine representatives.

\medskip
\noindent
The affine-normalization layer is now split across
\texttt{TernaryQuarticProof/AffineTransform.lean} and
\texttt{TernaryQuarticProof/AffineSocpTransform.lean}. The first file handles
the raw affine substitution $x \mapsto A x + b$: the induced algebra
homomorphism preserves the quadratic and quartic degree bounds and commutes
with the maps $A_u(\cdot)$, $\sigma$, and the residual after applying the
substitution coordinatewise. The second file upgrades the invertible case to an
actual transport statement: after moving the bilinear form by the inverse
equivalence, affine changes of variables preserve positive-definiteness,
admissibility, SOS quarticity, FOCP, SOCP, and the condition $r(u,p)=0$.

What still remains is the classification lemma producing the specific affine
and orthogonal normalization data for those remaining rank-$4$ factors. Once
that data exists, the representative theorems can be pulled back through these
two transport layers.

\medskip
\noindent
That pullback step is now also packaged explicitly in
\texttt{TernaryQuarticProof/RepresentativeTransport.lean}. The main theorem
there says: if, after applying an affine inverse on the polynomial side and an
orthogonal mixing on the factor side, the SOCP point becomes an already-solved
representative $u_{\mathrm{rep}}$, then the original point has zero residual.
So the remaining normalization problem is now sharply isolated to one kind of
statement:
\[
  \text{produce exact data with }
  \operatorname{mix}(M^\top,\; e^{-1}\!\cdot u) = u_{\mathrm{rep}}.
\]

\section*{Correction To The Previous Case Split}
The earlier five-case list is not complete. The constant-containing
$\dim(W \cap \mathbb{R}[x_1,x_2]_{\le 1}) = 2$ branch splits further.

\medskip
\noindent
Representative mixed-affine models recorded in the Julia script are:
\[
  (1,x_1,x_1^2,x_1x_2),\qquad
  (1,x_1,x_1x_2,x_2^2),\qquad
  (1,x_1,x_1^2,x_2^2).
\]
Their image ranks are respectively $13$, $14$, and $15$, with missing quotient
monomials
\[
  \{x_2^3,x_2^4\},\qquad \{x_1^4\},\qquad \varnothing.
\]
Moreover the first two mixed-affine representatives also admit explicit kernel
directions:
\[
  u=(1,x_1,x_1^2,x_1x_2),\qquad
  w_\ell=(0,0,-x_2\ell,x_1\ell),\qquad \ell=c+d x_2,
\]
for which
\[
  A_u(w_\ell)=0,\qquad
  \sigma(w_\ell)=(x_1^2+x_2^2)\ell^2 \equiv x_2^2\ell^2
  \equiv 2cd\,x_2^3 + d^2 x_2^4
  \pmod{\operatorname{im}(A_u)}.
\]
Hence this family generates the two missing quotient classes. For
\[
  u=(1,x_1,x_1x_2,x_2^2),
\]
the direction
\[
  w=(0,0,-x_1x_2,x_1^2)
\]
satisfies
\[
  A_u(w)=0,\qquad \sigma(w)=x_1^2x_2^2+x_1^4 \equiv x_1^4
  \pmod{\operatorname{im}(A_u)}.
\]
So the mixed-affine algebra itself now looks tractable; the old statement that
the proof was complete up to normalization was too optimistic because the
classification step still has to produce one of the formalized mixed-affine
plane types, not just a single concrete representative.

\medskip
\noindent
The current candidate normalization for the constant-containing
$\dim = 2$ branch is the following. After choosing coordinates so that
\[
  W \cap \mathbb{R}[x_1,x_2]_{\le 1} = \operatorname{span}(1,x_1),
\]
the remaining quadratic plane is a $2$-dimensional subspace of
\[
  \operatorname{span}(x_1^2,x_1x_2,x_2^2),
\]
hence the kernel of a nonzero binary quadratic
\[
  q = a x_1^2 + b x_1x_2 + c x_2^2
\]
in the dual $3$-space. Under shears and rescalings preserving the line
$\mathbb{R}x_1$, the quadratic $q$ should reduce to one of
\[
  x_2^2,\qquad x_1^2,\qquad x_1x_2,\qquad x_1^2 \pm x_2^2.
\]
The numerical sweep now recorded in
\texttt{julia/tq\_r4\_exploration\_2026\_04\_09.jl} matches this exactly:
\[
  x_2^2 \leadsto \operatorname{rank} 13,\qquad
  x_1^2 \leadsto \operatorname{rank} 14,
\]
while
\[
  x_1x_2,\qquad x_1^2 \pm x_2^2,\qquad x_1^2 + x_1x_2,
\]
and the tested generic examples all have image rank $15$, with numerical
errors again near $10^{-10}$ or smaller. So the only nonsurjective
mixed-affine representatives appear to be
\[
  (1,x_1,x_1^2,x_1x_2),\qquad (1,x_1,x_1x_2,x_2^2).
\]
The written reduction is therefore no longer mysterious: it should be the
binary-quadratic completion-of-squares argument above, implemented inside the
subgroup of affine changes preserving the line $\mathbb{R}x_1$.

\medskip
\noindent
The first algebraic half of this reduction is now formalized in
\texttt{TernaryQuarticProof/MixedAffineNormalization.lean}. That file defines
the shear $x_2 \mapsto t x_1 + x_2$ and the scaling $x_2 \mapsto s x_2$ as
explicit affine polynomial equivalences, and proves the key formulas
\[
  a x_1^2 + b x_1x_2 + c x_2^2
  \;\longmapsto\;
  \Bigl(a - \frac{b^2}{4c}\Bigr) x_1^2 + c x_2^2
  \qquad (c \neq 0),
\]
and
\[
  a x_1^2 + b x_1x_2
  \;\longmapsto\;
  b x_1x_2
  \qquad (b \neq 0),
\]
together with the diagonal scaling formula. In this round, the same file now
packages the sign cases as explicit normal-form equalities:
\[
  a x_1^2 + c x_2^2 \leadsto a(x_1^2+x_2^2)
  \qquad \text{when } a/c>0,
\]
and
\[
  a x_1^2 + c x_2^2 \leadsto a(x_1^2-x_2^2)
  \qquad \text{when } -a/c>0,
\]
after the scaling \(x_2 \mapsto \sqrt{a/c}\,x_2\) or
\(x_2 \mapsto \sqrt{-a/c}\,x_2\). It also combines these with the shear step,
so the cross-killed diagonal form now normalizes directly to the sum-square or
split-diagonal annihilator whenever the corresponding ratio is positive. The
same Lean file now packages the whole coefficient case split as a single theorem
\texttt{mixedAffineAnnihilator\_normal\_form\_cases}. So the remaining work is
no longer to discover the completion-of-squares algebra, but to connect these
explicit affine moves to a general rank-$4$ mixed-affine factor and then feed
the result into the already-formalized representative theorems.

\medskip
\noindent
There is one useful refinement of that plan. After the shear step, the
surjective non-rank-one case naturally appears as a split diagonal annihilator
\[
  x_1^2 - x_2^2,
\]
whose kernel plane is
\[
  \operatorname{span}(x_1x_2,\ x_1^2+x_2^2).
\]
This plane is not one of the two canonical surjective models already stated
above, but it is linearly equivalent to
\[
  \operatorname{span}(x_1^2,x_2^2)
\]
via the fixed change of variables
\[
  (x_1,x_2) \longmapsto (x_1+x_2,\ x_1-x_2).
\]
Under that map,
\[
  x_1x_2 \longmapsto x_1^2-x_2^2,
  \qquad
  x_1^2+x_2^2 \longmapsto 2(x_1^2+x_2^2),
\]
so any invertible basis of
\[
  \operatorname{span}(x_1x_2,\ x_1^2+x_2^2)
\]
transports to an invertible basis of
\[
  \operatorname{span}(x_1^2,x_2^2).
\]
Therefore the split diagonal case can be discharged by a fixed transport lemma,
and the remaining mixed-affine normalization can stay inside the explicit
translation-shear-scaling formulas already formalized.

\medskip
\noindent
The current plan for that last connection is to work in explicit quadratic
coordinates. Every quadratic polynomial can be written uniquely as
\[
  a_{00} + a_{10} x_1 + a_{01} x_2 + a_{20} x_1^2 + a_{11} x_1x_2 + a_{02} x_2^2.
\]
Inside the affine subgroup preserving the line $\mathbb{R}x_1$, the three basic
moves needed for the mixed-affine branch are:
\[
  x_2 \mapsto x_2 + t,\qquad
  x_2 \mapsto s x_1 + x_2,\qquad
  x_2 \mapsto s x_2.
\]
Their effect on the six coefficients is completely explicit. In particular, the
translation controls the affine tail $(a_{00},a_{10},a_{01})$, while the
shear-and-scale formulas act on the homogeneous triple
\[
  (a_{20},a_{11},a_{02})
\]
by the same completion-of-squares identities already proved in
\texttt{MixedAffineNormalization.lean}. The new file
\texttt{QuadraticCoordinateForm.lean} now also packages this at the level of
arbitrary quadratics: it defines the translation equivalence $x_2 \mapsto x_2+t$
inside the transport framework and proves the generic coefficient identities
\[
  \operatorname{coeff}_{x_2}\bigl(q(x_1,x_2+t)\bigr)
  =
  \operatorname{coeff}_{x_2}(q) + 2 t\, \operatorname{coeff}_{x_2^2}(q),
\]
\[
  \operatorname{coeff}_{x_1x_2}\bigl(q(x_1+t x_2,x_2)\bigr)
  =
  \operatorname{coeff}_{x_1x_2}(q) + 2 t\, \operatorname{coeff}_{x_2^2}(q),
\]
together with the corresponding $x_1^2$, $x_1x_2$, and $x_2^2$ formulas under
translation, shear, and scaling. So the remaining normalization problem
can be reframed as follows: extract from a general constant-containing
$\dim = 2$ factor the relevant coefficient data, kill the unwanted affine tail,
read off the binary-quadratic annihilator of the resulting homogeneous
$2$-plane, and then apply the already-formalized rank-$13$, rank-$14$, or
rank-$15$ plane theorem.

\medskip
\noindent
At this point the missing paper-level step is no longer ``find the right
normalization formulas'' but ``apply the already-formalized formulas with the
right chosen parameters.'' Concretely, for a quadratic
\[
  q = a_{00} + a_{10} x_1 + a_{01} x_2 + a_{20} x_1^2 + a_{11} x_1x_2 + a_{02} x_2^2,
\]
the intended normalization now runs as follows:
\[
  t_{\mathrm{tr}} = -\frac{a_{01}}{2 a_{02}}
  \qquad (a_{02} \neq 0)
\]
kills the $x_2$ coefficient under translation $x_2 \mapsto x_2+t_{\mathrm{tr}}$,
while
\[
  t_{\mathrm{sh}} = -\frac{a_{11}}{2 a_{02}}
  \qquad (a_{02} \neq 0)
\]
kills the $x_1x_2$ coefficient under the shear $x_2 \mapsto x_2+t_{\mathrm{sh}}x_1$,
leaving
\[
  a_{20} - \frac{a_{11}^2}{4 a_{02}}
\]
as the new $x_1^2$ coefficient. In the degenerate case $a_{02}=0$ but
$a_{11}\neq 0$, the parameter
\[
  t_{\mathrm{sh}} = -\frac{a_{20}}{a_{11}}
\]
instead kills the $x_1^2$ coefficient and reduces directly to the pure-cross
orbit. So the remaining formal gap is to convert these scalar coefficient
choices into relation data for the mixed-affine plane theorems.

\section*{Current Candidate Blueprint}
At the level of a paper proof, the present candidate argument is:

\begin{enumerate}
\item If the four coordinates of $u$ are linearly dependent over $\mathbb{R}$,
apply the already-formalized constant-relation lemma.
\item Otherwise let $W = \operatorname{span}(u_1,u_2,u_3,u_4)$ and classify by
the dimension of $W \cap \mathbb{R}[x_1,x_2]_{\le 1}$.
\item If $\dim(W \cap \mathbb{R}[x_1,x_2]_{\le 1}) = 3$, choose coefficient
  vectors spanning $1,x_1,x_2$ inside $W$ and apply the abstract theorem
  \texttt{residual\_eq\_zero\_of\_contains\_aff1}. No exact representative
  normalization is needed in this branch.
\item Otherwise use orthogonal changes in the factor coordinates together with
  affine changes of variables in $(x_1,x_2)$ to reduce the remaining stable
  branches to the following models:
  \[
    (x_1,x_2,x_1^2,x_1x_2),\quad
    (x_1,x_2,x_1^2,x_2^2),\quad
    (1,x_1^2,x_1x_2,x_2^2),\quad
    (x_1,x_1^2,x_1x_2,x_2^2).
  \]
\item In these stable models, show directly that every quadratic square lies in
  \[
    \operatorname{im}(A_u) + \operatorname{cone}(\sigma(\ker A_u))
  \]
  using explicit kernel families:
  \begin{itemize}
  \item $(x_1,x_2,x_1^2,x_1x_2)$:
    \[
      w_c=(-2x_2,x_1,0,1),\qquad
      w_q=(-x_2^2,x_1x_2,0,0);
    \]
  \item $(x_1,x_2,x_1^2,x_2^2)$:
    \[
      w=(-x_1,0,1,0);
    \]
  \item $(1,x_1^2,x_1x_2,x_2^2)$: surjective image, no kernel term needed;
  \item $(x_1,x_1^2,x_1x_2,x_2^2)$:
    \[
      w_\ell=(-x_2\ell,0,\ell,0),\qquad \ell=c+d x_2.
    \]
  \end{itemize}
\item The model
  \[
    (1,x_1^2,x_1x_2,x_2^2)
  \]
  is already formalized: its image is all quartics.
\item Treat the mixed-affine branch by separating the two nonsurjective models
  \[
    (1,x_1,x_1^2,x_1x_2),\qquad
    (1,x_1,x_1x_2,x_2^2),
  \]
  from the surjective rank-$15$ orbits. The representative quotient generators
  are now explicit: the first model uses the family
  \[
    w_\ell=(0,0,-x_2\ell,x_1\ell),\qquad \ell=c+d x_2,
  \]
  the second uses
  \[
    w=(0,0,-x_1x_2,x_1^2),
  \]
  and the hyperbolic rank-$15$ model
  \[
    (1,x_1,x_1^2,x_2^2)
  \]
  is formalized as surjective. The definite rank-$15$ orbit
  \[
    \operatorname{span}(x_1x_2,\ x_1^2-x_2^2)
  \]
  is now formalized as surjective at the plane level as well. The remaining
  mixed-affine gap is no longer the annihilator case split itself, since
  \texttt{MixedAffineNormalization.lean} now packages that as
  \texttt{mixedAffineAnnihilator\_normal\_form\_cases}. What is still missing is
  the written proof that every constant-containing $\dim = 2$ factor supplies
  the coefficient data and relation vectors needed to invoke that theorem and
  then land in one of these formalized plane types.
\item A smaller replacement for exact representative transport is now plausible
  in this branch: prove basis-invariant mixed-affine plane theorems. Instead of
  forcing the last two coordinates to be exactly
  \[
    (x_1^2,x_1x_2)\quad\text{or}\quad(x_1x_2,x_2^2),
  \]
  one can work with an arbitrary orthonormal coefficient-space basis of the
  quadratic plane
  \[
    x_1 \cdot \operatorname{span}(x_1,x_2)
    \qquad\text{or}\qquad
    x_2 \cdot \operatorname{span}(x_1,x_2),
  \]
  and build the kernel family directly from the syzygy
  \[
    (x_1 a)(b \ell) - (x_1 b)(a \ell)=0
  \]
  or its $x_2$-analogue. In these plane-type formulations, the image part
  depends only on the plane itself, while orthonormality restores the clean
  $\sigma = (a^2+b^2)\ell^2$ identity used in the representative proofs. The
  current Lean support for this approach is now in place: the new file
  \texttt{RepresentativeMixedAffinePlane.lean} proves that any explicit scalar
  relation $\sum c_i u_i=r$ yields an admissible image witness for $rq$, that
  two orthonormal coefficient-space relation directions satisfy
  \[
    \sigma\bigl(\mathrm{rel}_c(p)+\mathrm{rel}_d(q)\bigr)=p^2+q^2.
  \]
  More importantly, it now abstracts the whole image half of the rank-$14$
  mixed-affine proof: if a factor admits explicit relations for
  \[
    1,\qquad x_1,\qquad x_1x_2,\qquad x_2^2,
  \]
  then every quartic with zero $x_1^4$ coefficient lies in its admissible
  image. The certificate half is now abstracted too: if those same relations
  are accompanied by an orthonormal coefficient-space pair for the
  $(x_1x_2,x_2^2)$ directions, then the formalized kernel
  \[
    w=\mathrm{rel}_{-c_2}(t x_1x_2)+\mathrm{rel}_{c_3}(t x_1^2)
  \]
  matches the missing $x_1^4$ coefficient and forces zero residual for every
  SOCP. So the missing step is no longer the monomial image or kernel algebra
  itself, but the extraction of the correct canonical relation data from an
  arbitrary factor.
\item The rank-$13$ mixed-affine branch is now abstracted in the same style.
  The new plane-level theorem only assumes explicit scalar relations for
  \[
    1,\qquad x_1^2,\qquad x_1x_2,
  \]
  plus an orthonormal coefficient-space pair for the last two directions. Under
  those hypotheses, every quartic with vanishing $x_2^3$ and $x_2^4$
  coefficients lies in the admissible image, and the formalized kernel
  \[
    w=\mathrm{rel}_{-c_2}(x_2\ell)+\mathrm{rel}_{c_3}(x_1\ell),
    \qquad \ell=b+c x_2,
  \]
  matches the missing $(x_2^3,x_2^4)$ classes and forces zero residual for
  every SOCP. So both nonsurjective mixed-affine representatives are now
  available as basis-invariant plane certificates rather than only as concrete
  coordinate models.
\item The old stronger-basis attempt has now split into two clean formalized
  layers. First, \texttt{RepresentativeMixedAffinePlane.lean} now proves exact
  surjective plane theorems for any invertible basis of
  \[
    \operatorname{span}(x_1^2,x_2^2)
    \qquad\text{and}\qquad
    \operatorname{span}(x_1x_2,\ x_1^2-x_2^2),
  \]
  together with algebra-equivalence transport wrappers. Second, for the two
  nonsurjective planes
  \[
    \operatorname{span}(x_1x_2,\ x_2^2)
    \qquad\text{and}\qquad
    \operatorname{span}(x_1^2,\ x_1x_2),
  \]
  the file first proved orthogonal in-plane basis reductions back to the exact
  rank-$14$ and rank-$13$ certificates, and now those same nonsurjective
  theorems have been strengthened to allow orthogonal but not necessarily unit
  column data, again wrapped through arbitrary polynomial algebra
  equivalences. So the remaining gap has tightened again: one no longer needs
  to normalize the column lengths inside the mixed-affine plane, but one still
  needs the normalization/classification step to produce from a general
  mixed-affine factor either an invertible surjective plane basis or an
  orthonormal relation basis whose polynomial images lie in one of the two
  canonical nonsurjective planes. The intended bridge is now an explicit
  six-coefficient quadratic normal form plus formulas for the
  $x_2$-translation, $x_2$-shear, and $x_2$-scaling subgroup preserving
  $\mathbb{R}x_1$.

\medskip
\noindent
The sharper nonsurjective-plane picture has now partially been formalized. For
the rank-$14$ plane \(\operatorname{span}(x_1x_2,x_2^2)\), the Lean file now
proves the stronger determinant-based theorem:
\begin{itemize}
\item any orthonormal relation basis with polynomial images
  \[
    x_2(a x_1 + b x_2),\qquad x_2(c x_1 + d x_2),
  \]
  and
  \[
    \det\!\begin{pmatrix}a&b\\ c&d\end{pmatrix}\neq 0
  \]
  already
  forces zero residual;
\item the proof reconstructs the exact \(x_1x_2\) and \(x_2^2\) relations by
  the inverse \(2\times 2\) matrix and uses the kernel syzygy
  \[
    -x_1(c x_1+d x_2)\cdot x_2(a x_1+b x_2)
    +x_1(a x_1+b x_2)\cdot x_2(c x_1+d x_2)=0
  \]
  whose square still misses only the \(x_1^4\) coefficient, now scaled by
  \(a^2+c^2\neq 0\).
\end{itemize}
So the rank-$14$ branch no longer needs any extra polynomial-column
orthogonality inside the plane. The analogous rank-$13$ strengthening is now
formalized as well:
\begin{itemize}
\item for the rank-$13$ plane \(\operatorname{span}(x_1^2,x_1x_2)\), any
  orthonormal relation basis with polynomial images
  \[
    x_1(a x_1+b x_2),\qquad x_1(c x_1+d x_2)
  \]
  and
  \[
    \det\!\begin{pmatrix}a&b\\ c&d\end{pmatrix}\neq 0
  \]
  already forces zero residual;
\item the proof uses the kernel
  \[
    \mathrm{rel}_{-c_2}\bigl((c x_1+d x_2)(\beta+\gamma x_2)\bigr)
    + \mathrm{rel}_{c_3}\bigl((a x_1+b x_2)(\beta+\gamma x_2)\bigr),
  \]
  because
  \[
    -x_1(a x_1+b x_2)\cdot (c x_1+d x_2)(\beta+\gamma x_2)
    +x_1(c x_1+d x_2)\cdot (a x_1+b x_2)(\beta+\gamma x_2)=0.
  \]
  Its square still affects only the pure \(x_2^3\) and \(x_2^4\)
  coefficients modulo the already-formalized image, now scaled by
  \(b^2+d^2\neq 0\).
\end{itemize}
So neither nonsurjective mixed-affine plane now needs an extra polynomial-column
orthogonality condition. The normalization only has to produce the canonical
plane type and a nonzero determinant inside that plane, not a special
orthogonal basis of polynomial columns. Both determinant-based theorems are
also wrapped through the affine transport layer, so once a normalized factor
lands in either canonical plane type, the result pulls back directly to the
original SOCP point.
\item Apply the formalized admissible image-plus-cone certificate to conclude
$\sigma(u)=p$.
\end{enumerate}

\noindent
The main obstacle to a finished Lean proof is now precise: complete the
normalization/classification of the constant-containing $\dim = 2$ branch, then
assemble the final theorem from the already-formalized representative and
abstract branch theorems.

\medskip
\noindent
That six-monomial normal form is now formalized in
\texttt{TernaryQuarticProof/QuadraticNormalForm.lean}. It introduces the six
quadratic monomial exponents
\[
  1,\ x_1,\ x_2,\ x_1^2,\ x_1x_2,\ x_2^2
\]
as explicit \texttt{Fin 2 ->0 Nat} indices, proves that every support monomial of
an \texttt{IsQuadratic} polynomial is one of them, and concludes that every
quadratic is the sum of its six monomial-coefficient terms. This is still a
proof-serving finite-support lemma rather than a general coefficient API.

\medskip
\noindent
The mixed-affine analysis is now complete at the representative-or-plane level:
the rank-$13$ and rank-$14$ nonsurjective models are formalized both as concrete
representatives and as abstract plane certificates, now in a form that also
accepts non-unit orthogonal in-plane coefficient data and transports directly
across affine algebra equivalences, and the surjective rank-$15$ branch is
formalized for both real plane orbits
\[
  \operatorname{span}(x_1^2,x_2^2)
  \qquad\text{and}\qquad
  \operatorname{span}(x_1x_2,\ x_1^2-x_2^2).
\]
The new bridge layer is also formalized now: using the six-coefficient
quadratic normal form, Lean proves that a homogeneous quadratic satisfying one
of the canonical coefficient conditions
\[
  \operatorname{coeff}_{x_1x_2}=0,\qquad
  \operatorname{coeff}_{x_1^2}=0,\qquad
  \operatorname{coeff}_{x_2^2}=0,\qquad
  \operatorname{coeff}_{x_1^2}\pm \operatorname{coeff}_{x_2^2}=0
\]
already has the corresponding canonical plane expansion. The mixed-affine plane
file now wraps those coefficient lemmas into generic relation theorems, so once
normalization produces homogeneous quadratics in one of those coefficient
orbits, the existing rank-$13$, rank-$14$, or rank-$15$ plane certificates
apply directly.
So the remaining obstacle is no longer a missing mixed-affine representative
calculation. One correction is needed here: the polynomial identity
\texttt{mixedAffineAnnihilator\_normal\_form\_cases} describes the covariant
action of shears and rescalings on the annihilator polynomial
\(a x_1^2 + b x_1x_2 + c x_2^2\), while the actual classification of the
mixed-affine branch is governed by the dual action on the homogeneous
coefficient functionals
\[
  (\operatorname{coeff}_{x_1^2},\ \operatorname{coeff}_{x_1x_2},\
  \operatorname{coeff}_{x_2^2}).
\]
That dual action is now formalized in
\texttt{QuadraticCoordinateForm.lean}, and
\texttt{RepresentativeMixedAffinePlane.lean} already uses it in the two new
transport wrappers
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_const\_homQuadratics\_crossAnnihilator}
and
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_const\_x0\_homQuadratics\_annihilator\_disc\_zero}.
That discriminant-zero wrapper has now been strengthened further: it no longer
assumes the last two relation directions are already orthonormal in coefficient
space. A nonzero Gram determinant
\[
  \|c_2\|^2 \|c_3\|^2 - \langle c_2,c_3\rangle^2 \neq 0
\]
is enough, because \texttt{RepresentativeMixedAffinePlane.lean} now performs the
normalization to an orthonormal pair internally before invoking the canonical
rank-$14$ plane theorem.
The same file now also handles the purely diagonal dual relation
\(a\,\operatorname{coeff}_{x_1^2} + d\,\operatorname{coeff}_{x_2^2}=0\) by a
single \(x_2\)-scaling, via the wrappers
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_const\_homQuadratics\_diagAnnihilator\_sum}
and
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_const\_homQuadratics\_diagAnnihilator\_diff},
which reduce directly to the already-formalized sum-square and split-diagonal
plane theorems.
Moreover the same file now composes the cross-killing shear with these diagonal
scaling wrappers. The resulting theorems
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_const\_homQuadratics\_annihilator\_sum}
and
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_const\_homQuadratics\_annihilator\_diff}
start from a genuinely general dual annihilator
\[
  a\,\operatorname{coeff}_{x_1^2} +
  b\,\operatorname{coeff}_{x_1x_2} +
  c\,\operatorname{coeff}_{x_2^2} = 0
\]
with the appropriate sign on \(c - b^2/a\), and reduce it to the already
formalized rank-$15$ sum-square or split-diagonal planes. So the remaining
mixed-affine obstacle is now entirely the extraction/classification statement,
not another missing normal-form transport.
One more theorem-level reduction is now formalized in the same file:
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_homQuadratics\_plane\_nontrivial}.
Given explicit relations to \(1\), \(x_1\), and two homogeneous quadratics
\(q_2,q_3\), it forms the canonical dual annihilator directly from the
\(2\times 3\) coefficient matrix of \((q_2,q_3)\),
\[
  a = \det\!\begin{pmatrix}
    \operatorname{coeff}_{x_1x_2}(q_2) & \operatorname{coeff}_{x_2^2}(q_2) \\
    \operatorname{coeff}_{x_1x_2}(q_3) & \operatorname{coeff}_{x_2^2}(q_3)
  \end{pmatrix},
  \quad
  b = \det\!\begin{pmatrix}
    \operatorname{coeff}_{x_1^2}(q_2) & \operatorname{coeff}_{x_2^2}(q_2) \\
    \operatorname{coeff}_{x_1^2}(q_3) & \operatorname{coeff}_{x_2^2}(q_3)
  \end{pmatrix},
  \quad
  c = \det\!\begin{pmatrix}
    \operatorname{coeff}_{x_1^2}(q_2) & \operatorname{coeff}_{x_1x_2}(q_2) \\
    \operatorname{coeff}_{x_1^2}(q_3) & \operatorname{coeff}_{x_1x_2}(q_3)
  \end{pmatrix},
\]
and then dispatches automatically to the rank-$13$, rank-$14$, or rank-$15$
mixed-affine plane theorem according to whether \(a\), \(b\), and
\(c-b^2/a\) vanish or change sign.
However, a new coefficient-only Julia probe now shows that this is still not
the generic shape of the constant-containing mixed-affine branch before the
final normalization step. The new script
\texttt{julia/mixed\_affine\_tail\_dim\_2026\_04\_10.jl} perturbs each canonical
homogeneous pair by random affine tails and finds, in $20/20$ samples for each
of the rank-$13$, rank-$14$, and rank-$15$ bases, that
\[
  \dim(W \cap \mathbb{R}[x_1,x_2]_{\le 1}) = 2
  \qquad\text{but}\qquad
  \dim(W \cap \mathbb{R}[x_1,x_2]^{\mathrm{hom}}_2) = 1.
\]
So the next bridge should not expect a full $2$-dimensional homogeneous
quadratic plane to be present already. The generic target is instead:
\[
  1,\ x_1,\ h,\ q \in W,
\]
where \(h\) is homogeneous and \(q\) has one remaining affine tail. The
translation formulas now formalized in
\texttt{QuadraticCoordinateForm.lean}, including the new explicit
\(\operatorname{coeff}_{1}\) and \(\operatorname{coeff}_{x_1}\) identities under
\(x_2 \mapsto x_2+t\), are exactly the API needed to turn that one-line-plus-tail
configuration into the homogeneous-plane theorem above when the tail lies in the
right orbit. More concretely, the new Lean definition
\[
  H(q_2,q_3)
  :=
  \operatorname{coeff}_{x_2}(q_3)\,q_2
  -
  \operatorname{coeff}_{x_2}(q_2)\,q_3
\]
now canonically extracts the unique homogeneous quadratic line from any pair of
quadratics with zero constant and \(x_1\) coefficients: the
\(\operatorname{coeff}_{x_2}\) term cancels by construction, and the resulting
\(x_1^2\), \(x_1x_2\), and \(x_2^2\) coefficients are explicit \(2\times2\)
minors of the original coefficient matrix.
There is now a second concrete reduction at the written and numerical level as
well. The new Julia probe
\texttt{julia/mixed\_affine\_common\_translation\_2026\_04\_10.jl} isolates the
determinant-nondegenerate sub-branch and tests the explicit two-parameter
translation
\((x_1,x_2) \mapsto (x_1+u,x_2+v)\) predicted by the coefficient algebra, with
the identity
\[
  \operatorname{coeff}_{x_2}(q(\,\cdot+u,\cdot+v))
  =
  \operatorname{coeff}_{x_2}(q)
  +
  u\,\operatorname{coeff}_{x_1x_2}(q)
  +
  2v\,\operatorname{coeff}_{x_2^2}(q).
\]
So whenever two tailed quadratics \(q_2,q_3\) satisfy
\[
  \operatorname{coeff}_{x_1x_2}(q_2)\operatorname{coeff}_{x_2^2}(q_3)
  -
  \operatorname{coeff}_{x_2^2}(q_2)\operatorname{coeff}_{x_1x_2}(q_3)
  \neq 0,
\]
there is an explicit common translation killing both \(x_2\)-tails
simultaneously. Numerically this is now confirmed by
\texttt{julia/mixed\_affine\_common\_translation\_2026\_04\_10.jl}, which gives
tail residuals at the \(10^{-14}\) scale on random samples.
The minimal Lean algebra behind that statement is now formalized in
\texttt{QuadraticCoordinateForm.lean}: the file now contains the
\(x_1 \mapsto x_1+t\) and \(x_2 \mapsto x_2+t\) translation equivalences, the
exact \(x_2\)-tail coefficient formulas for each, and the nested two-step lemma
showing that the explicit Cramér-rule choice of \((u,v)\) kills the
\(\operatorname{coeff}_{x_2}\) term of both quadratics under the
\(x_1\)-then-\(x_2\) translation. It also now records the induced
\(\operatorname{coeff}_{1}\) and \(\operatorname{coeff}_{x_1}\) correction
terms after that same nested translation, so the affine part that must be
removed using the existing \(1\) and \(x_1\) relations is explicit in Lean.
That whole coordinate-level bridge is now formalized in
\texttt{RepresentativeMixedAffinePlane.lean} as
\[
  \texttt{residual\_eq\_zero\_of\_const\_x0\_tailedPair\_det},
\]
which proves that for
\[
  u=(1,x_1,q_2,q_3)
\]
the determinant condition
\[
  \operatorname{coeff}_{x_1x_2}(q_2)\operatorname{coeff}_{x_2^2}(q_3)
  -
  \operatorname{coeff}_{x_2^2}(q_2)\operatorname{coeff}_{x_1x_2}(q_3)
  \neq 0
\]
already implies \(\operatorname{residual}(p,u)=0\) for every SOCP. The proof
applies the explicit common translation, subtracts the induced \(1\)- and
\(x_1\)-corrections, computes the resulting Gram determinant in coefficient
space, and then dispatches to the already-formalized theorem
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_homQuadratics\_plane\_nontrivial}.
The same file also now contains the affine-equivalence wrapper
\texttt{residual\_eq\_zero\_of\_equiv\_const\_x0\_tailedPair\_det}, so this
certificate can be applied after a prior affine normalization step. It also now
contains the transport-shaped theorem
\texttt{residual\_eq\_zero\_of\_socp\_of\_eq\_mix\_affineEquiv\_const\_x0\_tailedPair\_det},
which packages the same branch in the exact ``affine normalization plus
orthogonal factor mixing'' form used by \texttt{RepresentativeTransport.lean}.
Finally, the same branch is now lifted one more step to explicit relation data:
\texttt{RepresentativeMixedAffinePlane.lean} now proves
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_tailedPair\_det} and its
equivariant version
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_const\_x0\_tailedPair\_det}.
The remaining interface mismatch on the mixed-affine side is now also gone for
the homogeneous-plane subcase itself. The same Lean file now exports
\texttt{residual\_eq\_zero\_of\_equiv\_const\_x0\_homQuadratics\_plane\_nontrivial}
and the transport-shaped theorem
\texttt{residual\_eq\_zero\_of\_socp\_of\_eq\_mix\_affineEquiv\_const\_x0\_homQuadratics\_plane\_nontrivial}.
So both constant-containing mixed-affine outputs now sit at the same top-level
surface: the classification step may hand back exact normalized coordinates
\[
  \operatorname{mix}(M^\top,e^{-1}\!\cdot u)=(1,x_1,q_2,q_3)
\]
either with a nontrivial homogeneous quadratic plane, or with a determinant
condition on a tailed pair, and the final transport layer can consume either
form directly without repackaging it into relation vectors first.
So once four coefficient vectors \(c_0,c_1,c_2,c_3 \in \mathbb{R}^4\) are
orthonormal and satisfy the explicit relations
\[
  \sum_i (c_0)_i u_i = 1,\qquad
  \sum_i (c_1)_i u_i = x_1,\qquad
  \sum_i (c_2)_i u_i = q_2,\qquad
  \sum_i (c_3)_i u_i = q_3,
\]
the tailed-pair branch no longer needs any hand-built \(4\times4\) matrix:
those theorems internally package the relation vectors into an orthogonal
\texttt{mix} matrix and invoke the transport form automatically.
The next numerical probe sharpened the still-open determinant-zero side of the
same mixed-affine branch. The new pure-linear-algebra script
\texttt{julia/mixed\_affine\_det\_zero\_2026\_04\_10.jl} tests the canonical
two-parameter families
\[
  (1,x_1,\mu x_1^2 + D,\nu x_1^2 + x_2)
\]
after the usual basis choice in the tailed two-plane, with
\[
  D \in \{x_1x_2,\ x_2^2,\ x_1^2+x_2^2,\ x_1^2-x_2^2\}.
\]
Across the sampled values \((\mu,\nu) \in
\{(-2,-1),(-1,1),(0,1),(1,1),(2,-1)\}\), the image-rank picture is completely
stable:
\[
  D=x_1x_2 \Longrightarrow \operatorname{rank}(A_u)=14
  \text{ with only } x_2^4 \text{ missing,}
\]
while
\[
  D \in \{x_2^2,\ x_1^2+x_2^2,\ x_1^2-x_2^2\}
  \Longrightarrow \operatorname{rank}(A_u)=15.
\]
The follow-up slice probe
\texttt{julia/mixed\_affine\_det\_zero\_slice\_2026\_04\_10.jl} sharpens that
further. For \(80\) random determinant-zero tailed families with completely
generic homogeneous line \(D=a x_1^2 + b x_1x_2 + c x_2^2\), every sample had
\(\operatorname{rank}(A_u)=15\). On the codimension-one slice \(c=0\) with
\(b \neq 0\), every sample had rank \(14\). On the slices \(b=0\) and
\(D=(u x_1 + v x_2)^2\), every sample had rank \(15\). So the
determinant-zero tailed branch no longer looks like an arbitrary new family.
Numerically it now appears to split by a much smaller invariant: the only
stable nonsurjective locus is the cross-type slice where the extracted
homogeneous line has vanishing \(x_2^2\)-coefficient after the canonical basis
choice, while the generic, \(x_2^2\)-type, and diagonal-type slices are already
surjective.
The new script
\texttt{julia/mixed\_affine\_det\_zero\_canonical\_a\_sweep\_2026\_04\_10.jl}
pushes that picture a bit further. It fixes the canonical families
\[
  (1,x_1,D,x_2 + a x_1^2)
\]
and varies \(a \in \{-1,1,2\}\). The result is again completely stable: the
cross family always has rank \(14\) with only \(x_2^4\) missing, while the
\(x_2^2\), sum-square, and split-diagonal families always have rank \(15\) with
no missing monomials. So the evidence is now tail-parameter invariant on those
canonical models, not just a one-off rank computation.
The new script
\texttt{julia/mixed\_affine\_det\_zero\_diag\_tail\_2026\_04\_10.jl} isolates
the simplest two-parameter diagonal family
\[
  (1,x_1,d x_1^2 + x_2^2,\ x_2 + a x_1^2)
\]
with \(d \in \{-2,-1,0,1,2\}\) and \(a \in \{-2,-1,1,2\}\). Every one of the
resulting \(20\) samples again has \(\operatorname{rank}(A_u)=15\) with no
missing quartic monomials. So the numerical picture is now even sharper: once
the extracted homogeneous line has a nonzero \(x_2^2\)-coefficient after the
canonical determinant-zero basis choice, the tailed branch looks uniformly
surjective, and only the cross-type slice with vanishing \(x_2^2\)-coefficient
still behaves like a genuine rank-\(14\) obstruction.
There is now also a small formal reduction on the Lean side that removes one
easy slice of this branch. In
\texttt{RepresentativeMixedAffinePlane.lean}, the new theorems
\texttt{relation\_x1\_of\_affineTail},
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_affineTail}, and
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_const\_x0\_affineTail}
show that whenever the fourth coordinate is already affine in \(x_1\) with
nonzero \(x_1\)-coefficient, the branch collapses immediately to the generic
\(\dim(W \cap \mathrm{Aff}_1)=3\) theorem by reconstructing an explicit
\(x_1\) relation. That generic branch is now also on the same top transport
surface as the others: \texttt{RepresentativeSpanThree.lean} now proves
\texttt{residual\_eq\_zero\_of\_socp\_of\_eq\_mix\_affineEquiv\_contains\_aff1},
so exact normalized coordinates of the form
\[
  \operatorname{mix}(M^\top,e^{-1}\!\cdot u)=(1,x_1,x_2,q)
\]
are enough to close the whole \(\dim(W \cap \mathrm{Aff}_1)=3\) branch
directly. There is now also a representative-level surjective theorem
in \texttt{RepresentativeMixedAffine.lean} for the first genuinely non-affine
determinant-zero family:
\[
  (1,x_1,x_2^2,x_2 + a x_1^2), \qquad a \neq 0.
\]
The new definitions and theorems
\texttt{mixedAffineTailX1SqRep},
\texttt{quartic\_in\_image\_mixedAffineTailX1SqRep}, and
\texttt{residual\_eq\_zero\_mixedAffineTailX1SqRep}
prove directly that every quartic lies in the image there. The proof is by an
explicit monomial check: every quartic monomial is immediate except
\(x_1^4\) and \(x_1^3x_2\), and those two come from the tailed coordinate by a
single correction against the \(x_1\) and \(x_2^2\) coordinates. Combined with
the affine-tail reduction, this means the entire \(x_2^2\)-type
determinant-zero family is already formalized at representative level. The new
exact relation theorems
\texttt{quartic\_in\_image\_of\_relations\_const\_x0\_x1sq\_x1PlusX0sq},
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_x1sq\_x1PlusX0sq}, and
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_const\_x0\_x1sq\_x1PlusX0sq}
lift that slice out of the concrete representative: once affine normalization
produces exact relations to \(1\), \(x_1\), \(x_2^2\), and \(x_2 + a x_1^2\),
the branch closes directly, and the final pullback through an algebra
equivalence is already packaged.
The second remaining determinant-zero candidate, the cross-type model
\[
  (1,x_1,x_1x_2,x_2 + a x_1^2), \qquad a \neq 0,
\]
is now also formalized in \texttt{RepresentativeMixedAffine.lean} as
\texttt{mixedAffineTailCrossRep}, with
\texttt{quartic\_in\_image\_mixedAffineTailCrossRep\_of\_coeff\_m04\_zero},
\texttt{mixedAffineTailCrossKerBase}, and
\texttt{residual\_eq\_zero\_mixedAffineTailCrossRep}. This is a rank-\(14\)
certificate of exactly the same shape as the already-solved codimension-one
mixed-affine representative: every quartic with zero \(x_2^4\) coefficient is
in the image, and the missing \(x_2^4\) class is supplied by a single explicit
kernel square. The same slice is now also lifted one step above the rigid
representative in \texttt{RepresentativeMixedAffinePlane.lean} via
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_x0x1\_x1PlusX0sq\_orthonormal}
and its equivariant version
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_const\_x0\_x0x1\_x1PlusX0sq\_orthonormal}.
There is now a further exact-relation lift in
\texttt{RepresentativeMixedAffine.lean}:
\texttt{quartic\_in\_image\_of\_relations\_const\_x0\_x0x1\_x1PlusX0sq\_of\_coeff\_m04\_zero},
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_x0x1\_x1PlusX0sq}, and
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_const\_x0\_x0x1\_x1PlusX0sq}.
These remove the orthonormality requirement from the usable theorem layer for
the cross slice: once affine normalization produces exact relations to
\(1\), \(x_1\), \(x_1x_2\), and \(x_2 + a x_1^2\), the same image-plus-kernel
argument closes directly. The key point is that the missing \(x_2^4\) class is
still supplied by a single explicit kernel direction, and its
\(\operatorname{coeff}_{x_2^4}\) depends only on the constant relation vector,
so no full coefficient-space orthonormal basis is needed at that stage. The
canonical determinant-zero mixed-affine branch is therefore no longer missing
any representative theorem, and the cross slice is no longer stuck even at the
orthonormal-relation level.
There is now one more useful lift in
\texttt{RepresentativeMixedAffinePlane.lean}:
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_crossTail} and its
equivariant version
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_const\_x0\_crossTail}. These
start from coefficient constraints rather than already-normalized exact images:
\(q_2\) is assumed to be a pure cross line, while \(q_3\) is assumed to have
zero \(x_2^2\)-coefficient and nonzero \(x_1\)- and \(x_1^2\)-coefficients.
Lean then performs the normalization internally: it rescales \(q_2\) to
\(x_1x_2\), subtracts the cross part from \(q_3\), rescales the remaining
tailed coordinate to \(x_2 + a x_1^2\), and dispatches to the exact relation
theorem above. So even the determinant-zero cross slice no longer needs the
classification statement to hand back fully normalized coordinates; a
coefficient-level pure-cross-plus-tail description now suffices.
The same file now also contains the symmetric square-line wrapper
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_x1sqTail} and its
equivariant version
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_const\_x0\_x1sqTail}. These
start from exact relations to \(1\), \(x_1\), a pure \(x_2^2\) line, and a
quadratic with zero \(x_1x_2\)-coefficient and nonzero \(x_2\)- and
\(x_1^2\)-coefficients. Lean then rescales the \(x_2^2\) line to the canonical
coordinate, removes the \(x_2^2\)-component from the tailed quadratic, rescales
what remains to \(x_2 + a x_1^2\), and invokes the already-formalized
\(x_2^2\)-type determinant-zero theorem in
\texttt{RepresentativeMixedAffine.lean}. So both nonsurjective
determinant-zero mixed-affine slices now have coefficient-level wrappers above
the exact normalized relation layer.
The next Lean step then covered the first genuinely diagonal determinant-zero
family at theorem level, not just numerically. In
\texttt{RepresentativeMixedAffine.lean} the new exact theorems
\texttt{quartic\_in\_image\_of\_relations\_const\_x0\_diagX1sq\_x1PlusX0sq},
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_diagX1sq\_x1PlusX0sq},
and
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_const\_x0\_diagX1sq\_x1PlusX0sq}
show directly that exact relations to
\[
  (1,\;x_1,\; d x_1^2 + x_2^2,\; x_2 + a x_1^2), \qquad a \neq 0,
\]
already force zero residual. The proof is again by explicit image generation:
the quartic monomials are produced in a short chain
\(x_1^4 \rightarrow x_1^3x_2 \rightarrow x_1^2x_2^2 \rightarrow x_1x_2^3
\rightarrow x_2^4\), with the lower-degree monomials supplied by the
\(x_1\)-relation. On top of that,
\texttt{RepresentativeMixedAffinePlane.lean} now contains the coefficient-level
wrapper
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_diagX1sqTail} and its
equivariant version
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_const\_x0\_diagX1sqTail}.
These start only from the coefficient conditions that \(q_2\) is a diagonal
homogeneous line with nonzero \(x_2^2\)-coefficient and that \(q_3\) has zero
\(x_1x_2\)- and \(x_2^2\)-coefficients but nonzero \(x_2\)- and
\(x_1^2\)-coefficients. Lean then rescales internally to the exact diagonal
model above and dispatches to the new exact theorem. So the diagonal
determinant-zero slice is no longer only a numerical byproduct of the
classification; one full theorem-level interface now exists for it too.
The remaining obstacle is now the written and formal extraction statement
showing that every constant-containing
$\dim(W \cap \mathbb{R}[x_1,x_2]_{\le 1}) = 2$ factor reduces to one of these
already-formalized plane types and supplies a nontrivial homogeneous quadratic
line together with the extra affine-tail relation needed to invoke this
canonical cross-product theorem. In other words, the next bridge no longer has
to \emph{find} the homogeneous line; it only has to classify how the remaining
tailed quadratic sits relative to that explicit extracted line, and in the
nondegenerate determinant sub-branch it only has to extract exact normalized
coordinates of the form \((1,x_1,q_2,q_3)\), or equivalently exact relation
data for that same tuple, together with the same determinant hypothesis after
affine normalization. So the remaining gap there is no longer the translation
algebra, the cross-slice kernel certificate, or their integration with the
mixed-affine theorem layer; it is the
upstream normalization/classification step producing that coordinate picture
from a general admissible factor. The diagonal
determinant-zero slices no longer look like separate obstructions; if they need
additional Lean work at all, it should only be because they simplify the final
classification step, not because the numerics suggest another missing case.

\medskip
\noindent
The low-affine representative analysis is now complete too in
\texttt{TernaryQuarticProof/RepresentativeLowAffine.lean}. The three stable
models
\[
  (x_1,x_2,x_1^2,x_2^2),\qquad
  (x_1,x_2,x_1^2,x_1x_2),\qquad
  (x_1,x_1^2,x_1x_2,x_2^2)
\]
are all formalized at the representative level. The first two use explicit
codimension-one or codimension-two image-plus-kernel corrections at the
constant and quartic-end classes, while the third uses the affine kernel line
\[
  w_\ell = (-x_2\ell,0,\ell,0), \qquad \ell = c + d x_2,
\]
to match the missing quotient classes of $1$ and $x_2$. So the only remaining
representative-level model outside the mixed-affine normalization problem is
the generic
\[
  (1,x_1,x_2,h)
\]
branch.

\medskip
\noindent
This low-affine upgrade needed one correction. The original writeup treated
only two \(\dim=2\) no-constant branches as canonical, but the exploration log
already contained the additional rank-\(14\) definite plane
\[
  \operatorname{span}(x_1x_2,\ x_1^2-x_2^2),
\]
for example through the probe
\[
  (x_1,x_2,x_1^2-x_2^2,x_1x_2),
\]
which numerically has image rank \(14\). The file
\texttt{TernaryQuarticProof/RepresentativeLowAffine.lean} no longer stops at
the exact representatives
\[
  (x_1,x_2,x_1^2,x_2^2)
  \qquad\text{and}\qquad
  (x_1,x_2,x_1^2,x_1x_2).
\]
It now proves basis-invariant plane theorems from explicit relations to
\(x_1,x_2\) together with any invertible basis of the canonical quadratic
planes
\[
  \operatorname{span}(x_1^2,x_2^2)
  \qquad\text{and}\qquad
  \operatorname{span}(x_1^2,x_1x_2).
\]
In addition, the new exact-relation theorem
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_x0x1\_diffsq}
and its determinant plane wrapper
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_x0x1\_diffsqPlane}
now formalize the definite rank-\(14\) orbit
\[
  \operatorname{span}(x_1x_2,\ x_1^2-x_2^2)
\]
as well. The split rank-\(14\) orbit
\[
  \operatorname{span}(x_1x_2,\ x_1^2+x_2^2)
\]
is now packaged by the fixed transport theorem
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_x0x1\_sumsqPlane},
which sends it to the already-formalized coprime plane
\(\operatorname{span}(x_1^2,x_2^2)\). So the final normalization step no
longer needs exact representative equalities in the low-affine \(\dim=2\)
cases either: it only has to recover the linear part and the resulting plane
type.

The remaining uncertainty in that low-affine \(\dim=2\) picture is now much
smaller numerically than it was earlier in the project. The new script
\texttt{julia/low\_affine\_hom\_plane\_2026\_04\_10.jl} samples twenty random
normalized homogeneous planes
\[
  (x_1,x_2,q_2,q_3),
  \qquad q_2,q_3 \in \operatorname{span}(x_1^2,x_1x_2,x_2^2),
  \qquad \dim \operatorname{span}(q_2,q_3)=2,
\]
and every sampled case has image rank \(14\) with only the constant monomial
missing. The same script also records the canonical controls:
\[
  \operatorname{span}(x_1^2,x_1x_2)
  \quad\leadsto\quad \text{rank }13,
  \qquad
  \operatorname{span}(x_1^2,x_2^2),\ 
  \operatorname{span}(x_1x_2,x_1^2-x_2^2)
  \quad\leadsto\quad \text{rank }14.
\]
So the written proof target is now sharper: the genuinely exceptional
low-affine plane still looks like the common-factor slice
\(\operatorname{span}(x_1^2,x_1x_2)\), while the generic homogeneous plane
appears to collapse to the already-formalized rank-\(14\) behavior.

\medskip
\noindent
The linear low-affine model has now also been lifted through the same
\(3\times 3\) homogeneous reconstruction. The same file no longer uses only
the rigid representative
\[
  (x_1,x_1^2,x_1x_2,x_2^2),
\]
but proves the exact-relation theorem for any admissible factor $u$ with
coefficient vectors satisfying
\[
  \sum_i c_{0,i}u_i = x_1,\qquad
  \sum_i c_{1,i}u_i = x_1^2,\qquad
  \sum_i c_{2,i}u_i = x_1x_2,\qquad
  \sum_i c_{3,i}u_i = x_2^2.
\]
The image reduction is now abstract at that level, and the kernel-line
argument has been lifted to the same exact relation data. The new
\texttt{exact\_relations\_of\_homQuadBasis\_det} lemma now reconstructs those
exact canonical relations from any independent homogeneous quadratic triple, so
the linear low-affine branch also has a basis-invariant homogeneous-basis
theorem and an affine-equivalence transport wrapper. It now also has the
transport-shaped theorem
\texttt{residual\_eq\_zero\_of\_socp\_of\_eq\_mix\_affineEquiv\_x0\_homQuadBasis\_det}.
So if the final classification step returns exact normalized coordinates
\[
  \operatorname{mix}(M^\top,e^{-1}\!\cdot u)=(x_1,q_1,q_2,q_3)
\]
with each \(q_j\) lying in the homogeneous quadratic span
\(\operatorname{span}(x_1^2,x_1x_2,x_2^2)\) and with nonzero coefficient
matrix determinant, this low-affine linear branch can now be consumed directly
at the same top transport layer as the mixed-affine branches.

The constant \(\dim(W \cap \mathrm{Aff}_1)=1\) branch now sits on that same
surface too. In \texttt{RepresentativeSurjective.lean}, the new theorems
\texttt{residual\_eq\_zero\_of\_socp\_of\_eq\_mix\_mapVec\_const\_homQuadBasis\_det}
and
\texttt{residual\_eq\_zero\_of\_socp\_of\_eq\_mix\_affineEquiv\_const\_homQuadBasis\_det}
consume exact normalized coordinates
\[
  \operatorname{mix}(M^\top,e^{-1}\!\cdot u)=(1,q_1,q_2,q_3)
\]
whenever \((q_1,q_2,q_3)\) is a nondegenerate homogeneous quadratic basis.
So both \(\dim(W \cap \mathrm{Aff}_1)=1\) branches are now available in the
same transport-shaped form.

\medskip
\noindent
The determinant-zero mixed-affine interface has now been lifted one more step
as well. The file \texttt{RepresentativeMixedAffinePlane.lean} now contains the
new theorems
\[
  \texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_mixedAffineTailHomLine\_crossTail},
\]
\[
  \texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_mixedAffineTailHomLine\_x1sqTail},
\]
and
\[
  \texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_mixedAffineTailHomLine\_diagX1sqTail},
\]
together with their affine-equivalence wrappers. These no longer require the
classification step to hand back an explicit relation to the canonical
homogeneous line itself. Instead, starting from exact relations to
\[
  1,\ x_1,\ q_2,\ q_3,
\]
with \(q_2,q_3\) having zero constant and \(x_1\)-linear coefficients, Lean
now forms the canonical extracted line
\[
  H(q_2,q_3)
  =
  \operatorname{coeff}_{x_2}(q_3)\,q_2
  -
  \operatorname{coeff}_{x_2}(q_2)\,q_3
\]
internally and dispatches to the already-formalized cross, \(x_2^2\)-type, or
diagonal determinant-zero tail theorem according to the coefficient pattern of
\(H(q_2,q_3)\). So the remaining mixed-affine extraction step no longer has to
produce a separate homogeneous-line relation in those determinant-zero
subbranches; it only has to supply the tailed pair and verify the extracted
line's coefficient case.
The diagonal slice has now been weakened one step further as well. The same
file also proves
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_diagX1sqTail\_det} and its
equivariant version, so the classification step no longer has to pre-kill the
\(x_2^2\)-coefficient of the tailed quadratic in the diagonal case. It is now
enough to know that the homogeneous line has nonzero \(x_2^2\)-coefficient, the
tail has zero \(x_1x_2\)-coefficient and nonzero \(x_2\)-coefficient, and the
determinant
\[
  \operatorname{coeff}_{x_2^2}(q_2)\operatorname{coeff}_{x_1^2}(q_3)
  -
  \operatorname{coeff}_{x_1^2}(q_2)\operatorname{coeff}_{x_2^2}(q_3)
\]
is nonzero. Lean then subtracts the required multiple of the diagonal line
internally and dispatches to the previous diagonal tail theorem.
The extracted-line layer has now been upgraded in the same direction. The same
file now also proves determinant-zero refinements
\[
  \texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_mixedAffineTailHomLine\_crossTail\_det},
\]
\[
  \texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_mixedAffineTailHomLine\_x1sqTail\_det},
\]
and
\[
  \texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_mixedAffineTailHomLine\_diagX1sqTail\_det},
\]
again with equivariant wrappers. In the cross and \(x_2^2\)-type cases, the
vanishing of the tailed-pair determinant for the extracted line now
automatically forces the remaining offending tail coefficient to vanish, so the
classification step no longer has to provide that coefficient cancellation as a
separate hypothesis. In the diagonal case, the same extracted-line determinant
zero forces the \(x_1x_2\)-coefficient of the tail to vanish, while the second
diagonal determinant is passed directly to the already-formalized
\texttt{diagX1sqTail\_det} theorem.
The theorem surface is now one step closer to the natural pair data as well.
\texttt{QuadraticCoordinateForm.lean} now proves the determinant identities
\[
  \det\!\begin{pmatrix}
    \operatorname{coeff}_{x_1x_2}(H(q_2,q_3)) &
    \operatorname{coeff}_{x_2^2}(H(q_2,q_3)) \\
    \operatorname{coeff}_{x_1x_2}(q_3) &
    \operatorname{coeff}_{x_2^2}(q_3)
  \end{pmatrix}
  =
  \operatorname{coeff}_{x_2}(q_3)
  \det\!\begin{pmatrix}
    \operatorname{coeff}_{x_1x_2}(q_2) &
    \operatorname{coeff}_{x_2^2}(q_2) \\
    \operatorname{coeff}_{x_1x_2}(q_3) &
    \operatorname{coeff}_{x_2^2}(q_3)
  \end{pmatrix},
\]
and likewise for the \((x_2^2,x_1^2)\) determinant. Using those identities,
\texttt{RepresentativeMixedAffinePlane.lean} now exports the stronger wrappers
\[
  \texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_mixedAffineTailHomLine\_crossTail\_origDetZero},
\]
\[
  \texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_mixedAffineTailHomLine\_x1sqTail\_origDetZero},
\]
and
\[
  \texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_mixedAffineTailHomLine\_diagX1sqTail\_origDetZero},
\]
again with equivariant wrappers. So the remaining classification no longer has
to rewrite determinant-zero conditions from the original tailed pair
\((q_2,q_3)\) into a second determinant involving the extracted line
\(H(q_2,q_3)\); Lean now performs that conversion internally.
The diagonal determinant-zero slice has now been compressed one step further.
The new theorem
\[
  \texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_mixedAffineTailHomLine\_diagX1sqTail\_origDetZero\_cases}
\]
and its equivariant wrapper no longer require the second diagonal determinant as
input. If that determinant is nonzero, Lean calls the existing diagonal tail
theorem. If it vanishes, Lean subtracts the corresponding multiple of the
extracted homogeneous line and shows that the remaining coordinate is affine,
so the branch falls back to \texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_affineTail}
and hence to the already-solved \(\dim(W \cap \mathrm{Aff}_1)=3\) case. So the
classification step now only has to identify the diagonal extracted-line type
and the original determinant-zero relation; the secondary diagonal determinant
is handled internally by a Lean case split.
The same affine-tail fallback is now formalized for the other two
determinant-zero extracted-line slices as well. The new theorems
\[
  \texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_mixedAffineTailHomLine\_crossTail\_origDetZero\_cases}
\]
and
\[
  \texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_mixedAffineTailHomLine\_x1sqTail\_origDetZero\_cases},
\]
again with equivariant wrappers, no longer require a separate nonvanishing
\(x_1^2\)-tail coefficient. If that coefficient is present, Lean invokes the
existing nonsurjective cross or \(x_2^2\)-type theorem. If it vanishes, Lean
subtracts the corresponding multiple of the extracted homogeneous line and
falls back to the affine-tail theorem. So all three determinant-zero extracted
line types are now handled from the same natural data: the extracted-line
coefficient pattern, the original determinant-zero relation, and the nonzero
linear tail.
That layer is now also available at the same top transport interface used by
the determinant-nondegenerate tailed-pair theorem. For each of the three
determinant-zero extracted-line types, \texttt{RepresentativeMixedAffinePlane.lean}
now exports coordinate-level equivariant theorems and
\(\mathrm{eq\_mix}\)-plus-affine-equivalence pullback theorems. So the
remaining classification step no longer has to descend to relation vectors on
the transformed factor: it is enough to produce exact normalized coordinates
\[
  \operatorname{mix}(M^\top, e^{-1}\!\cdot u) = (1,x_0,q_2,q_3)
\]
together with the relevant coefficient pattern for
\(H(q_2,q_3)=\texttt{mixedAffineTailHomLine}(q_2,q_3)\), and Lean now finishes
the determinant-zero branch directly from that data.
The cross-type determinant-zero slice has now been weakened once more. The same
file now proves
\[
  \texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_mixedAffineTailHomLine\_crossShear\_origDetZero\_cases},
\]
together with its equivariant and
\(\operatorname{mix}(M^\top,e^{-1}\!\cdot u)\)-transport wrappers. So the
classification step no longer has to return a homogeneous line that is already
exactly proportional to \(x_0x_1\). It is now enough that the extracted line
\(H(q_2,q_3)\) satisfy
\[
  \operatorname{coeff}_{x_1^2} H(q_2,q_3)=0,\qquad
  \operatorname{coeff}_{x_0x_1} H(q_2,q_3)\neq 0,
\]
along with the usual determinant-zero relation for \((q_2,q_3)\) and a nonzero
\(x_1\)-tail coefficient on \(q_3\). Lean then applies the explicit internal
\(x_1\)-shear killing the \(x_0^2\)-component of \(H(q_2,q_3)\), subtracts the
induced \(x_0\)-tail from the transformed pair, and dispatches automatically to
the already-formalized pure-cross determinant-zero theorem. So even the
cross-type branch no longer needs a separate external ``normalize to the
pure-cross representative'' lemma; the remaining work is entirely upstream
extraction and transport.
The complementary determinant-zero side now admits the same kind of internal
normalization. The file also proves
\[
  \texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_mixedAffineTailHomLine\_diagShear\_origDetZero\_cases},
\]
again with equivariant and
\(\operatorname{mix}(M^\top,e^{-1}\!\cdot u)\)-transport wrappers. So the
classification step no longer has to externally shear the extracted line
\(H(q_2,q_3)\) into a pure \(x_1^2\)-type or diagonal representative. It is
now enough to produce normalized coordinates \((1,x_0,q_2,q_3)\) with zero
constant and \(x_0\)-linear parts in \(q_2,q_3\), the original
determinant-zero relation for the pair, nonzero \(x_1\)-tail in \(q_3\), and
\[
  \operatorname{coeff}_{x_1^2} H(q_2,q_3)\neq 0.
\]
Lean then applies the internal \(x_1\)-shear killing the cross coefficient of
\(H(q_2,q_3)\), subtracts the induced \(x_0\)-tails, re-extracts the
homogeneous line, and dispatches automatically to the already-formalized
diagonal/\(x_1^2\)-type determinant-zero theorem. So the determinant-zero
mixed-affine branch is now much closer to a coarse extracted-line split, but
one residual pure-\(x_0^2\) slice still remains outside the currently
formalized wrappers. The new numerical probe
\texttt{julia/mixed\_affine\_det\_zero\_x0sq\_tail\_2026\_04\_10.jl} shows:
\[
  \texttt{mixed\_affine\_det\_zero\_x0sq\_tail\_probe}
  =
  (\texttt{samples}=20,\ \texttt{generic\_ranks}=[15],\
   \texttt{low\_affine\_ranks}=[13],\
   \texttt{low\_affine\_missing}=[(1,3),(0,4)]).
\]
So when the extracted line is pure \(x_0^2\), the generic determinant-zero
family looks surjective, while the special low-affine sub-slice with vanishing
\(x_0x_1\)- and \(x_1^2\)-terms still behaves like the known rank-\(13\)
family. The remaining classification therefore appears to split into three
determinant-zero outputs:
\(\operatorname{coeff}_{x_1^2} H\neq 0\),
\(\operatorname{coeff}_{x_1^2} H=0\) with
\(\operatorname{coeff}_{x_0x_1} H\neq 0\), and the residual pure-\(x_0^2\)
line.

The next Lean step removed the generic pure-\(x_0^2\) branch from that list.
In \texttt{RepresentativeMixedAffinePlane.lean}, the new theorems
\texttt{residual\_eq\_zero\_of\_const\_x0\_x0sqTail\_m02\_nonzero},
\texttt{residual\_eq\_zero\_of\_equiv\_const\_x0\_x0sqTail\_m02\_nonzero}, and
their \texttt{eq\_mix\_affineEquiv} wrapper cover exact normalized coordinates
\((1,x_0,q_2,q_3)\) where \(q_2\) is a pure \(x_0^2\) line and
\(\operatorname{coeff}_{x_1^2}(q_3)\neq 0\). Lean now performs the common
\(x_1\)-translation killing the \(x_1\)-tail of \(q_3\), subtracts the
induced constant and \(x_0\) parts by explicit coordinate relations, and then
falls through to the already-formalized homogeneous-plane theorem. So the
pure-\(x_0^2\) determinant-zero branch is no longer open in its generic
\(\operatorname{coeff}_{x_1^2}(q_3)\neq 0\) case; the genuinely remaining
pure-\(x_0^2\) work is the complementary slice
\(\operatorname{coeff}_{x_1^2}(q_3)=0\), which matches the low-affine-looking
rank-\(13\) behavior seen numerically.

The next round removed another generic part of that complementary slice. The
same file now also proves
\texttt{residual\_eq\_zero\_of\_const\_x0\_x0sqTail\_m11\_nonzero},
\texttt{residual\_eq\_zero\_of\_equiv\_const\_x0\_x0sqTail\_m11\_nonzero}, and
their \texttt{eq\_mix\_affineEquiv} wrapper. These cover exact normalized
coordinates \((1,x_0,q_2,q_3)\) where \(q_2\) is still a pure \(x_0^2\) line,
\(\operatorname{coeff}_{x_1^2}(q_3)=0\), but
\(\operatorname{coeff}_{x_0x_1}(q_3)\neq 0\). Lean now performs the common
\(x_0\)-translation killing the \(x_1\)-tail of \(q_3\), subtracts the induced
constant and \(x_0\) parts from both \(q_2\) and \(q_3\), and then dispatches
to the already-formalized homogeneous-plane theorem. So the genuinely
remaining pure-\(x_0^2\) determinant-zero case is the stricter low-affine
special slice
\[
  \operatorname{coeff}_{x_1^2}(q_3)=0,\qquad
  \operatorname{coeff}_{x_0x_1}(q_3)=0,
\]
which is exactly the numerically rank-\(13\) family.

That last special slice is now formalized too at the exact-coordinate level.
The file adds the direct theorem
\texttt{residual\_eq\_zero\_of\_const\_x0\_x0sqTail\_m11\_m02\_zero}, together
with the coarser wrapper
\texttt{residual\_eq\_zero\_of\_const\_x0\_x0sqTail\_m02\_zero} and their
equivariant/\(\operatorname{mix}(M^\top,e^{-1}\!\cdot u)\)-transport versions.
So once exact normalized coordinates
\((1,x_0,q_2,q_3)\) are available with \(q_2\) a pure \(x_0^2\)-type line and
\(\operatorname{coeff}_{x_1^2}(q_3)=0\), Lean now performs the remaining
internal split automatically:
\(\operatorname{coeff}_{x_0x_1}(q_3)\neq 0\) falls to the translated
homogeneous-plane theorem, while
\(\operatorname{coeff}_{x_0x_1}(q_3)=0\) reduces either to the affine-tail
\(\operatorname{contains\_aff1}\) theorem or to a direct constant relation.
So the pure-\(x_0^2\) determinant-zero branch is no longer waiting on another
representative calculation; the remaining gap there is only the upstream
normalization/extraction statement, together with the still stricter
\(\operatorname{coeff}_{x_1^2}(q_3)\neq 0\) theorem surface.

The next Lean step removes even that remaining asymmetry in the theorem
surface. The file now also proves
\texttt{residual\_eq\_zero\_of\_const\_x0\_mixedAffineTailHomLine\_x0sqTail\_cases}
and its equivariant/\(\operatorname{mix}(M^\top,e^{-1}\!\cdot u)\)-transport
wrappers. So the classification step no longer has to hand back a separately
chosen coordinate \(q_2\) which is already the pure \(x_0^2\) line. It is now
enough to produce normalized coordinates \((1,x_0,q_2,q_3)\) for which the
extracted homogeneous line
\[
  H(q_2,q_3)=\texttt{mixedAffineTailHomLine}(q_2,q_3)
\]
is pure \(x_0^2\) with nonzero coefficient. Lean then performs an internal
orthogonal \(2\times 2\) rotation in the last two factor coordinates, sending
\((q_2,q_3)\) to a pair \((q_H,q_P)\) with \(q_H\) exactly the normalized pure
\(x_0^2\) line, and dispatches automatically to the already-formalized
\texttt{x0sqTail\_m02\_nonzero} or \texttt{x0sqTail\_m02\_zero} branch according
to \(\operatorname{coeff}_{x_1^2}(q_P)\). So the pure-\(x_0^2\) slice is now on
the same extracted-line transport surface as the cross, \(x_1^2\)-type, and
diagonal determinant-zero branches.

The next Lean round compresses those determinant-zero extracted-line outputs
one step further. The file now proves
\texttt{residual\_eq\_zero\_of\_const\_x0\_mixedAffineTailHomLine\_origDetZero\_cases}
and its equivariant/\(\operatorname{mix}(M^\top,e^{-1}\!\cdot u)\)-transport
wrappers. So the classification step no longer has to decide in advance which
determinant-zero extracted-line subtype occurs. It is now enough to produce
exact normalized coordinates \((1,x_0,q_2,q_3)\) with
\(\operatorname{coeff}_{1}(q_i)=\operatorname{coeff}_{x_0}(q_i)=0\), the
original determinant-zero relation
\[
  \operatorname{coeff}_{x_0x_1}(q_2)\operatorname{coeff}_{x_1^2}(q_3)-
  \operatorname{coeff}_{x_1^2}(q_2)\operatorname{coeff}_{x_0x_1}(q_3)=0,
\]
and a nonzero \(x_1\)-tail in \(q_3\). Lean then forms the extracted line
\(H(q_2,q_3)=\texttt{mixedAffineTailHomLine}(q_2,q_3)\) internally and performs
the entire coarse split by itself: if \(H=0\), it closes by a direct constant
relation; if \(\operatorname{coeff}_{x_1^2} H=0\) with
\(\operatorname{coeff}_{x_0x_1} H\neq 0\), it dispatches to the cross-shear
theorem; if \(\operatorname{coeff}_{x_1^2} H\neq 0\), it dispatches to the
diagonal/\(x_1^2\)-type theorem; and if \(H\) is pure \(x_0^2\), it dispatches
to the pure-\(x_0^2\) theorem. So the remaining classification burden is now
only to extract the normalized tailed pair itself, not to refine the
determinant-zero branch into separate extracted-line outputs.

The next Lean round compresses the constant-containing mixed-affine branch
itself one level further. The file now proves
\texttt{residual\_eq\_zero\_of\_const\_x0\_pair\_coeff\_m00\_m10\_zero} and
its equivariant/\(\operatorname{mix}(M^\top,e^{-1}\!\cdot u)\)-transport
wrappers. So once exact normalized coordinates
\[
  (1,x_0,q_2,q_3)
\]
are available with \(q_2,q_3\) quadratic and
\(\operatorname{coeff}_{1}(q_i)=\operatorname{coeff}_{x_0}(q_i)=0\), the
classification step no longer has to provide any further determinant,
tail-placement, or extracted-line subtype data. Lean now performs the whole
internal split itself: if the original determinant is nonzero, it dispatches
to the tailed-pair theorem; if the determinant is zero and one quadratic has a
nonzero \(x_1\)-tail, it dispatches to the coarse determinant-zero wrapper
after swapping the last two factor coordinates if needed; if both \(x_1\)-tail
coefficients vanish but the homogeneous plane is nontrivial, it dispatches to
the homogeneous-plane theorem; and if that plane collapses to a line, it
closes by a direct constant relation. So the mixed-affine constant-containing
\(\dim=2\) branch now asks only for the exact normalized coordinate picture
itself.

The next Lean round strengthens the determinant-nondegenerate sub-branch again.
The file now proves
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_tailedPair\_det\_general}
and its equivariant wrapper. This removes the last orthogonality hypothesis
from that branch: once affine normalization yields exact relation data
\[
  \sum_i c_{0,i} u_i = 1,\qquad
  \sum_i c_{1,i} u_i = x_0,\qquad
  \sum_i c_{2,i} u_i = q_2,\qquad
  \sum_i c_{3,i} u_i = q_3,
\]
with \(q_2,q_3\) quadratic and
\(
\operatorname{coeff}_{x_0x_1}(q_2)\operatorname{coeff}_{x_1^2}(q_3) -
\operatorname{coeff}_{x_1^2}(q_2)\operatorname{coeff}_{x_0x_1}(q_3)\neq 0,
\)
Lean now performs the common translation, affine-tail correction, and the
corrected-pair Gram nondegeneracy argument internally. So the determinant-
nondegenerate mixed-affine branch no longer needs an orthonormal coefficient
basis or an exact \((1,x_0,q_2,q_3)\) coordinate realization.

The next Lean round packages the remaining purely homogeneous
\((1,x_0,q_2,q_3)\) sub-branch at relation level. The file now proves
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_homQuadratics\_cases},
together with the supporting lemmas that (i) a nontrivial homogeneous plane
forces Gram nondegeneracy of the relation pair, and (ii) the complementary
dependent homogeneous-line case closes by an explicit constant relation.
As a result, the exact-coordinate theorem
\texttt{residual\_eq\_zero\_of\_const\_x0\_pair\_coeff\_m00\_m10\_zero} no
longer carries a long inlined proof for the ``both \(x_1\)-tails vanish''
branch; it now delegates that whole homogeneous subcase to the new
relation-level theorem. This does not yet finish the global classification,
but it reduces the remaining gap to the genuinely tailed normalization and
transport data.

The next Lean round lifts the nondependent pure-\(x_0^2\) extracted-line
subcases to relation level too. The file now proves
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_x0sqTail\_m02\_nonzero}
and
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_x0sqTail\_m11\_nonzero}.
These replay the old exact-coordinate proofs directly from relation data:
after the relevant \(x_1\)- or \(x_0\)-translation and affine-tail
correction, Lean lands in a nontrivial homogeneous quadratic plane and closes
through
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_homQuadratics\_plane\_nontrivial\_noGram}.
So the pure-\(x_0^2\) determinant-zero branch is no longer missing its
relation-level lift in the two genuinely plane-like cases. The only remaining
missing lift there is the residual affine/dependent subcase
\(\operatorname{coeff}_{x_1^2}(q_3)=\operatorname{coeff}_{x_0x_1}(q_3)=0\),
which is exactly the branch where the old exact proof closed by constructing a
constant relation.

The next Lean round closes that remaining relation-level gap under the
ambient tail hypothesis that actually arises in the determinant-zero extracted
line branch. The file now proves
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_x0sqTail\_m11\_m02\_zero\_of\_m01\_nonzero}
and the coarser wrapper
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_x0sqTail\_m02\_zero\_of\_m01\_nonzero}.
These use the fact that in the extracted-line setting the corrected tail still
has nonzero \(x_1\)-coefficient, so the final affine/dependent slice falls
directly to the already-formalized relation-level affine-tail theorem rather
than needing a separate constant-relation witness.

The same round also lifts the whole pure-\(x_0^2\) extracted-line branch to
relation level:
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_mixedAffineTailHomLine\_x0sqTail\_cases}
and its equivariant wrapper. So the classification layer no longer has to hand
back exact coordinates even for that slice; exact relation data for
\((1,x_0,q_2,q_3)\) together with the extracted-line coefficient conditions are
already enough.

The next compression packages the entire determinant-zero tailed mixed-affine
branch at relation level, with one natural independence hypothesis on the two
tailed relation vectors. The file now proves
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_mixedAffineTailHomLine\_origDetZero\_cases\_gram},
which takes exact relation data for \(1,x_0,q_2,q_3\), the determinant-zero
condition on \((q_2,q_3)\), a nonzero \(x_1\)-tail in \(q_3\), and nonzero
Gram determinant of the relation pair \((c_2,c_3)\). Lean then performs the
entire coarse split internally: the cross-like, diagonal/\(x_1^2\)-like, and
pure-\(x_0^2\) extracted-line cases dispatch to their existing relation-level
theorems, while the vanishing extracted-line case closes by a constant relation
whose nontriviality is forced by the Gram hypothesis.

Finally, the file now proves the coarse mixed-affine relation theorem
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_pair\_coeff\_m00\_m10\_zero\_gram}
and its equivariant wrapper. So for the constant-containing mixed-affine
\(\dim=2\) branch, the theorem-level target is no longer an exact normalized
coordinate picture \((1,x_0,q_2,q_3)\). It is enough to extract exact relation
data to \(1,x_0,q_2,q_3\), with \(q_2,q_3\) quadratic and zero constant and
\(x_0\) coefficients, together with nonzero Gram determinant of the two
quadratic relation vectors. From there Lean performs the determinant split,
tail swap, extracted-line normalization, homogeneous-plane fallback, and
dependent constant-relation closure internally.

The next Lean round packages the affine normalization itself. The file
\texttt{MixedAffineNormalization.lean} now contains the explicit affine
equivalences \texttt{affineLineEquiv} and \texttt{linearPairEquiv}, sending an
arbitrary nonconstant affine line to \(x_0\) and an arbitrary independent pair
of linear forms to \((x_0,x_1)\). On top of that,
\texttt{RepresentativeSpanThree.lean} now proves the equivariant relation-level
contains-\(\mathrm{Aff}_1\) theorem
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_contains\_aff1} and the
direct wrapper
\texttt{residual\_eq\_zero\_of\_relations\_const\_affinePair\_det}. Likewise,
\texttt{RepresentativeMixedAffinePlane.lean} now proves
\texttt{residual\_eq\_zero\_of\_relations\_const\_affineLine\_pair\_coeff\_m00\_m10\_zero\_gram}.
So the remaining classification step no longer has to pre-normalize these
branches to exact relations with \(x_0\) or \((x_0,x_1)\); it only has to
extract a constant together with an affine basis and the quadratic kernel
data.

The next Lean round collapses the constant-containing \(\dim=2\) branch even
further. The same file now proves
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0} and
\texttt{residual\_eq\_zero\_of\_relations\_const\_affineLine}. The proof no
longer asks the classification step to hand back any quadratic pair at all.
Once relations to \(1\) and to one nonconstant affine line are available,
Lean forms the normalized coefficient map
\[
  c \longmapsto
  \bigl(\operatorname{coeff}_1,\operatorname{coeff}_{x_0}\bigr)
  \bigl(\sum_i c_i u_i\bigr),
\]
extracts a basis of its kernel internally, and feeds the resulting quadratic
pair into the already-formalized coarse mixed-affine relation theorem. So the
mixed-affine constant-containing \(\dim=2\) branch is no longer an external
quadratic-normalization problem; at theorem level it has reduced to the single
task of extracting \(1\) together with one nonconstant affine line.

The same normalization pattern is now available on the low-affine side as
well. In \texttt{RepresentativeLowAffine.lean} the new theorems
\texttt{residual\_eq\_zero\_of\_relations\_linearForm\_homQuadBasis\_det},
\texttt{residual\_eq\_zero\_of\_relations\_linearPair\_x0sq\_x0x1Plane},
\texttt{residual\_eq\_zero\_of\_relations\_linearPair\_x0x1\_x1sqPlane},
\texttt{residual\_eq\_zero\_of\_relations\_linearPair\_x0sq\_x1sqPlane},
\texttt{residual\_eq\_zero\_of\_relations\_linearPair\_x0x1\_diffsqPlane}, and
\texttt{residual\_eq\_zero\_of\_relations\_linearPair\_x0x1\_sumsqPlane}
transport the linear and plane low-affine branches across arbitrary nonzero
linear forms and arbitrary independent linear pairs. So the remaining
classification does not need to normalize low-affine branches to exact
\((x_0)\) or \((x_0,x_1)\) coordinates either; it only has to extract the
corresponding linear basis and transformed homogeneous quadratic data.
Moreover the linear \(\dim=1\) branch now also has the top transport wrapper
\texttt{residual\_eq\_zero\_of\_socp\_of\_eq\_mix\_affineEquiv\_linearForm\_homQuadBasis\_det},
so it lands on the same \(\operatorname{mix}\)-plus-affine-equivariant surface
as the already-normalized \(x_0\)-based branch.

This low-affine plane layer is now also symmetric in the common-factor slice:
the span \(\operatorname{span}(x_0x_1,x_1^2)\) is formalized directly, not
only via an external variable swap back to
\(\operatorname{span}(x_0^2,x_0x_1)\). So the remaining classifier no longer
needs to decide in advance which coordinate ordering presents the repeated-line
plane.

The next round removes the same external homogeneous-basis burden from the
rank-one affine-image branches. In
\texttt{RepresentativeAffineRankOne.lean} the theorems
\texttt{residual\_eq\_zero\_of\_relations\_const\_affineRankOne} and
\texttt{residual\_eq\_zero\_of\_relations\_x0\_affineRankOne} show that it is
enough to know two things:
\begin{itemize}
\item one exact relation to \(1\) or to \(x_0\), and
\item that the affine coefficient map
\[
  c \longmapsto
  \bigl(\operatorname{coeff}_1,\operatorname{coeff}_{x_0},
    \operatorname{coeff}_{x_1}\bigr)\Bigl(\sum_i c_i u_i\Bigr)
\]
has one-dimensional image generated by the matching constant or \(x_0\)
coefficient vector.
\end{itemize}
Lean then handles the rest internally. If the relation map is not injective it
falls back immediately to the already-formalized constant-relation certificate.
Otherwise it takes a basis of the kernel of the affine coefficient map inside
\(\mathbf{R}^4\), proves the resulting three relations are homogeneous and
linearly independent, reconstructs the exact \((x_0^2,x_0x_1,x_1^2)\)
relations by an internal \(3 \times 3\) coefficient-matrix argument, and
applies the existing constant or linear representative theorem.

That rank-one layer is now wider on the nonzero-constant side too. In
\texttt{RepresentativeSurjective.lean} the new theorems
\texttt{quartic\_in\_image\_of\_relations\_constX0\_x0sq\_x0x1\_x1sq},
\texttt{residual\_eq\_zero\_of\_relations\_constX0\_x0sq\_x0x1\_x1sq}, and
\texttt{residual\_eq\_zero\_of\_relations\_constX0\_homQuadBasis\_det} show
that once one has an exact relation to \(\alpha + x_0\) with \(\alpha \neq 0\)
and an invertible homogeneous quadratic basis, the branch is actually
surjective: Lean reconstructs \(x_0\), \(1\), and \(x_1\) directly inside the
admissible image and then reaches every quartic monomial without any kernel
certificate. On top of that, \texttt{RepresentativeAffineRankOne.lean} now
packages the same endpoint as
\texttt{residual\_eq\_zero\_of\_relations\_constX0\_affineRankOne}, so the
rank-one affine-image branch no longer needs to collapse all nonzero-constant
affine lines to the pure constant theorem.

The supporting numerical probe
\texttt{julia/affine\_rank\_one\_nonzero\_const\_2026\_04\_11.jl} matches that
shape. For the normalized model
\[
  (\alpha + x_0,\ q_1,\ q_2,\ q_3),
  \qquad \operatorname{span}(q_1,q_2,q_3)=
    \operatorname{span}(x_0^2,x_0x_1,x_1^2),
\]
all tested nonzero constants \(\alpha \in
\{-3,-2,-1,-\tfrac12,\tfrac12,1,2,3\}\) gave image rank \(15\), and \(20\)
random homogeneous bases at \(\alpha=1\) also gave rank \(15\). Only the
\(\alpha=0\) limit returns to the old linear rank-\(13\) pattern, missing the
constant and \(x_1\) classes.

However, the remaining \(\dim=1\) exact-affine branch is now visibly broader
than the pure affine-rank-one surface. The new file
\texttt{julia/affine\_dim\_one\_tailed\_2026\_04\_11.jl} now records both image
ranks and the exact-affine dimension for several normalized ``one exact affine
line plus one tailed affine direction'' families. The first three probes were
\[
  (x_0,\ x_1+\lambda x_0x_1,\ x_0^2,\ x_1^2),\qquad
  (x_0,\ x_1+\lambda x_0x_1,\ x_0^2,\ x_0x_1),\qquad
  (x_0,\ x_1+\lambda x_0^2,\ x_0^2,\ x_1^2).
\]
Numerically, only the first of those is genuinely exact-affine \(\dim=1\): the
script records
\[
  \texttt{cross\_tail\_exact\_affine\_dims}=[1],\qquad
  \texttt{common\_factor\_tail\_exact\_affine\_dims}=[2],\qquad
  \texttt{x0sq\_tail\_exact\_affine\_dims}=[2].
\]
So the second and third families already belong to the exact-affine
\(\dim=2\) branch, which is closed in Lean, and only the cross family remains
as a genuine new \(\dim=1\) endpoint among those three.

That cross slice is now formalized. In
\texttt{RepresentativeAffineRankOne.lean} the new theorem
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1PlusAX0x1\_x0sq\_x1sq}
proves zero residual from exact relations to
\[
  (x_0,\ x_1 + a x_0x_1,\ x_0^2,\ x_1^2),\qquad a \ne 0.
\]
The image part uses the explicit identity
\[
  x_1 = (x_1 + a x_0x_1)(1-a x_0) + a^2 x_0^2 x_1,
\]
and the kernel part matches the missing constant coefficient with the same
\(x_0/x_0^2\)-based admissible kernel direction used in the previous codimension
one branches.

The updated script also shows that two other families are genuinely
exact-affine \(\dim=1\):
\[
  (x_0,\ x_1+\lambda x_1^2,\ x_0^2,\ x_0x_1),\qquad
  (x_0,\ x_1+\lambda x_0^2,\ x_0x_1,\ x_1^2).
\]
For every tested nonzero \(\lambda \in
\{-3,-2,-1,-\tfrac12,\tfrac12,1,2,3\}\), the first has image rank \(13\)
with missing pure \(x_1\)-powers
\((0,0),(0,1),(0,2),(0,3),(0,4)\), while the second has image rank \(14\) with
only the constant monomial missing.

The second of those is now formalized too. The theorem
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1PlusAX0sq\_x0x1\_x1sq}
in \texttt{RepresentativeAffineRankOne.lean} closes
\[
  (x_0,\ x_1 + a x_0^2,\ x_0x_1,\ x_1^2),\qquad a \ne 0,
\]
again by an explicit image-plus-kernel certificate. The image part first
recovers \(x_1\) from \(x_1 + a x_0^2\) using the exact \(x_0\) relation, then
reaches all nonconstant monomials; the kernel part uses the exact
\(x_0x_1\)-relation to build the missing constant-matching admissible
direction.

The first slice is now formalized as well. The theorem
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1PlusAX1sq\_x0sq\_x0x1}
closes
\[
  (x_0,\ x_1 + a x_1^2,\ x_0^2,\ x_0x_1),\qquad a \ne 0.
\]
Its image proof splits quartics into a pure-\(x_1\) part, handled by the
divisibility identity through \(x_1 + a x_1^2\), and the complementary
\(x_0\)-divisible part, handled directly from the exact relations to
\(x_0\), \(x_0^2\), and \(x_0x_1\). The kernel proof uses a two-parameter
\(x_0/x_0^2\)-based family whose constant coefficient and the extracted
\texttt{x1sqTailFunctional} can be matched independently on each SOS square.

So the remaining classifier still cannot identify the exact-affine \(\dim=1\)
branch with the pure affine-rank-one surface, but there is no remaining
normalized tailed endpoint. What remains is the upstream extraction step that
routes a general exact-affine \(\dim=1\) factor into one of the already solved
surfaces.

The transport surface for those three tailed endpoints is now aligned with the
rest of the proof as well. In
\texttt{RepresentativeAffineRankOne.lean} the new theorems
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_x0\_x1PlusAX0x1\_x0sq\_x1sq},
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_x0\_x1PlusAX0sq\_x0x1\_x1sq},
and
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_x0\_x1PlusAX1sq\_x0sq\_x0x1}
show that after any polynomial algebra equivalence preserving quadratic and
quartic degree, these normalized affine-dim-one coordinate pictures can be
pulled back directly to the original SOCP point. So the remaining classifier no
longer has to package separate transported relation vectors for the tailed
\(\dim=1\) branches; exact normalized coordinates after affine normalization are
enough.

The next extraction lemma is now also internalized. In
\texttt{ExactAffineClassification.lean}, the new theorem
\texttt{exists\_exactAffine\_affineLine\_of\_dimOne\_noConst} shows that if the
exact affine kernel has dimension \(1\), the scalar relation map is injective,
and there is no exact constant relation, then Lean can already extract a
genuine nonconstant exact affine line
\[
  r + a x_0 + b x_1,\qquad a^2+b^2 \ne 0.
\]
So the remaining \(\dim=1\) classifier no longer has to reprove from scratch
that the surviving exact affine relation really has nonzero linear part.

The next retained structural step is now formalized too. The same file proves
\texttt{exists\_relations\_x0\_homQuadBasis\_of\_exactAffineDimOne}: once an
exact-affine \(\dim=1\) branch has been normalized so that one exact relation
is \(x_0\), Lean constructs three further exact relations of the form
\[
  \alpha_{20} + \beta_{20}x_1 + x_0^2,\qquad
  \alpha_{11} + \beta_{11}x_1 + x_0x_1,\qquad
  \alpha_{02} + \beta_{02}x_1 + x_1^2.
\]
So the remaining \(\dim=1\) classifier no longer has to rediscover the
homogeneous quadratic basis: it only has to analyze the six scalar tail
coefficients \((\alpha_{ij},\beta_{ij})\).

Moreover the pure affine-rank-one subcase above that normalization is already
packaged. The theorem
\texttt{residual\_eq\_zero\_of\_relations\_x0\_affineRankOne\_of\_finrank\_range\_one}
shows that if the affine coefficient image is already one-dimensional after
normalizing the exact line to \(x_0\), Lean closes the branch immediately
through the existing affine-rank-one theorem. So the genuinely remaining
\(\dim=1\) work is now concentrated in the affine-image rank-\(2\) tailed
subcase, not the pure rank-\(1\) image.

Two constant-tailed affine-\(\dim=1\) endpoints are now formalized. In
\texttt{RepresentativeAffineRankOne.lean}, the theorem
\texttt{residual\_eq\_zero\_of\_relations\_x0\_onePlusAX0sqBX1sq\_x0x1\_x1sq}
shows that exact relations
\[
  x_0,\qquad 1 + a x_0^2 + b x_1^2,\qquad x_0x_1,\qquad x_1^2,
  \qquad a \ne 0,
\]
already force zero residual. So the affine-\(\dim=1\) classifier no longer has
to treat the mixed \(x_0^2/x_1^2\) constant-tail family as an open endpoint.
The same file now also proves
\texttt{residual\_eq\_zero\_of\_relations\_x0\_onePlusAX0x1\_x0sq\_x1sq}, so
\[
  x_0,\qquad 1 + a x_0x_1,\qquad x_0^2,\qquad x_1^2
\]
is now formalized directly as a surjective exact-affine \(\dim=1\) endpoint as
well: Lean first reconstructs \(x_1\) from the constant-cross relation using
the exact \(x_0\) and \(x_1^2\) relations, then reconstructs \(1\), and from
there reaches every quartic monomial in the admissible image.

The numerical side now matches that split more sharply. The Julia file
\texttt{julia/affine\_dim\_one\_tailed\_2026\_04\_11.jl} records
\texttt{affine\_dim\_one\_constant\_probe}, whose retained summary is:
\[
  \texttt{const\_x0sq\_diag\_ranks}=[15],\qquad
  \texttt{const\_cross\_ranks}=[15],\qquad
  \texttt{const\_x1sq\_common\_ranks}=[13].
\]
So among the visible pure constant tails, the families
\[
  (x_0,\ 1 + a x_0^2 + b x_1^2,\ x_0x_1,\ x_1^2)
  \quad\text{and}\quad
  (x_0,\ 1 + a x_0x_1,\ x_0^2,\ x_1^2)
\]
are now both in the proved column and also look surjective in every tested
nonzero scalar sample, while the repeated-line
family
\[
  (x_0,\ 1 + a x_1^2,\ x_0^2,\ x_0x_1)
\]
stays rank \(13\) with the entire \(x_1\)-axis monomial strip missing.

That remaining repeated-line constant-tail endpoint is now formalized too. The
same file proves
\texttt{residual\_eq\_zero\_of\_relations\_x0\_onePlusAX1sq\_x0sq\_x0x1}, with
the matching equivariant wrapper. The proof mirrors the earlier
\(x_1 + a x_1^2\) repeated-line branch but uses the exact relations
\[
  x_0^2,\qquad x_0x_1
\]
to build a two-parameter admissible-kernel family whose \(\sigma\)-image moves
the pure \(x_1\)-quartic strip into the divisibility conditions for
\((1 + a x_1^2)\). So all three visible pure constant-tail affine-\(\dim=1\)
endpoints are now in the proved column:
\[
  (x_0,\ 1 + a x_0^2 + b x_1^2,\ x_0x_1,\ x_1^2),\qquad
  (x_0,\ 1 + a x_0x_1,\ x_0^2,\ x_1^2),\qquad
  (x_0,\ 1 + a x_1^2,\ x_0^2,\ x_0x_1).
\]

The same repeated-line constant endpoint is now stable under an additional
linear \(x_1\)-tail. In \texttt{RepresentativeAffineRankOne.lean} the new
theorem
\texttt{residual\_eq\_zero\_of\_relations\_x0\_onePlusBX1PlusAX1sq\_x0sq\_x0x1}
proves zero residual from exact relations to
\[
  (x_0,\ 1 + b x_1 + a x_1^2,\ x_0^2,\ x_0x_1),
  \qquad a \ne 0.
\]
The proof performs the internal translation \(x_1 \mapsto x_1 - b/(2a)\):
after translation the mixed constant-linear relation becomes either a pure
\(x_1^2\) relation or a \((1 + a' x_1^2)\) relation, while the translated
\(x_0x_1\) relation is repaired by subtracting the induced \(x_0\) term using
the exact \(x_0\) relation. So this mixed repeated-line slice now reduces
completely to the already solved pure repeated-line endpoints.

The Julia file \texttt{julia/affine\_dim\_one\_tailed\_2026\_04\_11.jl} now
also records
\[
  \texttt{affine\_dim\_one\_constant\_linear\_probe},
\]
and for all tested nonzero \(a,b \in \{-2,-1,1,2\}\) it returns
\[
  \texttt{const\_linear\_x1sq\_common\_ranks}=[13],\qquad
  \texttt{const\_linear\_x1sq\_common\_missing}
    =[ \{(0,0),(0,1),(0,2),(0,3),(0,4)\} ].
\]
So numerically this new mixed repeated-line family behaves exactly like the
pure \(1 + a x_1^2\) slice, which matches the translation-based Lean proof.

That repeated-line constant surface is now also available one step above exact
normal form. In \texttt{RepresentativeAffineRankOne.lean}, the theorem
\texttt{residual\_eq\_zero\_of\_relations\_x0\_onePlusBX1PlusAX1sq\_x0sq\_x0x1Plane}
starts from exact relations to
\[
  x_0,\qquad 1 + b x_1 + a x_1^2,\qquad
  r x_0^2 + s x_0x_1,\qquad t x_0^2 + w x_0x_1
\]
with \(a \ne 0\) and \(rw-st \ne 0\), reconstructs exact \(x_0^2\) and
\(x_0x_1\) internally by Cramer's rule, and then dispatches to the translated
repeated-line endpoint. So the remaining exact-affine \(\dim=1\) classifier no
longer has to hand back a pre-normalized \((x_0^2,x_0x_1)\) pair in that
slice.

The same repeated-line reduction is now available for the mixed
constant-diagonal surface as well. The new theorem
\texttt{residual\_eq\_zero\_of\_relations\_x0\_onePlusAX0sqBX1sq\_x0x1\_x1sqPlane},
again with its equivariant wrapper, starts from exact relations to
\[
  x_0,\qquad 1 + a x_0^2 + b x_1^2,\qquad
  r x_0x_1 + s x_1^2,\qquad t x_0x_1 + w x_1^2
\]
with \(a \ne 0\) and \(rw-st \ne 0\), reconstructs exact \(x_0x_1\) and
\(x_1^2\) internally, and then dispatches to the already formalized endpoint
\((x_0,\ 1 + a x_0^2 + b x_1^2,\ x_0x_1,\ x_1^2)\). So the
exact-affine \(\dim=1\) classifier no longer has to hand the common-factor
constant slice an exact repeated-line basis either.

There is now also a direct exact-normalized mixed-tail endpoint in the same
file:
\texttt{quartic\_in\_image\_of\_relations\_x0\_onePlusAX0sq\_x1PlusBX0x1\_x1sq}
and
\texttt{residual\_eq\_zero\_of\_relations\_x0\_onePlusAX0sq\_x1PlusBX0x1\_x1sq}.
From exact relations to
\[
  x_0,\qquad 1 + a x_0^2,\qquad x_1 + b x_0x_1,\qquad x_1^2
\]
with \(a \ne 0\), Lean now proves image surjectivity directly. The proof first
reconstructs \(1\), \(x_1\), \(x_0^2\), and \(x_0x_1\), then manufactures the
remaining monomials \(x_0^2x_1\), \(x_0^3x_1\), and \(x_0^4\) by multiplying
the \(x_0\)- and \(1+a x_0^2\)-relations by quadratic cofactors. So if the
exact-affine \(\dim=1\) classifier normalizes a branch to this mixed surface,
it no longer needs any further in-plane reconstruction before closing.

The same file now also exports the diagonal-plane analogues
\texttt{residual\_eq\_zero\_of\_relations\_x0\_onePlusAX0x1\_x0sq\_x1sqPlane}
and
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1PlusAX0x1\_x0sq\_x1sqPlane},
again with matching equivariant wrappers. So in the constant-cross and
linear-cross affine-rank-one slices, the classifier no longer needs exact
\((x_0^2,x_1^2)\) relations either; determinant-nondegenerate plane data in
\(\operatorname{span}(x_0^2,x_1^2)\) is now enough.

The same reduction is now available for the remaining linear repeated-line
endpoints too. In \texttt{RepresentativeAffineRankOne.lean}, the new theorems
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1PlusAX0sq\_x0x1\_x1sqPlane}
and
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1PlusAX1sq\_x0sq\_x0x1Plane},
again with matching equivariant wrappers, lift
\[
  (x_0,\ x_1 + a x_0^2,\ x_0x_1,\ x_1^2)
  \quad\text{and}\quad
  (x_0,\ x_1 + a x_1^2,\ x_0^2,\ x_0x_1)
\]
from exact repeated-line bases to arbitrary determinant-nondegenerate bases of
their repeated-line planes. So the exact-affine \(\dim=1\) classifier no
longer has to reconstruct exact \((x_0x_1,x_1^2)\) or exact
\((x_0^2,x_0x_1)\) before entering those linear-tail branches.

The same file now also lifts the pure \(x_1\)-tail branches one step above
their already normalized homogeneous planes. The new theorems
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1Plus\_homQuadratics\_x0sq\_x1sqPlane},
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1Plus\_homQuadratics\_x0x1\_x1sqPlane},
and
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1Plus\_homQuadratics\_x0sq\_x0x1Plane},
again with equivariant wrappers, now accept an exact \(x_0\)-relation, one
exact quadratic relation whose affine part is exactly \(x_1\), and an arbitrary
determinant-nondegenerate basis of the relevant homogeneous plane. Lean
internally reconstructs the canonical homogeneous basis by Cramer's rule,
subtracts the in-plane homogeneous part of the tailed relation, and dispatches
to the already solved exact linear-tail endpoints. So in the tail-rank \(1\)
exact-affine \(\dim=1\) branch, the classifier no longer has to manufacture a
separately normalized tailed coordinate once the homogeneous plane type is
known.

The remaining exact-affine \(\dim=1\) gap is therefore no longer a missing
constant-tail endpoint; it is the classifier that packages a general
rank-\(2\) affine-image tailed branch into one of the already solved
constant-tail or \(x_1\)-tail surfaces.

One more exact-affine \(\dim=1\) classifier output is now internalized in
\texttt{ExactAffineClassification.lean}. The theorem
\texttt{residual\_eq\_zero\_of\_exactAffineDimOne\_tailRangeZero} closes the
tail-rank \(0\) subcase: once an exact \(x_0\) relation is fixed and the tail
map from the \(x_0\)-kernel to the \((1,x_1)\)-coefficients has zero image,
Lean recovers exact relations to
\[
  x_0^2,\qquad x_0x_1,\qquad x_1^2
\]
and dispatches directly to the existing
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x0sq\_x0x1\_x1sq} theorem. So
the remaining \(\dim=1\) classifier no longer needs to treat the pure
homogeneous-tail subcase as open.

The new theorem
\texttt{residual\_eq\_zero\_of\_exactAffineDimOne\_affineLine\_tailRangeZero}
pushes that one step outward. The classifier no longer has to pre-normalize
the unique exact affine line to exact \(x_0\) coordinates before using the
tail-rank \(0\) closure: from any exact affine relation
\[
  r + a x_0 + b x_1,\qquad a^2+b^2\neq 0,
\]
Lean applies \texttt{affineLineEquiv}, proves exact-affine dimension \(1\) is
preserved under that normalization, invokes the existing exact-\(x_0\)
tail-rank \(0\) theorem on the transformed point, and transports the residual
equality back. So this subcase is now closed from the same affine-line-level
input that the outer exact-affine classifier naturally produces.

The same file also now proves
\texttt{exists\_x0\_tail\_const\_x1\_hom\_of\_exactAffineDimOne\_rangeTwo}.
So in the opposite tail-rank \(2\) extreme, Lean can already normalize the
\(x_0\)-kernel to exact relations whose \((1,x_1)\)-tails are
\[
  1,\qquad x_1,\qquad 0,
\]
leaving a single nonzero pure homogeneous quadratic relation as the remaining
classifier datum. The new strengthening
\texttt{exists\_x0\_tail\_const\_x1\_hom\_independent\_of\_exactAffineDimOne\_rangeTwo}
pushes that one step further: Lean now also proves that neither the normalized
constant relation minus \(1\) nor the normalized \(x_1\)-tail relation minus
\(x_1\) can collapse onto that homogeneous line, since either collapse would
manufacture an exact \(1\) or exact \(x_1\) relation and contradict
\(\dim(\mathrm{ExactAff})=1\). So the tail-rank \(2\) branch is no longer just
a raw extractor; it already arrives with the two key independence facts needed
for the remaining classifier. The next retained strengthening
\texttt{exists\_x0\_tail\_const\_x1\_hom\_basis\_of\_exactAffineDimOne\_rangeTwo}
packages the whole homogeneous side: Lean now proves that
\[
  q_0-1,\qquad q_1-x_1,\qquad q_2
\]
are linearly independent, since any nontrivial dependence would manufacture an
exact affine relation of the form \(r+b x_1\), impossible when the exact-affine
space is one-dimensional and already contains \(x_0\). So the tail-rank \(2\)
output is now not only normalized but also a genuine three-dimensional
homogeneous basis on the quadratic side. On the numerical side, the new script
\texttt{julia/affine\_dim\_one\_mixed\_tails\_2026\_04\_11.jl} records
\texttt{affine\_dim\_one\_mixed\_tail\_probe}, and in all \(20\) retained
random tail-rank \(2\) samples it returns
\[
  \texttt{tail\_ranks}=[2],\qquad
  \texttt{exact\_affine\_dims}=[1],\qquad
  \texttt{image\_ranks}=[15],\qquad
  \texttt{missing}=[\varnothing].
\]
The file now also proves
\texttt{exists\_x0\_tail\_nonzero\_homPair\_of\_finrank\_range\_one}. So in the
intermediate tail-rank \(1\) case, Lean extracts exact relations
\[
  q_{\mathrm{tail}},\qquad q_1,\qquad q_2
\]
inside the \(x_0\)-kernel such that \(q_{\mathrm{tail}}\) has nonzero
\((1,x_1)\)-tail, while \(q_1\) and \(q_2\) are pure homogeneous quadratics
and are linearly independent. That means the rank-\(1\) tail case is no longer
an unstructured kernel problem; it has been reduced to classifying one tailed
quadratic relative to a \(2\)-dimensional homogeneous plane.
The new strengthening
\texttt{exists\_x0\_tail\_nonzero\_hom\_basis\_of\_exactAffineDimOne\_rangeOne}
pushes the same branch one step further: after subtracting the explicit affine
tail from \(q_{\mathrm{tail}}\), Lean now proves that
\[
  q_{\mathrm{tail}}-(r+b x_1),\qquad q_1,\qquad q_2
\]
are linearly independent homogeneous quadratics. Any nontrivial dependence
would again manufacture an exact affine relation of the form \(r'+b' x_1\),
impossible in a one-dimensional exact-affine space already containing \(x_0\).
So the tail-rank \(1\) output is no longer only ``one tailed quadratic plus a
homogeneous plane''; it already carries the full homogeneous-basis datum behind
that picture.

The new script
\texttt{julia/affine\_dim\_one\_tail\_rank\_one\_2026\_04\_11.jl} records
\texttt{affine\_dim\_one\_tail\_rank\_one\_probe}. In the retained samples it
returns
\[
  \texttt{mixed\_image\_ranks}=[15],\qquad
  \texttt{const\_image\_ranks}=[15],\qquad
  \texttt{linear\_image\_ranks}=[14],\qquad
  \texttt{linear\_missing}=[\{(0,0)\}].
\]
So numerically the generic rank-\(1\) tail family already looks surjective in
the mixed and constant directions, while the pure \(x_1\)-tail direction still
shows the codimension-one missing constant monomial pattern already visible in
the normalized affine-rank-one endpoint theorems.
That numerical picture does not yet identify the final Lean abstraction by
itself. After inspecting the current theorem inventory, the remaining
\(\dim=1\) gap is not only the diagonal-chart reducers: the exact-affine
extractors now return a full homogeneous basis, so the open step is a
basis-level affine-rank-one normalization theorem that routes that basis into
one of the solved plane/chart surfaces, or an equivalent generic surjective
surface.

The affine-rank-one plane layer has also been widened on the constant side.
In \texttt{RepresentativeAffineRankOne.lean}, the new theorems
\texttt{residual\_eq\_zero\_of\_relations\_x0\_onePlus\_homQuadratics\_x0x1\_x1sqPlane},
\texttt{residual\_eq\_zero\_of\_relations\_x0\_onePlus\_homQuadratics\_x0sq\_x1sqPlane},
and
\texttt{residual\_eq\_zero\_of\_relations\_x0\_onePlus\_homQuadratics\_x0sq\_x0x1Plane}
show that the classifier no longer needs to reconstruct an exact normalized
constant-tailed relation before entering the affine-rank-one endpoint layer.
Once it produces exact \(x_0\), a quadratic with affine part \(1\), and a
determinant-nondegenerate basis of one canonical homogeneous plane, Lean now
recovers the exact plane basis internally and reduces to the already solved
constant affine-rank-one endpoints.

The new classifier-facing bridge in \texttt{ExactAffineClassification.lean}
then packages the first exact-affine \(\dim=1\), tail-rank \(1\) subfamily
directly above those endpoint surfaces:
\texttt{residual\_eq\_zero\_of\_relations\_x0\_homQuadBasis\_singleConstTail}
and
\texttt{residual\_eq\_zero\_of\_relations\_x0\_homQuadBasis\_singleX1Tail}.
These start from the explicit canonical hom-basis relations
\[
  r_{20} + b_{20}x_1 + x_0^2,\qquad
  r_{11} + b_{11}x_1 + x_0x_1,\qquad
  r_{02} + b_{02}x_1 + x_1^2
\]
and close all six support-one tail patterns, i.e. when exactly one of
\((r_{20},r_{11},r_{02})\) is nonzero or exactly one of
\((b_{20},b_{11},b_{02})\) is nonzero. So the remaining \(\dim=1\) classifier
no longer needs to handle those isolated pure-coordinate tail outputs by hand.
The same layer now also proves
\texttt{residual\_eq\_zero\_of\_relations\_x0\_homQuadBasis\_singleMixedTail\_x1sq},
which closes the first genuinely mixed support-one slice: only the
\(x_1^2\)-direction carries tails, and it carries both a nonzero constant term
and a nonzero \(x_1\)-term. Lean rescales that single relation to the normal
form \(1 + b x_1 + a x_1^2\), keeps the exact \(x_0^2\) and \(x_0x_1\)
relations untouched, and dispatches directly to the already-formalized
translated repeated-line affine-rank-one theorem. So the remaining tail-rank
\(1\) classifier is no longer missing the mixed repeated-line support slice
either.
I then packaged the whole support-on-\(x_1^2\) family into the single theorem
\texttt{residual\_eq\_zero\_of\_relations\_x0\_homQuadBasis\_tailOnX1sq}. So
the classifier no longer has to split the \(x_1^2\)-supported tail case into
pure-constant, pure-\(x_1\), and mixed repeated-line subbranches before
entering the affine-rank-one layer: that one support test is now enough.
The remaining support-one mixed slices are now packaged as well. The new exact
affine-rank-one endpoints
\texttt{residual\_eq\_zero\_of\_relations\_x0\_onePlusBX1PlusAX0sq\_x0x1\_x1sq}
and
\texttt{residual\_eq\_zero\_of\_relations\_x0\_onePlusBX1PlusAX0x1\_x0sq\_x1sq},
together with the classifier-facing wrappers
\texttt{residual\_eq\_zero\_of\_relations\_x0\_homQuadBasis\_singleMixedTail\_x0sq},
\texttt{residual\_eq\_zero\_of\_relations\_x0\_homQuadBasis\_singleMixedTail\_x0x1},
\texttt{residual\_eq\_zero\_of\_relations\_x0\_homQuadBasis\_tailOnX0sq},
and
\texttt{residual\_eq\_zero\_of\_relations\_x0\_homQuadBasis\_tailOnX0x1},
show that Lean now closes the whole support-one mixed tail family, not just the
\(x_1^2\) repeated-line slice. So once the extracted tail is supported on
exactly one canonical homogeneous direction, the classifier no longer has to
split any mixed constant-plus-\(x_1\) subcase by hand.
The updated Julia file
\texttt{julia/affine\_dim\_one\_tailed\_2026\_04\_11.jl} now also records
\texttt{affine\_dim\_one\_support\_one\_mixed\_probe}. On the retained scalar
grid \(a,b \in \{-2,-1,1,2\}\), it returns
\[
  \texttt{mixed\_x0sq\_ranks}=[15],\qquad
  \texttt{mixed\_x0sq\_missing}=[\varnothing],\qquad
  \texttt{mixed\_x0x1\_ranks}=[15],\qquad
  \texttt{mixed\_x0x1\_missing}=[\varnothing].
\]
So numerically the two newly formalized support-one mixed slices are uniformly
surjective across that test grid, matching the new Lean theorems.
The affine-rank-one plane layer now absorbs the same mixed normalization step.
In \texttt{RepresentativeAffineRankOne.lean}, the new theorems
\texttt{residual\_eq\_zero\_of\_relations\_x0\_onePlusBX1Plus\_homQuadratics\_x0x1\_x1sqPlane},
\texttt{residual\_eq\_zero\_of\_relations\_x0\_onePlusBX1Plus\_homQuadratics\_x0sq\_x1sqPlane},
and
\texttt{residual\_eq\_zero\_of\_relations\_x0\_onePlusBX1Plus\_homQuadratics\_x0sq\_x0x1Plane}
show that once the classifier has produced exact \(x_0\), one quadratic with
affine part \(1 + b x_1\), and a determinant-nondegenerate basis of one
canonical homogeneous plane, Lean now cancels the in-plane homogeneous terms
internally and dispatches directly to the already solved mixed endpoints. So
the remaining exact-affine \(\dim=1\) gap is no longer mixed-tail cleanup on a
known plane chart; it is the upstream routing of the extracted tail-rank \(1\)
and tail-rank \(2\) outputs into those plane charts or an equivalent generic
surjective theorem.

The mixed constant-plus-\(x_1\) surface now also absorbs the coprime cross
chart above exact plane data. The new theorem
\texttt{residual\_eq\_zero\_of\_relations\_x0\_onePlusBX1Plus\_homQuadratics\_crossDet\_zero}
performs the same internal \(x_1\)-shear as the constant-tail cross theorem,
but then repairs the induced \(x_0\)-linear term in the tailed relation by a
single subtraction of the exact \(x_0\) relation before dispatching to the
existing \texttt{x0sq\_x1sqPlane} mixed theorem.

The pure \(x_1\)-tail surface now has the same coprime-cross reduction above
exact plane data. The new theorem
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1Plus\_homQuadratics\_crossDet\_zero}
again performs the internal \(x_1\)-shear and repairs the induced \(x_0\)-term
by subtracting the exact \(x_0\) relation. Its only extra hypothesis is the
expected one: after that shear, the tailed quadratic must still have nonzero
\(x_0x_1\)-coefficient so that Lean can dispatch to the solved
\texttt{x1Plus\_homQuadratics\_x0sq\_x1sqPlane} surface.

The next repeated-line/common-factor chart has now also narrowed sharply. The
new Julia probe
\texttt{julia/affine\_dim\_one\_common\_factor\_chart\_2026\_04\_12.jl} shows
that in the affine-rank-one common-factor family, the transformed tailed
relation's \(x_0^2\)-coefficient behaves as a real classifier boundary:
\(\texttt{m20}\neq 0\) is the only behavior seen in the retained
exact-affine \(\dim=1\) samples, while the apparent \(\texttt{m20}=0\) slice
appears only when the exact-affine dimension jumps to \(2\). On the formal
side, \texttt{RepresentativeAffineRankOne.lean} now proves
\texttt{residual\_eq\_zero\_of\_relations\_x0\_onePlusBX1Plus\_homQuadratics\_commonFactorChart},
which performs the common-factor \(x_1\)-shear internally, reconstructs the
\((x_0x_1,x_1^2)\)-plane, repairs the induced \(x_0\)-term in the tailed
relation, and dispatches to the existing mixed repeated-line plane theorem.
Because \(b\) is arbitrary, this single theorem also covers the pure
constant-tail common-factor chart as the case \(b=0\).
The same file now also proves the pure \(x_1\)-tail companion
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1Plus\_homQuadratics\_commonFactorChart},
with the same internal common-factor shear and the same \(x_0\)-term repair.
So the repeated-line/common-factor chart is no longer open on the
affine-rank-one side: constant, mixed, and pure \(x_1\) tails all reduce
internally to the existing \((x_0x_1,x_1^2)\)-plane surfaces.

The remaining affine-rank-one diagonal complement is now numerically pinned
down too. The probe
\texttt{julia/affine\_dim\_one\_diagonal\_chart\_2026\_04\_12.jl} samples the
two canonical diagonal-plane outputs
\(\operatorname{span}(x_0x_1,x_0^2-x_1^2)\) and
\(\operatorname{span}(x_0x_1,x_0^2+x_1^2)\), with the tailed relation carrying
a nonzero complementary diagonal component so that the exact-affine dimension
stays \(1\). Its retained output is
\[
  \texttt{affine\_dim\_one\_diagonal\_chart\_probe}
  =
  (\texttt{samples}=30,\ 
   \texttt{diag\_sum\_const.ranks}=[15],\ 
   \texttt{diag\_sum\_x1.ranks}=[14],\ 
   \texttt{diag\_sum\_x1.missing}=[\{(0,0)\}],\ 
   \texttt{diag\_sum\_mixed.ranks}=[15],\ 
   \texttt{diag\_diff\_const.ranks}=[15],\ 
   \texttt{diag\_diff\_x1.ranks}=[14],\ 
   \texttt{diag\_diff\_x1.missing}=[\{(0,0)\}],\ 
   \texttt{diag\_diff\_mixed.ranks}=[15]).
\]
So both diagonal charts match the pattern already visible in the rank-\(1\)
tail probe: the constant and mixed tails look generically surjective, while
the pure \(x_1\)-tail direction still shows the same codimension-one missing
constant monomial as the retained affine-rank-one linear endpoint.

The new script
\texttt{julia/affine\_dim\_one\_diagonal\_complement\_2026\_04\_12.jl}
shows that this is not just a chart artifact. On the exact canonical
complement models
\[
  (x_0,\ 1 + a(x_0^2+x_1^2),\ x_0x_1,\ x_0^2-x_1^2),\qquad
  (x_0,\ x_1 + a(x_0^2+x_1^2),\ x_0x_1,\ x_0^2-x_1^2),
\]
their mixed \(1+b x_1\) companions, and the corresponding
\(\operatorname{span}(x_0x_1,\ x_0^2+x_1^2)\) versions with the complementary
tail \(a(x_0^2-x_1^2)\), the retained output is
\[
  \texttt{affine\_dim\_one\_diagonal\_complement\_probe}
  =
  (\texttt{diag\_sum\_const.ranks}=[15],\ 
   \texttt{diag\_sum\_x1.ranks}=[14],\ 
   \texttt{diag\_sum\_x1.missing}=[\{(0,0)\}],\ 
   \texttt{diag\_sum\_mixed.ranks}=[15],\ 
   \texttt{diag\_diff\_const.ranks}=[15],\ 
   \texttt{diag\_diff\_x1.ranks}=[14],\ 
   \texttt{diag\_diff\_x1.missing}=[\{(0,0)\}],\ 
   \texttt{diag\_diff\_mixed.ranks}=[15]).
\]
So the remaining diagonal theorem is now pinned down to an exact affine-rank-one
plane family: the pure \(x_1\)-tail diagonal-complement models are the only
stable nonsurjective outputs, while the constant and mixed complement models
already look uniformly surjective on their exact canonical representatives.

The next Lean round has already absorbed the local diagonal normal-form
algebra needed for that branch. In
\texttt{RepresentativeAffineRankOne.lean} there are now explicit expansion
lemmas for the constant, pure \(x_1\), and mixed \(1+b x_1\) affine-rank-one
tails relative to the two diagonal bases
\(\operatorname{span}(x_0x_1,x_0^2-x_1^2)\) and
\(\operatorname{span}(x_0x_1,x_0^2+x_1^2)\). So the remaining diagonal-chart
work is no longer blocked on repeated coefficient rewrites from
\(\texttt{quadForm}\); what remains is the actual chart reduction theorems
that feed those normal forms into the already solved endpoint surfaces.

That scaling layer is now formalized too for the pure \(x_1\)-tail branch. The
new theorems
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1Plus\_homQuadratics\_diag\_sum\_zero}
and
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1Plus\_homQuadratics\_diag\_diff\_zero}
start from exact relations to \(x_0\), a tailed quadratic \(q_1\), and a
homogeneous pair \(q_2,q_3\) satisfying the corresponding diagonal relation
\(a\,\operatorname{coeff}_{20} + d\,\operatorname{coeff}_{02}=0\). Lean then
performs the internal \(x_1\)-scaling by \(\sqrt{d/a}\) or \(\sqrt{-d/a}\),
normalizes the \(x_1\)-coefficient of the tailed relation back to \(1\),
reconstructs the \((x_0x_1,\ x_0^2-x_1^2)\) or
\((x_0x_1,\ x_0^2+x_1^2)\) plane by Cramer's rule, and dispatches to the
already solved diagonal plane surfaces. So the affine-rank-one diagonal
complement is no longer missing its scaling algebra either; what remains is the
outer shear/chart layer that produces these diagonal-relation hypotheses from a
general homogeneous pair.

In the next round the same file was lifted to the transport surface via
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_x0\_x1Plus\_homQuadratics\_diag\_sum\_zero}
and
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_x0\_x1Plus\_homQuadratics\_diag\_diff\_zero}.
So later classifier code no longer has to rebuild the \texttt{dotTransport} and
\texttt{mapVec} bookkeeping by hand once it has normalized into these
diagonal-scaling hypotheses.

The pure constant diagonal-complement branch now has the same scaling layer.
The new theorems
\texttt{residual\_eq\_zero\_of\_relations\_x0\_onePlus\_homQuadratics\_diag\_sum\_zero}
and
\texttt{residual\_eq\_zero\_of\_relations\_x0\_onePlus\_homQuadratics\_diag\_diff\_zero},
again with matching equivariant wrappers, start from exact relations to
\(x_0\), a quadratic \(q_1\) with affine part exactly \(1\), and a homogeneous
pair satisfying the same diagonal coefficient relation
\(a\,\operatorname{coeff}_{20}+d\,\operatorname{coeff}_{02}=0\). Lean now
performs the internal \(x_1\)-scaling, converts the homogeneous pair to the
\((x_0x_1,\ x_0^2-x_1^2)\) or \((x_0x_1,\ x_0^2+x_1^2)\) plane by Cramer's
rule, and dispatches directly to the already solved pure-constant diagonal
plane theorems. So the constant affine-rank-one diagonal complement is no
longer blocked on scaling algebra either; what remains there is the outer chart
reduction and the mixed \(1+b x_1\) companion.

So the remaining \(\dim=1\) work is now sharply split: tail-rank \(0\) is
proved, tail-rank \(1\) and tail-rank \(2\) both have structural extractor
theorems, and the genuinely open classifier work is the basis-level routing of
those extracted homogeneous-basis outputs into the affine-rank-one theorem
surface above. The diagonal-complement reducers are one visible part of that
layer, not the whole missing abstraction. One additional piece of that surface
is now closed on the constant side: in
\texttt{RepresentativeAffineRankOne.lean} the new theorems
\texttt{residual\_eq\_zero\_of\_relations\_x0\_onePlus\_homQuadratics\_x0x1\_diffsqPlane}
and
\texttt{residual\_eq\_zero\_of\_relations\_x0\_onePlus\_homQuadratics\_x0x1\_sumsqPlane},
with matching equivariant wrappers, start from exact relations to \(x_0\), a
quadratic \(q_1\) with affine part exactly \(1\), and an invertible basis of
\(\operatorname{span}(x_0x_1,x_0^2-x_1^2)\) or
\(\operatorname{span}(x_0x_1,x_0^2+x_1^2)\). Lean reconstructs the canonical
\((x_0x_1,\ x_0^2 \mp x_1^2)\) basis by Cramer's rule, subtracts the
corresponding homogeneous part from \(q_1\) to recover an exact constant
relation, and then dispatches to the existing mixed-affine plane theorems. So
the exact constant diagonal-complement slice is no longer merely numerical; the
remaining affine-rank-one diagonal work is concentrated in the pure \(x_1\)
and mixed \(1+b x_1\) outputs above the final classifier layer.
those outputs into the already solved endpoint or generic-surjective surfaces,
with the diagonal charts now reduced to a concrete Lean target rather than an
unclassified numerical branch.

\section*{Pending Step}
To finish the theorem, it remains to formalize the remaining normalization and
classification data for the branches not yet reduced to a one-line interface:
most notably the final extraction/classification interface itself. The
low-affine homogeneous side is no longer blocked at the chart level. On top of
the cross-product scalars \texttt{lowHomQuadPlaneA/B/C} and their relation
lemmas, \texttt{RepresentativeLowAffine.lean} first proves the normalization
charts
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_homQuadratics\_crossDet\_zero}
and
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_homQuadratics\_commonFactorChart},
each with the matching transport wrappers. The first reduces any homogeneous
low-affine plane whose
\((m_{11},m_{02})\)-minor vanishes but whose \((m_{20},m_{02})\)-minor is
nonzero to the already-formalized \(\operatorname{span}(x_0^2,x_1^2)\) plane
after an internal \(x_1\)-shear. The second covers the repeated-line chart
where the same shear kills the \(x_0^2\) coefficients and lands in the already
formalized \(\operatorname{span}(x_0x_1,x_1^2)\) plane. The unsuccessful larger
dispatcher attempt was still backed out rather than left as a non-terminating
proof script. The file then adds the scale-only inner theorem surface for the
remaining diagonal branch:
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_homQuadratics\_diag\_sum\_zero}
and
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_homQuadratics\_diag\_diff\_zero},
again with equivariant and \texttt{linearPair} wrappers. These reduce the
post-shear diagonal relation directly to the already formalized
\(\operatorname{span}(x_0x_1,x_0^2-x_1^2)\) and
\(\operatorname{span}(x_0x_1,x_0^2+x_1^2)\) planes after an internal
\(x_1\)-scaling. On top of those, the file now proves the outer diagonal chart
wrappers
\texttt{...diagSumChart} and \texttt{...diagDiffChart}, then packages the whole
homogeneous low-affine branch into the single theorem
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_homQuadratics\_plane\_nontrivial},
again with equivariant and \texttt{linearPair} wrappers. So the low-affine
homogeneous classifier is now internal to Lean: once one has exact relations to
a linear pair and to a nontrivial homogeneous quadratic plane, Lean dispatches
automatically to the common-factor or generic rank-\(14\) branch.

The low-affine plane surface is now slightly cleaner as well. The same file now
proves that any linearly independent pair of homogeneous quadratics already has
nontrivial plane data, via the public criterion
\texttt{lowHomQuadPlane\_nontrivial\_of\_independent\_pair}, and packages that
into the direct wrapper
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_homQuadratics\_independent}.
So later classifier code no longer needs to carry a local proof that one of the
three cross-product minors \texttt{lowHomQuadPlaneA/B/C} is nonzero.

That branch now also reaches the same top transport surface as the other main
branches. In the same file, the new theorems
\texttt{residual\_eq\_zero\_of\_socp\_of\_eq\_mix\_mapVec\_x0\_x1\_homQuadratics\_plane\_nontrivial}
and
\texttt{residual\_eq\_zero\_of\_socp\_of\_eq\_mix\_affineEquiv\_x0\_x1\_homQuadratics\_plane\_nontrivial}
show that after affine normalization and orthogonal factor mixing, an exact
coordinate picture \((x_0,x_1,q_2,q_3)\) with a nontrivial homogeneous plane is
already enough to close the SOCP branch directly.

The remaining task is therefore no longer to add more low-affine chart
theorems. It is to extract, from an arbitrary admissible rank-\(4\) factor, the
appropriate affine or linear basis and the matching mixed-affine, low-affine,
or rank-one relation package, and then assemble the root declaration in
\texttt{TernaryQuarticProof.lean}. The admissible image-plus-cone certificate
and the representative/plane-level no-spurious-SOCP theorems are already in
place.

The next Lean step finally starts to package that extraction itself. In
\texttt{ExactAffineClassification.lean} the exact affine relations inside
\(\operatorname{span}(u_0,u_1,u_2,u_3)\) are encoded as the kernel of the
homogeneous coefficient map
\[
  c \longmapsto
  \bigl(\operatorname{coeff}_{x_0^2},
    \operatorname{coeff}_{x_0x_1},
    \operatorname{coeff}_{x_1^2}\bigr)\Bigl(\sum_i c_i u_i\Bigr).
\]
This already closes two classifier outputs directly. The theorem
\texttt{residual\_eq\_zero\_of\_exactAffineDimThree} shows that if this exact
affine kernel has dimension \(3\), then Lean reconstructs exact relations to
\(1\), \(x_0\), and \(x_1\) internally and dispatches to the existing
\(\dim(W \cap \mathrm{Aff}_1)=3\) certificate. The theorem
\texttt{residual\_eq\_zero\_of\_exactAffineDimTwo\_contains\_one} does the same
for the constant-containing \(\dim = 2\) output: from a single exact constant
relation and the fact that the exact affine kernel has dimension \(2\), Lean
proves there is a second independent exact affine line and closes through the
coarse mixed-affine theorem
\texttt{residual\_eq\_zero\_of\_relations\_const\_affineLine}. So two global
classification branches are now expressed directly in the kernel-of-homogeneous
map language rather than by hand-picked representative coordinates.

The next extraction step is now also internalized there for the no-constant
\(\dim = 2\) branch, even though the full SOCP theorem for that branch is not
yet assembled. The new theorem
\texttt{exists\_exactAffine\_affinePair\_of\_dimTwo\_noConst} proves that if the
exact affine kernel has dimension \(2\) and contains no exact constant
relation, then its linear-coefficient projection is automatically an
isomorphism onto \(\mathbf{R}^2\). Lean therefore reconstructs exact relations
to the normalized affine pair \(x_0 + r_0\) and \(x_1 + r_1\) internally. So
the remaining gap in that branch is no longer affine-pair extraction; it is the
next low-affine or translated-homogeneous quadratic package sitting above that
pair.

That next package theorem is now formalized at the low-affine homogeneous
surface too. In \texttt{RepresentativeLowAffine.lean}, the new theorem
\texttt{residual\_eq\_zero\_of\_relations\_affinePair\_homQuadratics\_independent}
starts from exact relations to \(x_0 + r_0\), \(x_1 + r_1\), and a quadratic
pair \(q_2,q_3\). If translating by \((-r_0,-r_1)\) makes both quadratics
homogeneous and linearly independent, Lean now performs that translation
internally, proves the translated homogeneous plane is nontrivial, and
dispatches to the existing low-affine homogeneous theorem surface. So the
remaining no-constant \(\dim=2\) classifier gap is now narrower again: it only
has to extract the right quadratic pair above the internally reconstructed
affine pair, not reprove the translation step or the low-affine homogeneous
reduction.

The linear-algebra side of that same branch is now tighter too. In
\texttt{ExactAffineClassification.lean}, the theorems
\texttt{translatedLinearCoeffMap\_surjective\_of\_affinePair} and
\texttt{translatedLinearCoeffKernel\_finrank\_two\_of\_affinePair} show that
once Lean has reconstructed an affine pair \(x_0 + r_0\), \(x_1 + r_1\), the
common translation by \((-r_0,-r_1)\) automatically yields a surjective linear
coefficient map onto \((x_0,x_1)\), and hence a two-dimensional kernel of
relations with zero translated \(x_0\)- and \(x_1\)-coefficients. So the
remaining no-constant \(\dim=2\) extraction problem is now more precise: it is
no longer to find the translated zero-linear-tail space, but to understand the
constant-term functional on that already-formalized \(2\)-dimensional kernel
and route the resulting cases to the mixed-affine or low-affine theorem
surfaces.

One side of that split is now formalized too. The theorem
\texttt{residual\_eq\_zero\_of\_relations\_affinePair\_translatedKernel\_constZero}
in \texttt{ExactAffineClassification.lean} assumes an exact affine pair
\(x_0+r_0\), \(x_1+r_1\) and the condition that every relation in the
translated two-dimensional zero-linear-tail kernel already has zero translated
constant coefficient. Lean then chooses a basis of that kernel, proves the two
translated basis images are homogeneous and linearly independent, and closes the
branch through the previously formalized theorem
\texttt{residual\_eq\_zero\_of\_relations\_affinePair\_homQuadratics\_independent}.
So the no-constant \(\dim=2\) classifier gap is now cut exactly in half: only
the complementary nonzero-constant split on the translated kernel remains.

That complementary half is now packaged more explicitly too. The new theorem
\texttt{exists\_translatedKernel\_constSplit\_of\_affinePair} in the same file
starts from an exact affine pair and the hypothesis that the translated
zero-linear-tail kernel contains some relation with nonzero constant term.
Lean then normalizes that functional internally: it produces one translated
kernel relation with constant coefficient \(1\) and a second nonzero translated
kernel relation with constant coefficient \(0\). So the remaining no-constant
\(\dim=2\) branch is no longer waiting on a vague existential split. It is now
reduced to the constant-perturbed low-affine closure problem for an explicit
normalized pair \((1+H_1, H_2)\) with zero linear tails.

The latest normalized probe suggests that this remaining half is narrower than
it first appeared. The script
\texttt{julia/low\_affine\_const\_perturbation\_2026\_04\_10.jl} samples the
model
\[
  (x_0,\ x_1,\ a + H_1,\ b + H_2)
\]
with \((H_1,H_2)\) spanning a random nontrivial homogeneous low-affine plane.
All \(20\) random samples came back with image rank \(15\). Among the control
families, the constant-perturbed coprime, diagonal, and split planes are also
rank \(15\), while the constant-perturbed common-factor slice remains rank
\(14\), missing only \(x_1^4\). So the likely remaining theorem shape is not a
generic low-affine obstruction with constants; it is a surjective generic
deformation plus a special repeated-line/common-factor subcase.

The first formal theorem in that direction is now in place. The file
\texttt{TernaryQuarticProof/RepresentativeLowAffine.lean} proves
\texttt{quartic\_in\_image\_of\_relations\_x0\_x1\_onePlusX0sq\_x1sq} and
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_onePlusX0sq\_x1sq}. So the
constant-perturbed coprime slice
\[
  (x_0,\ x_1,\ 1 + x_0^2,\ x_1^2)
\]
is no longer only a numerical observation: Lean now proves full quartic
surjectivity for that exact-relation surface. That theorem now also sits on the
same affine-pair transport layer as the other low-affine branches: the same file
proves
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_x0\_x1\_onePlusX0sq\_x1sq}
and
\texttt{residual\_eq\_zero\_of\_relations\_affinePair\_onePlusX0sq\_x1sq}.
So once classifier code identifies the coprime constant-perturbed case above an
exact affine pair \(x_0+r_0,\ x_1+r_1\), Lean now performs the common
translation internally instead of requiring hand-normalized coordinates.

The same file now also formalizes the split constant-perturbed slice. The new
theorems
\texttt{quartic\_in\_image\_of\_relations\_x0\_x1\_onePlusX0x1\_diffsq},
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_onePlusX0x1\_diffsq},
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_x0\_x1\_onePlusX0x1\_diffsq},
and
\texttt{residual\_eq\_zero\_of\_relations\_affinePair\_onePlusX0x1\_diffsq}
show that the model
\[
  (x_0,\ x_1,\ 1 + x_0x_1,\ x_0^2 - x_1^2)
\]
is quartic-surjective and already sits on the same affine-pair transport
surface. So in the nonzero-constant exact-affine \(\dim=2\) branch, both the
coprime and split rank-\(15\) deformations are now available directly in Lean;
the remaining constant-perturbed low-affine work is concentrated in the
repeated-line/common-factor slice and the final classifier that routes into
these theorem surfaces.

That repeated-line slice is now formalized too. The same file proves
\texttt{quartic\_in\_image\_of\_relations\_x0\_x1\_onePlusX0sq\_x0x1\_of\_coeff\_m04\_zero},
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_onePlusX0sq\_x0x1},
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_x0\_x1\_onePlusX0sq\_x0x1},
and
\texttt{residual\_eq\_zero\_of\_relations\_affinePair\_onePlusX0sq\_x0x1}.
These show that the model
\[
  (x_0,\ x_1,\ 1 + x_0^2,\ x_0x_1)
\]
has exactly the expected rank-\(14\) behavior at theorem level: every quartic
with zero \(x_1^4\) coefficient lies in the admissible image, and the missing
\(x_1^4\) class is removed by the existing common-factor quartic kernel. So the
nonzero-constant exact-affine \(\dim=2\) branch is no longer missing a low-rank
endpoint; what remains is the classifier that decides which translated-kernel
output lands in the coprime, split, or repeated-line theorem surface.

The same file now also formalizes the symmetric repeated-line target
\[
  (x_0,\ x_1,\ 1 + x_0x_1,\ x_1^2).
\]
The new theorems
\texttt{quartic\_in\_image\_of\_relations\_x0\_x1\_onePlusX0x1\_x1sq\_of\_coeff\_m40\_zero},
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_onePlusX0x1\_x1sq},
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_x0\_x1\_onePlusX0x1\_x1sq},
and
\texttt{residual\_eq\_zero\_of\_relations\_affinePair\_onePlusX0x1\_x1sq}
put the repeated-line/common-factor branch on both canonical constant-perturbed
orientations. That now matches the output shape of the already formalized
homogeneous common-factor chart, so the remaining gap is no longer a missing
endpoint theorem but the final classifier that routes the exact-affine
translated-kernel data into these surfaces.

That classifier now has one more formalized output. In
\texttt{TernaryQuarticProof/RepresentativeLowAffine.lean} the theorem
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_onePlus\_homQuadratics\_dependent}
handles the dependent constant-perturbed low-affine input directly: if
\[
  q_2 = 1 + H,\qquad q_3 = tH,\qquad q_3 \neq 0,
\]
with both relations carrying zero linear part, Lean combines \(q_2\) and \(q_3\)
to recover an exact constant relation and closes via the existing
\(\operatorname{span}(1,x_0,x_1)\) theorem. On top of that,
\texttt{TernaryQuarticProof/ExactAffineClassification.lean} now contains
\texttt{residual\_eq\_zero\_of\_relations\_affinePair\_translatedKernel\_constSplit\_dependent},
which transports the translated-kernel constant-split data from the no-constant
exact-affine \(\dim=2\) branch into that dependent theorem. So the remaining
no-constant exact-affine \(\dim=2\) gap is no longer the whole nonzero-constant
half; it is the complementary independent constant-perturbed branch.

That interface has now been tightened one step further. In
\texttt{RepresentativeLowAffine.lean}, the theorem
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_onePlus\_homQuadratics\_not\_independent}
shows that the dependent branch does not need the explicit vanishing equations
\(\texttt{lowHomQuadPlaneA}=\texttt{lowHomQuadPlaneB}=\texttt{lowHomQuadPlaneC}=0\)
as input: it is enough that the homogeneous pair
\((q_2-1,q_3)\) is not linearly independent and that \(q_3 \neq 0\). Lean now
extracts the scalar-multiple relation internally and falls back to the older
dependent theorem. On top of that,
\texttt{ExactAffineClassification.lean} now contains
\texttt{residual\_eq\_zero\_of\_relations\_affinePair\_translatedKernel\_constSplit\_not\_independent},
so the exact-affine classifier can branch directly on
\(\operatorname{LinearIndependent}(q_2-1,q_3)\). This sharpens the remaining
open work again: the no-constant exact-affine \(\dim=2\) branch is now reduced
to its genuinely independent constant-perturbed half.

The exact coprime constant-perturbed endpoint has also been widened from the
single normalized model \(1 + x_0^2\) to the full nonzero scalar family
\[
  (x_0,\ x_1,\ 1 + \alpha x_0^2,\ x_1^2), \qquad \alpha \neq 0.
\]
In \texttt{RepresentativeLowAffine.lean}, the new theorems
\texttt{quartic\_in\_image\_of\_relations\_x0\_x1\_onePlusAX0sq\_x1sq},
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_onePlusAX0sq\_x1sq},
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_x0\_x1\_onePlusAX0sq\_x1sq},
and
\texttt{residual\_eq\_zero\_of\_relations\_affinePair\_onePlusAX0sq\_x1sq}
show that the coefficient of the square line does not need to be normalized to
\(1\) before entering the exact theorem surface. The old
\texttt{onePlusX0sq\_x1sq} theorems now sit as the \(\alpha = 1\) wrappers.
This should be the right endpoint for the still-open independent translated
kernel branch, where the classifier is expected to produce a constant relation
of the form \(1 + \alpha x_0^2\) rather than literally \(1 + x_0^2\).

The accompanying numerical probe
\texttt{julia/low\_affine\_const\_perturbation\_2026\_04\_10.jl} sharpens that
picture further:
\[
\texttt{low\_affine\_const\_perturbation\_probe}
= (\texttt{samples} = 20,\ \texttt{generic\_ranks} = [15],\ \dots).
\]
Its control cases show that the coprime, diagonal, and split constant-perturbed
surfaces are all rank \(15\), while the only stable nonsurjective control is
the common-factor slice \((x_0,\ x_1,\ 1 + x_0^2,\ x_0x_1)\), which has rank
\(14\). So the remaining independent translated-kernel classifier now looks
more asymmetric than before: generically it should land in a surjective
constant-perturbed surface, and only the repeated-line/common-factor output
appears to need the codimension-one theorem layer.

That repeated-line layer is now widened on both canonical orientations. In
\texttt{RepresentativeLowAffine.lean}, the theorems
\texttt{quartic\_in\_image\_of\_relations\_x0\_x1\_onePlusAX0sq\_x0x1\_of\_coeff\_m04\_zero},
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_onePlusAX0sq\_x0x1}, and
\texttt{residual\_eq\_zero\_of\_relations\_affinePair\_onePlusAX0sq\_x0x1}
now cover
\[
  (x_0,\ x_1,\ 1 + \alpha x_0^2,\ x_0x_1),\qquad \alpha \neq 0,
\]
and the symmetric theorems
\texttt{quartic\_in\_image\_of\_relations\_x0\_x1\_onePlusBX0x1\_x1sq\_of\_coeff\_m40\_zero},
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_onePlusBX0x1\_x1sq}, and
\texttt{residual\_eq\_zero\_of\_relations\_affinePair\_onePlusBX0x1\_x1sq}
cover
\[
  (x_0,\ x_1,\ 1 + \beta x_0x_1,\ x_1^2),\qquad \beta \neq 0.
\]
So the common-factor constant-perturbed branch no longer depends on normalizing
those coefficients to \(1\) before it can enter Lean.

The new scalar sweep
\texttt{julia/low\_affine\_common\_factor\_scalars\_2026\_04\_10.jl} matches
that formal widening:
\[
\texttt{low\_affine\_common\_factor\_scalar\_probe}
= (\texttt{x0sq\_tail} = [\dots],\ \texttt{x1sq\_tail} = [\dots]).
\]
Across the tested values
\(\alpha,\beta \in \{-3,-2,-1,-\tfrac12,\tfrac12,1,2,3\}\), every
\((1+\alpha x_0^2,\ x_0x_1)\) sample had rank \(14\) with only \(x_1^4\)
missing, and every \((1+\beta x_0x_1,\ x_1^2)\) sample had rank \(14\) with
only \(x_0^4\) missing. This sharpens the remaining classifier target again:
the repeated-line/common-factor output is now numerically and formally stable
across its whole nonzero scalar family.

The same low-affine file now also contains the determinant-nondegenerate plane
wrapper
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_onePlusAX0sq\_x0x1Plane}.
So the classifier no longer needs to hand back a specially normalized common-factor
pair of the exact form \((1+\alpha x_0^2,\ x_0x_1)\). It is enough to produce
exact relations to
\[
  x_0,\quad x_1,\quad 1 + a x_0^2 + b x_0x_1,\quad c x_0^2 + d x_0x_1
\]
with \(ad-bc \neq 0\); Lean now performs the internal elimination, including
the \(d=0\) swap-to-\((1+\beta x_0x_1,\ x_1^2)\) branch and the \(d \neq 0\)
\(x_1\)-shear branch, before dispatching to the already-formalized scalar
families.

That common-factor plane surface is now symmetric and transport-ready as well.
The same file now proves the swapped determinant-nondegenerate wrapper
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_onePlusBX0x1\_x1sqPlane}
and the missing equivariant wrapper
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_x0\_x1\_onePlusAX0sq\_x0x1Plane}.
So the repeated-line/common-factor output no longer has an orientation mismatch
between its \((x_0^2,x_0x_1)\) and \((x_0x_1,x_1^2)\) plane presentations.

The same low-affine file now also contains the constant-perturbed common-factor
chart theorem
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_onePlus\_homQuadratics\_commonFactorChart}.
It assumes exact relations to \(x_0\), \(x_1\), a quadratic \(q_2\) with
constant coefficient \(1\) and zero linear terms, and a quadratic \(q_3\) with
zero constant and linear terms, together with the same cross-product chart data
\(\texttt{lowHomQuadPlaneA} \neq 0\) and
\(\texttt{lowHomQuadPlaneC} - \texttt{lowHomQuadPlaneB}^2 /
\texttt{lowHomQuadPlaneA} = 0\). Lean then performs the internal \(x_1\)-shear,
proves that both transformed quadratics have zero \(x_0^2\) coefficient, and
dispatches directly to the already formalized
\texttt{...onePlusBX0x1\_x1sqPlane} surface. So the classifier no longer needs
to pre-normalize the repeated-line/common-factor output all the way down to a
scalar representative even in the constant-perturbed branch.

On top of that, \texttt{ExactAffineClassification.lean} now proves
\texttt{residual\_eq\_zero\_of\_relations\_affinePair\_translatedKernel\_constSplit\_commonFactorChart}.
So the exact-affine \(\dim = 2\), no-constant, independent translated-kernel
branch is no longer open on the repeated-line/common-factor chart. Once the
global classifier extracts an affine pair, performs the common translation, and
produces a constant-split translated-kernel basis whose homogeneous part lands
in that chart, Lean closes the branch automatically. The remaining open work is
therefore the complementary independent translated-kernel outputs, together
with the final assembly in \texttt{TernaryQuarticProof.lean}.

That chart theorem now also sits on the same transport surface as the rest of
the low-affine layer. \texttt{RepresentativeLowAffine.lean} now proves
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_x0\_x1\_onePlus\_homQuadratics\_commonFactorChart}
and
\texttt{residual\_eq\_zero\_of\_relations\_affinePair\_onePlus\_homQuadratics\_commonFactorChart}.
So future classifier steps no longer need to inline the common translation just
to enter that chart; they can hand the branch directly to the same
\texttt{equiv}/\texttt{affinePair} interface used elsewhere.

The coprime constant-perturbed cross chart is now formalized in the same way.
\texttt{RepresentativeLowAffine.lean} now proves
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_onePlus\_homQuadratics\_crossDet\_zero},
its \texttt{equiv} wrapper, and
\texttt{residual\_eq\_zero\_of\_relations\_affinePair\_onePlus\_homQuadratics\_crossDet\_zero}.
So when the translated-kernel constant-split pair has vanishing
\texttt{lowHomQuadPlaneA} but nonzero \texttt{lowHomQuadPlaneB}, Lean now
performs the internal \(x_1\)-shear, kills the mixed term, reconstructs a
constant-perturbed diagonal pair, and dispatches directly to the already
formalized \texttt{...onePlusAX0sq\_x1sqPlane} endpoint.

On top of that, \texttt{ExactAffineClassification.lean} now proves
\texttt{residual\_eq\_zero\_of\_relations\_affinePair\_translatedKernel\_constSplit\_crossDet\_zero}.
So the exact-affine \(\dim = 2\), no-constant, independent translated-kernel
branch is no longer open on the coprime cross chart either. The remaining open
independent low-affine work is therefore the complementary diagonal/split
constant-perturbed outputs, together with the final global assembly.

One more of those complementary outputs is now discharged. First,
\texttt{RepresentativeLowAffine.lean} adds
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_onePlusX0x1\_sumsqPlane},
which transports the constant-perturbed \((x_0x_1,\ x_0^2+x_1^2)\)-plane
through the fixed split-diagonal change of variables to the already solved
\texttt{...onePlusAX0sq\_x1sqPlane} surface. On top of that, the same file now
proves
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_onePlus\_homQuadratics\_diagDiffChart},
so when the translated homogeneous pair satisfies
\[
  \texttt{lowHomQuadPlaneA} \neq 0,\qquad
  0 <
    -(\texttt{lowHomQuadPlaneC} -
      \texttt{lowHomQuadPlaneB}^2/\texttt{lowHomQuadPlaneA}) /
      \texttt{lowHomQuadPlaneA},
\]
Lean performs the internal \(x_1\)-shear and \(x_1\)-scaling, reconstructs a
constant-perturbed \((x_0x_1,\ x_0^2+x_1^2)\)-plane, and closes the branch.

That chart theorem is now lifted into
\texttt{ExactAffineClassification.lean} as
\texttt{residual\_eq\_zero\_of\_relations\_affinePair\_translatedKernel\_constSplit\_diagDiffChart}.
So the exact-affine \(\dim = 2\), no-constant, independent translated-kernel
branch is no longer open on the diagonal-difference/sumsq side either. The
remaining independent low-affine work is therefore the opposite-sign
trace-free branch, namely the \texttt{diagSumChart} output, together with the
final global assembly.

That last low-affine split is now internal too. In
\texttt{RepresentativeLowAffine.lean} the new theorem
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_onePlus\_homQuadratics\_independent}
packages the whole constant-perturbed independent low-affine branch: from
exact relations to \(x_0\), \(x_1\), \(q_2\), \(q_3\), the coefficient
conditions
\[
  \operatorname{coeff}_1(q_2)=1,\quad
  \operatorname{coeff}_{x_0}(q_2)=\operatorname{coeff}_{x_1}(q_2)=0,\quad
  \operatorname{coeff}_1(q_3)=\operatorname{coeff}_{x_0}(q_3)=
  \operatorname{coeff}_{x_1}(q_3)=0,
\]
and linear independence of \((q_2-1,q_3)\), Lean now performs the full inner
case split automatically. It dispatches to the repeated-line
\(\operatorname{span}(x_0^2,x_0x_1)\) plane when
\(\texttt{lowHomQuadPlaneA}=\texttt{lowHomQuadPlaneB}=0\), to the coprime
cross chart when \(\texttt{lowHomQuadPlaneA}=0\) but
\(\texttt{lowHomQuadPlaneB}\neq 0\), and to the common-factor, diagonal-sum,
or diagonal-difference charts when \(\texttt{lowHomQuadPlaneA}\neq 0\).
The matching affine-pair wrapper
\texttt{residual\_eq\_zero\_of\_relations\_affinePair\_onePlus\_homQuadratics\_independent}
puts that result on the same transport surface as the rest of the low-affine
theory.

On the affine-rank-one side, one more constant-tail chart is now internal too.
In \texttt{RepresentativeAffineRankOne.lean} the new theorem
\texttt{residual\_eq\_zero\_of\_relations\_x0\_onePlus\_homQuadratics\_crossDet\_zero}
handles the coprime cross chart for a single exact affine line. From exact
relations to \(x_0\), one quadratic \(q_1\) with affine part \(1\), and a pure
homogeneous pair \((q_2,q_3)\) with
\(\texttt{lowHomQuadPlaneA}(q_2,q_3)=0\) and
\(\texttt{lowHomQuadPlaneB}(q_2,q_3)\neq 0\), Lean now performs the internal
\(x_1\)-shear, converts the homogeneous pair to an
\((x_0^2,x_1^2)\)-plane basis, and dispatches to the already formalized
\texttt{onePlus\_homQuadratics\_x0sq\_x1sqPlane} theorem. So this affine-rank-one
branch no longer needs a classifier that pre-normalizes the homogeneous pair all
the way down to explicit \((x_0^2,x_1^2)\) coordinates before entering the
endpoint layer.

The repeated-line/common-factor chart is now internal on the same pure
constant surface. The new theorem
\texttt{residual\_eq\_zero\_of\_relations\_x0\_onePlus\_homQuadratics\_commonFactorChart}
performs the internal \(x_1\)-shear when
\(\texttt{lowHomQuadPlaneA}(q_2,q_3)\neq 0\) and the post-shear diagonal term
vanishes, rewrites the homogeneous pair into
\(\operatorname{span}(x_0x_1,x_1^2)\), and dispatches to the already solved
\texttt{onePlus\_homQuadratics\_x0x1\_x1sqPlane} theorem. So the pure
constant affine-rank-one layer is now internal on both the coprime cross and
common-factor charts; the remaining diagonal-complement work is separate.

On top of that, \texttt{ExactAffineClassification.lean} now proves
\texttt{residual\_eq\_zero\_of\_relations\_affinePair\_translatedKernel\_constSplit\_independent}
and the global classifier theorem
\texttt{residual\_eq\_zero\_of\_exactAffineDimTwo\_noConst}. So the whole
exact-affine \(\dim=2\), no-constant branch is now closed in Lean: once the
exact affine kernel has dimension \(2\) and contains no exact constant
relation, Lean reconstructs an exact affine pair, analyzes the translated
zero-linear-tail kernel, and dispatches internally to either the homogeneous
low-affine branch or the constant-perturbed low-affine branch. The remaining
global work is no longer in the exact-affine \(\dim=2\) classifier. It is now
the exact-affine \(\dim=1\) classifier: one must either exclude or directly
formalize the tailed families with one exact affine line and a second affine
direction carrying quadratic tail, and then assemble the resulting endpoint
theorems in \texttt{TernaryQuarticProof.lean}.

\section*{Planned File Map}
\begin{itemize}
\item \texttt{TernaryQuarticProof/Socp.lean}: abstract SOCP consequences.
\item \texttt{TernaryQuarticProof/Certificate.lean}: admissible transformed
image-plus-cone certificate.
\item \texttt{TernaryQuarticProof/FactorTransform.lean}: orthogonal
factor-coordinate invariance.
\item \texttt{TernaryQuarticProof/QuadraticNormalForm.lean}: finite-support
normal form for two-variable quadratics;
\item one future algebra module for the remaining representative image and
image-plus-kernel decompositions;
\item \texttt{TernaryQuarticProof.lean}: final theorem assembly.
\end{itemize}

\end{document}
