\documentclass[11pt]{article}
\usepackage{amsmath,amssymb}
\usepackage[margin=1in]{geometry}

\title{Ternary Quartic Rank-4 Exploration Log}
\date{2026-04-09}

\begin{document}
\maketitle

\section*{Current Strategy}
The current proof search is split into two layers:
\begin{enumerate}
\item isolate the abstract SOCP certificate lemmas that depend only on the
definitions in \texttt{TernaryQuartic.lean};
\item use Julia to determine what algebraic decomposition should replace the
univariate factor-peeling argument in the ternary quartic rank-4 setting.
\end{enumerate}

The key modeling difference from the univariate reference is that the SOCP
conditions only quantify over admissible directions, i.e. rank-4 vectors of
quadratic polynomials. Any image-plus-cone certificate therefore has to keep all
witness directions quadratic.

\section*{Lean Progress}
Added the following reusable proof modules.
\begin{itemize}
\item \texttt{TernaryQuarticProof/Socp.lean}: bilinear simp lemmas, the FOCP
consequence $\langle \sigma(u), \sigma(u)-p \rangle = 0$, and positivity lemmas
for the objective.
\item \texttt{TernaryQuarticProof/Certificate.lean}: admissible
image-plus-cone definitions and the abstract theorem that any admissible
transformed image-plus-cone decomposition forces zero residual at a SOCP.
\item The same file now also contains two concrete reductions:
  \begin{itemize}
  \item if $p \in \operatorname{im}(A_u)$, then the residual vanishes directly;
  \item if the four coordinates of $u$ satisfy a nontrivial constant linear
  relation, then every SOS quartic lies in the admissible kernel cone and the
  residual again vanishes.
  \end{itemize}
\item The same certificate layer now also contains an indexed-family version of
  the admissible image-plus-kernel-square argument, so SOS decompositions can
  be summed directly over \texttt{Fin k} without deduplicating kernel
  directions into a \texttt{Finset}.
\item \texttt{TernaryQuarticProof/FactorTransform.lean}: orthogonal mixing of
  the four factor coordinates, with formal proofs that admissibility, the
  linearized image map, $\sigma$, the residual, FOCP, and SOCP are preserved
  under orthogonal recombination.
\item \texttt{TernaryQuarticProof/RepresentativeSurjective.lean}: the explicit
  representative
  \[
    (1, x_1^2, x_1x_2, x_2^2)
  \]
  is now proved surjective on quartics by a monomial-by-monomial image
  construction. This discharges the constant $\dim(W \cap \mathbb{R}[x]_{\le 1})
  = 1$ branch at the representative level.
\item \texttt{TernaryQuarticProof/QuadraticNormalForm.lean}: a proof-serving
  six-monomial support theorem for two-variable quadratics. It formalizes that
  any \texttt{IsQuadratic} polynomial has support contained in exactly the six
  monomials
  \[
    1,\ x_1,\ x_2,\ x_1^2,\ x_1x_2,\ x_2^2,
  \]
  and hence equals the sum of its six monomial-coefficient terms.
\item \texttt{TernaryQuarticProof/RepresentativeMixedAffine.lean}: the
  two nonsurjective mixed-affine representatives
  \[
    (1, x_1, x_1x_2, x_2^2),
    \qquad
    (1, x_1, x_1^2, x_1x_2)
  \]
  are now fully discharged at the representative level. For the rank-$14$ model
  the file proves both that any quartic with zero $x_1^4$ coefficient lies in
  $\operatorname{im}(A_u)$ and that every SOCP at this representative has zero
  residual. For the rank-$13$ model it proves the analogous codimension-two
  image theorem with vanishing $x_2^3$ and $x_2^4$ coefficients, the quadratic
  square coefficient identities for those two classes, and the representative
  no-spurious-SOCP theorem using the explicit kernel line
  \[
    (0,0,-x_2(b+c x_2),x_1(b+c x_2)).
  \]
  The same file now also formalizes the surjective mixed-affine model
  \[
    (1, x_1, x_1^2, x_2^2),
  \]
  by a direct quartic image-surjectivity argument.
\item \texttt{TernaryQuarticProof/RepresentativeSpanThree.lean}: the generic
  representative
  \[
    (1,x_1,x_2,h)
  \]
  with arbitrary homogeneous quadratic $h$ is now also discharged at the
  representative level. The file splits a quadratic factor as
  $q=q_{\le 1}+q_{\mathrm{hom}}$, forms the kernel direction
  \[
    (-q_{\mathrm{hom}},a,b,0),\qquad q_{\mathrm{hom}}=x_1 a + x_2 b,
  \]
  proves that the corrected square
  \[
    q^2-\sigma(-q_{\mathrm{hom}},a,b,0)
  \]
  has total degree at most $3$, and then sums these cubic image terms over an
  SOS decomposition using the indexed-family certificate.
\item \texttt{TernaryQuarticProof/AffineTransform.lean}: a new affine
  substitution layer. It formalizes the polynomial algebra homomorphism
  induced by an affine change of variables, proves that affine substitutions do
  not increase total degree, and deduces preservation of the
  \texttt{IsQuadratic} and \texttt{IsQuartic} bounds. The same file also proves
  that applying such a homomorphism coordinatewise commutes with $A_u$,
  $\sigma$, and the residual.
\item \texttt{TernaryQuarticProof/AffineSocpTransform.lean}: the invertible
  half of the affine story is now formalized too. The file transports the
  bilinear form across a polynomial algebra equivalence, proves preservation of
  positive-definiteness, FOCP, SOCP, admissibility, SOS quarticity, and zero
  residual, and packages these statements for explicit affine inverses.
\item \texttt{TernaryQuarticProof/RepresentativeTransport.lean}: a new
  packaging layer now combines the affine transport and orthogonal mixing
  invariance. Its main theorem says that once exact normalization data are
  available in the form
  \[
    \operatorname{mix}(M^\top, e^{-1}\!\cdot u) = u_{\mathrm{rep}},
  \]
  any representative-level no-spurious-SOCP theorem immediately pulls back to
  the original SOCP point.
\end{itemize}

These files build with
\begin{verbatim}
lake build TernaryQuarticProof.Basic
\end{verbatim}
and the full root project builds with \texttt{lake build}. The verification
harness still fails at
\begin{verbatim}
Unknown identifier `TernaryQuartic.ternaryQuartic_rankFour_no_spurious_socp`
\end{verbatim}
because the final theorem has not been assembled in
\texttt{TernaryQuarticProof.lean}.

\section*{Julia Experiments}
Recorded script:
\begin{verbatim}
julia/tq_r4_exploration_2026_04_09.jl
\end{verbatim}

Random rank-4 sweep command:
\begin{verbatim}
cd /agent-workspace/julia
julia --project=. tq_r4_exploration_2026_04_09.jl
\end{verbatim}

Observed random sweep output:
\begin{verbatim}
random_rank4_errs = [6.351865185047683e-11, 8.956485306788336e-12,
2.6531504105737066e-11, 1.305916626190534e-10, 7.704433630342462e-11]
\end{verbatim}

Interpretation: all five sampled rank-4 instances were numerically consistent
with ``no counterexample found'' for the fixed-$u$ SDP search.

Observed basis sweep summary:
\begin{verbatim}
basis_sweep_summary = (cases = 15, counters = 0, unknowns = 0)
\end{verbatim}

Each of the 15 monomial-basis rank-4 subspaces of the quadratic space produced
an \texttt{INFEASIBLE} homogeneous SDP, again consistent with the conjectured
absence of spurious SOCPs.

Representative dual probe (same recorded Julia script, basis
$u = (1, x_2, x_1, x_2^2)$):
\begin{verbatim}
dual_probe_status = (termination = INFEASIBLE,
  dual = INFEASIBILITY_CERTIFICATE, grad_constraints = 24)
\end{verbatim}

This confirms that CSDP returns an explicit infeasibility certificate for at
least one basis case, so the next exploration step should be to decode the
gradient/PSD/SOS dual objects into a symbolic admissible image-plus-cone
certificate.

Image-rank scan for the 15 monomial bases showed ranks only in
$\{13,14,15\}$. In particular, several cases are already surjective
(\texttt{rank = 15}), hence formally reducible to the new Lean lemma
\texttt{residual\_eq\_zero\_of\_in\_admissible\_image}.

Normal-form probe:
\begin{verbatim}
normal_form_probe = (u = [1, x1, x2, x1^2 + x2^2], rank = 13, err ~ 1e-11)
normal_form_probe = (u = [1, x1, x2, x1^2 - x2^2], rank = 13, err ~ 1e-12)
normal_form_probe = (u = [1, x1, x2, x1*x2 + x1^2], rank = 13, err ~ 1e-12)
normal_form_probe = (u = [1, x1, x2, x1*x2 + x2^2], rank = 13, err ~ 1e-11)
normal_form_probe = (u = [x1, x2, x1^2 + x2^2, x1*x2], rank = 14, err ~ 1e-12)
normal_form_probe = (u = [x1, x2, x1^2 - x2^2, x1*x2], rank = 14, err ~ 1e-11)
\end{verbatim}

This suggests the main classification is by the dimension of the affine
lower-degree part of $\operatorname{span}(u_1,\dots,u_4)$ rather than by the
factorization type of the extra quadratic.

Explicit kernel-family probe for $u=(1,x_2,x_1,x_2^2)$:
\begin{verbatim}
w = (-x1^2, 0, x1, 0),  A_u(w)=0,  sigma(w)=x1^4 + x1^2
\end{verbatim}

Since $x_1^2 \in \operatorname{im}(A_u)$ via the first coordinate $1$, this
shows that $x_1^4$ is represented modulo $\operatorname{im}(A_u)$ by an
admissible kernel square.

\section*{Current Hypothesis}
The abstract SOCP step is not the blocker. The missing ingredient is a
ternary-quartic algebraic statement of the following form:
\[
  p \in \operatorname{im}(A_{u_{\mathrm{img}}}) +
      \operatorname{cone}(\sigma(\ker A_u))
\]
with all witnesses admissible, for every admissible rank-4 factor $u$ and every
SOS quartic $p$.

This is the multivariate replacement for the univariate factor-peeling/hgroup
reduction. The admissibility restriction rules out naively importing the
univariate kernel generators, because multiplying a quadratic factor by a
nonconstant polynomial leaves the quadratic direction space.

The current best candidate for the remaining proof is a normal-form
classification by the filtered dimensions
\[
  \dim(W \cap \mathbb{R}), \qquad \dim(W \cap \mathbb{R}[x_1,x_2]_{\le 1}),
  \qquad W = \operatorname{span}(u_1,\dots,u_4).
\]
The newly formalized dependent case handles $\dim W < 4$, and the surjective
case handles image rank $15$. The whole
$\dim(W \cap \mathbb{R}[x]_{\le 1}) = 3$ branch is now solved in Lean through
an abstract ``contains $1,x_1,x_2$'' theorem, not just at one representative.
The unresolved cases are therefore normalization statements reducing the
independent $\dim = 2$ configurations of image rank $13$ or $14$ to the
already-formalized representatives.

Current canonical candidates after orthogonal mixing of the four coordinates and
affine changes of variables:
\begin{itemize}
\item $\dim(W \cap \mathbb{R}[x]_{\le 1}) = 3$:
  \[
    u \sim (1, x_1, x_2, h), \qquad h \in \mathbb{R}[x_1,x_2]_2^{\mathrm{hom}}.
  \]
  More strongly, this branch no longer needs exact normalization to that
  representative. It suffices that $1,x_1,x_2 \in W$. The kernel family is:
  if $h_0 = x_1 a + x_2 b$ with $a,b$ linear, then
  \[
    w^{\mathrm{raw}}
    =
    \operatorname{rel}(-c_0,h_0)+\operatorname{rel}(c_1,a)+\operatorname{rel}(c_2,b),
  \]
  where $\sum_i c_{0,i}u_i = 1$, $\sum_i c_{1,i}u_i = x_1$, and
  $\sum_i c_{2,i}u_i = x_2$. After scaling by
  $\bigl(\sqrt{\sum_i c_{0,i}^2}\bigr)^{-1}$, the corrected square has degree
  at most $3$, so every quadratic square is in
  $\operatorname{im}(A_u)+\operatorname{cone}(\sigma(\ker A_u))$ without
  further normalization data.
\item $\dim(W \cap \mathbb{R}[x]_{\le 1}) = 2$, common-factor case:
  \[
    u \sim (x_1, x_2, x_1^2, x_1 x_2).
  \]
  The image misses exactly the classes of $1$ and $x_2^4$ in the monomial
  model. Explicit kernel directions:
  \[
    w_c = (-2x_2, x_1, 0, 1), \qquad
    w_q = (-x_2^2, x_1 x_2, 0, 0).
  \]
  Then $\sigma(w_c) \equiv 1$ and $\sigma(w_q) \equiv x_2^4$ modulo
  $\operatorname{im}(A_u)$.
\item $\dim(W \cap \mathbb{R}[x]_{\le 1}) = 2$, coprime case:
  \[
    u \sim (x_1, x_2, x_1^2, x_2^2)
  \]
  or any equivalent rank-$14$ model. The image already contains every positive
  degree monomial; only the constant class remains, and it is represented by
  \[
    w = (-x_1, 0, 1, 0), \qquad \sigma(w)=x_1^2+1 \equiv 1.
  \]
\item $\dim(W \cap \mathbb{R}[x]_{\le 1}) = 1$, constant case:
  \[
    u \sim (1, x_1^2, x_1 x_2, x_2^2),
  \]
  which is already surjective.
\item $\dim(W \cap \mathbb{R}[x]_{\le 1}) = 1$, linear case:
  \[
    u \sim (x_1, x_1^2, x_1 x_2, x_2^2).
  \]
  The image misses exactly the classes of $1$ and $x_2$. For any affine linear
  $\ell = c + d x_2$,
  \[
    w_\ell = (-x_2 \ell, 0, \ell, 0)
  \]
  lies in the kernel, and
  \[
    \sigma(w_\ell) = (1+x_2^2)\ell^2 \equiv c^2 + 2 c d\, x_2
    \pmod{\operatorname{im}(A_u)}.
  \]
  This matches the quotient classes of an arbitrary quadratic square.
\item \textbf{Correction.} The previous five-case list is not complete. When
  $\dim(W \cap \mathbb{R}[x]_{\le 1}) = 2$ and $1 \in W$, there is an
  additional mixed-affine subbranch. The representative probes
  \[
    (1,x_1,x_1^2,x_1x_2),\qquad
    (1,x_1,x_1x_2,x_2^2),\qquad
    (1,x_1,x_1^2,x_2^2)
  \]
  have image ranks $13$, $14$, and $15$ respectively, with missing monomial
  classes
  \[
    \{x_2^3,x_2^4\},\qquad \{x_1^4\},\qquad \varnothing.
  \]
  The first two also admit explicit kernel probes:
  \[
    w = (0,0,-x_2^2,x_1x_2)
  \]
  for the rank-$13$ model, giving
  \[
    A_u(w)=0,\qquad \sigma(w)=x_2^4+x_1^2x_2^2,
  \]
  and
  \[
    w = (0,0,-x_1x_2,x_1^2)
  \]
  for the rank-$14$ model, giving
  \[
    A_u(w)=0,\qquad \sigma(w)=x_1^2x_2^2+x_1^4.
  \]
  Modulo $\operatorname{im}(A_u)$ these produce the missing $x_2^4$ and $x_1^4$
  classes, and the rank-$13$ family extends to
  \[
    w_\ell=(0,0,-x_2\ell,x_1\ell),\qquad \ell=c+d x_2,
  \]
  which generates both missing classes $x_2^3$ and $x_2^4$ modulo the image.
  A further parabolic probe of the same branch gives
  \[
    (1,x_1,x_1x_2,x_1^2-x_2^2),\qquad
    (1,x_1,x_1x_2,x_1^2+x_2^2)
  \]
  both with image rank $15$. This suggests the following sharper
  classification: after fixing
  \[
    W \cap \mathbb{R}[x_1,x_2]_{\le 1} = \operatorname{span}(1,x_1),
  \]
  the remaining quadratic plane is the kernel of a binary quadratic
  $a x_1^2 + b x_1x_2 + c x_2^2$, and only the annihilators $x_2^2$ and
  $x_1^2$ give nonsurjective image ranks ($13$ and $14$ respectively). The
  new annihilator-family sweep in the checked-in Julia script supports the full
  completion-of-squares reduction:
  \begin{itemize}
  \item if $c \neq 0$, shear $x_2 \mapsto x_2 - \frac{b}{2c}x_1$ kills the
    cross term;
  \item if the resulting $x_1^2$ coefficient vanishes, the branch is the
    rank-$13$ model $x_2^2$;
  \item if that coefficient is nonzero, rescaling gives $x_1^2 \pm x_2^2$,
    numerically always surjective;
  \item if $c = 0$ but $b \neq 0$, the model reduces to $x_1x_2$, again
    numerically surjective;
  \item if $c = 0$ and $b = 0$, the model is the rank-$14$ case $x_1^2$.
  \end{itemize}
  The shear-and-scale formulas underlying this reduction are now formalized in
  \texttt{TernaryQuarticProof/MixedAffineNormalization.lean}. That file proves
  the explicit affine identities sending
  \[
    a x_1^2 + b x_1x_2 + c x_2^2
  \]
  to
  \[
    \Bigl(a - \frac{b^2}{4c}\Bigr) x_1^2 + c x_2^2
    \qquad (c \neq 0),
  \]
  and to
  \[
    b x_1x_2
    \qquad (c=0,\ b \neq 0),
  \]
  inside the affine subgroup preserving the line $\mathbb{R}x_1$.
  In this round the same file is tightened further: it now also proves the
  diagonal sign normalizations
  \[
    a x_1^2 + c x_2^2 \leadsto a(x_1^2+x_2^2)
    \qquad \text{when } a/c>0,
  \]
  and
  \[
    a x_1^2 + c x_2^2 \leadsto a(x_1^2-x_2^2)
    \qquad \text{when } -a/c>0,
  \]
  via the scalings
  \(x_2 \mapsto \sqrt{a/c}\,x_2\) and
  \(x_2 \mapsto \sqrt{-a/c}\,x_2\), and it combines those with the shear step
  into explicit shear-plus-scale normal-form equalities. So the algebraic
  normalization layer now already covers the three post-shear outcomes that the
  mixed-affine classification needs: rank-$14$, sum-square, and split-diagonal.
  So the remaining classification gap is specifically the constant-containing
  $\dim = 2$ branch rather than the whole independent nonsurjective case.
\end{itemize}

The remaining gap is no longer the representative algebra in these cases; it is
the normalization step that extracts, from a general admissible factor, the
quadratic annihilator and the coefficient vectors needed to apply those
formalized shear/scale identities inside Lean. After that data is produced, the
transport layer can pull back the already-formalized representative theorems.

One smaller alternative is now on the table for the mixed-affine branch:
replace exact representative transport by basis-invariant plane theorems.
For a plane of the form
\[
  x_1 \cdot \operatorname{span}(x_1,x_2)
  \qquad\text{or}\qquad
  x_2 \cdot \operatorname{span}(x_1,x_2),
\]
the image part depends only on the plane itself, not on the chosen basis of
its two coordinates. If one chooses an orthonormal coefficient-space basis of
that plane, the same syzygy as in the representative proofs produces kernel
families with a clean
\[
  \sigma = (a^2+b^2)\ell^2
\]
formula. This may avoid the last orthogonal-normalization obstruction in the
constant-containing $\dim = 2$ branch.

The first formal Lean infrastructure for that alternative is now added in
\texttt{TernaryQuarticProof/RepresentativeMixedAffinePlane.lean}. It proves:
\begin{itemize}
\item any explicit scalar relation $\sum_i c_i u_i=r$ produces an admissible
  image witness for $r q$;
\item if coefficient vectors $c,d \in \mathbb{R}^4$ are orthonormal, then
  their relation directions satisfy
  \[
    \sigma\bigl(\mathrm{rel}_c(p)+\mathrm{rel}_d(q)\bigr)=p^2+q^2.
  \]
\item the whole image half of the rank-$14$ mixed-affine representative proof
  is now abstracted away from the concrete representative: if a factor admits
  explicit relations for $1$, $x_1$, $x_1x_2$, and $x_2^2$, then every quartic
  with zero $x_1^4$ coefficient lies in its admissible image.
\item the residual argument for that branch is now abstracted as well: with the
  same relation data and an orthonormal coefficient-space pair for the
  $(x_1x_2,x_2^2)$ directions, the kernel
  \[
    w=\mathrm{rel}_{-c_2}(t x_1x_2)+\mathrm{rel}_{c_3}(t x_1^2)
  \]
  has
  \[
    \operatorname{coeff}_{x_1^4}(\sigma(w))=t^2
  \]
  and the admissible image-plus-cone certificate gives zero residual.
\item the rank-$13$ branch is now abstracted in the same basis-invariant way:
  from explicit relations for $1$, $x_1^2$, and $x_1x_2$, together with an
  orthonormal coefficient-space pair for the last two directions, the file now
  proves both the codimension-two image theorem and the full SOCP certificate.
  The abstract kernel is
  \[
    w=\mathrm{rel}_{-c_2}(x_2\ell)+\mathrm{rel}_{c_3}(x_1\ell),
    \qquad \ell=b+c x_2,
  \]
  and its $(x_2^3,x_2^4)$ coefficients match the representative formulas
  $(2bc,c^2)$.
\item the same file now also formalizes the surjective rank-$15$ plane orbits
  directly: any invertible basis of
  \[
    \operatorname{span}(x_1^2,x_2^2)
    \qquad\text{or}\qquad
    \operatorname{span}(x_1x_2,\ x_1^2-x_2^2)
  \]
  yields a no-spurious-SOCP theorem, and both statements are wrapped through
  the affine-equivalence transport layer.
\item the stronger follow-up abstraction is now split in two. The fully
  arbitrary orthonormal-basis version for the canonical planes
  $x_1 \cdot \operatorname{span}(x_1,x_2)$ and
  $x_2 \cdot \operatorname{span}(x_1,x_2)$ still ran into unstable coefficient
  bookkeeping for products of general linear forms and remains reverted. But a
  cleaner partial theorem is now formalized: if the quadratic relations differ
  from the monomial pairs $(x_1x_2,x_2^2)$ or $(x_1^2,x_1x_2)$ by an orthogonal
  $2 \times 2$ basis change inside that plane, then Lean reconstructs the
  exact monomial relations and reduces back to the existing rank-$14$ or
  rank-$13$ mixed-affine theorem. So the obstruction has narrowed from
  ``arbitrary plane-basis algebra'' to ``produce orthogonal in-plane basis
  data from a general mixed-affine factor.''
\item that new partial strengthening is now connected to the affine transport
  layer as well: the same file now wraps the rank-$14$ and rank-$13$ orthogonal
  in-plane basis theorems through an arbitrary polynomial algebra equivalence.
  In practical terms, once an affine normalization sends a mixed-affine factor
  to one of these orthogonal plane-basis configurations, Lean can conclude
  zero residual for the original SOCP point without building a separate
  transport proof each time.
\end{itemize}
So the remaining work for the plane strategy is now geometric rather than
algebraic: identify the right orthonormal relation data canonically from a
general mixed-affine factor.

The newest strategy adjustment is therefore to split the old ``normalization''
problem into two formally separate layers:
\begin{itemize}
\item affine variable-change transport on the polynomial side;
\item orthogonal factor-coordinate mixing on the rank-$4$ side.
\end{itemize}
The second layer was already formalized in
\texttt{TernaryQuarticProof/FactorTransform.lean}. The first layer is now
itself split into a homomorphism file
\texttt{TernaryQuarticProof/AffineTransform.lean} and an equivalence-transport
file \texttt{TernaryQuarticProof/AffineSocpTransform.lean}. This shift came
from a concrete obstruction: orthogonal mixing acts on coefficient vectors in
$\mathbb{R}^4$, so it does not by itself force exact coordinates $1$, $x_1$,
or $x_2$ to appear even when the corresponding affine lines lie in
$\operatorname{span}(u_1,\dots,u_4)$.

The next proof-serving abstraction is therefore more concrete: introduce an
explicit six-coordinate normal form for quadratics,
\[
  a_{00} + a_{10} x_1 + a_{01} x_2 + a_{20} x_1^2 + a_{11} x_1x_2 + a_{02} x_2^2,
\]
and formalize how the affine subgroup preserving $\mathbb{R}x_1$ acts on those
coordinates. The aim is not a generic symbolic algebra package; it is a small
normalization API tailored to the remaining constant-containing mixed-affine
branch. In particular, one wants explicit formulas for:
\begin{itemize}
\item translation in $x_2$, to absorb affine tails into the already-controlled
  $\operatorname{span}(1,x_1)$ part;
\item shear $x_2 \mapsto x_2 + t x_1$, to perform the completion-of-squares
  reduction on the homogeneous coefficients;
\item scaling $x_2 \mapsto s x_2$, to normalize the remaining nonzero diagonal
  coefficient.
\end{itemize}
This should reduce the last unresolved branch to coefficient extraction and a
binary-quadratic orbit calculation, rather than to another ad hoc representative
proof.

That coefficient layer is now strengthened in Lean. The same file
\texttt{TernaryQuarticProof/QuadraticCoordinateForm.lean} now also:
\begin{itemize}
\item defines the actual affine equivalence for translation $x_2 \mapsto x_2+t$
  inside the existing transport framework;
\item proves generic coefficient identities for arbitrary quadratics under that
  translation, for example
  \[
    \operatorname{coeff}_{x_2}(q(x_1,x_2+t))
    =
    \operatorname{coeff}_{x_2}(q)+2t\,\operatorname{coeff}_{x_2^2}(q),
  \]
  while the homogeneous coefficients
  \[
    \operatorname{coeff}_{x_1^2},\ \operatorname{coeff}_{x_1x_2},\ \operatorname{coeff}_{x_2^2}
  \]
  remain unchanged;
\item proves the analogous generic coefficient-update rules under shear and
  scaling.
\end{itemize}
So the mixed-affine normalization problem is now expressed in Lean directly at
the level of coefficient functionals on arbitrary quadratics, rather than only
through hand-expanded representative formulas.

The next strategy refinement is now very narrow: formalize the coefficient-
selected normalization parameters themselves. For a quadratic with coefficients
\[
  (a_{00},a_{10},a_{01},a_{20},a_{11},a_{02}),
\]
the planned Lean lemmas are exactly the scalar choices used in the paper proof:
\begin{itemize}
\item if $a_{02}\neq 0$, translate by
  \[
    t_{\mathrm{tr}}=-\frac{a_{01}}{2a_{02}}
  \]
  to kill the $x_2$ term;
\item if $a_{02}\neq 0$, shear by
  \[
    t_{\mathrm{sh}}=-\frac{a_{11}}{2a_{02}}
  \]
  to kill the $x_1x_2$ term and replace the $x_1^2$ coefficient by
  \[
    a_{20}-\frac{a_{11}^2}{4a_{02}};
  \]
\item if $a_{02}=0$ but $a_{11}\neq 0$, shear by
  \[
    t_{\mathrm{sh}}=-\frac{a_{20}}{a_{11}}
  \]
  to kill the $x_1^2$ term and reduce to the pure-cross orbit.
\end{itemize}
This algebra is no longer only a blueprint item: the same file now packages it
as the Lean theorem
\texttt{mixedAffineAnnihilator\_normal\_form\_cases}, giving a direct case split
on the annihilator coefficients.

\section*{Additional Recorded Experiments}
The checked-in Julia script now also records:
\begin{verbatim}
low_affine_probe = (u = [1, x1^2, x1*x2, x2^2], rank = 15, err ~ 1e-11)
low_affine_probe = (u = [x1, x1^2, x1*x2, x2^2], rank = 13, err ~ 1e-10)
low_affine_probe = (u = [x1, x2, x1^2, x1*x2], rank = 13, err ~ 1e-11)
low_affine_probe = (u = [x1, x2, x1^2 + x2^2, x1*x2], rank = 14, err ~ 1e-12)
\end{verbatim}

These are the representative models behind the five canonical cases above.

The same Julia script now also records the mixed-affine branch:
\begin{verbatim}
mixed_affine_probe = (u = [1, x1, x1^2, x1*x2], rank = 13,
  missing = [x2^3, x2^4], err ~ 1e-10)
mixed_affine_probe = (u = [1, x1, x1*x2, x2^2], rank = 14,
  missing = [x1^4], err ~ 1e-11)
mixed_affine_probe = (u = [1, x1, x1^2, x2^2], rank = 15,
  missing = [], err ~ 1e-11)
mixed_affine_kernel_probe = (u = [1, x1, x1^2, x1*x2],
  w = [0, 0, -x2^2, x1*x2], Auw = 0,
  sigmaw = x2^4 + x1^2*x2^2)
mixed_affine_kernel_probe = (u = [1, x1, x1*x2, x2^2],
  w = [0, 0, -x1*x2, x1^2], Auw = 0,
  sigmaw = x1^2*x2^2 + x1^4)
mixed_affine_parabolic_probe = (u = [1, x1, x1*x2, x1^2 - x2^2],
  rank = 15, err ~ 1e-12)
mixed_affine_parabolic_probe = (u = [1, x1, x1*x2, x1^2 + x2^2],
  rank = 15, err ~ 1e-11)
mixed_affine_annihilator_probe = (annihilator = x2^2,
  rank = 13, missing = [x2^3, x2^4], err ~ 1e-10)
mixed_affine_annihilator_probe = (annihilator = x1^2,
  rank = 14, missing = [x1^4], err ~ 1e-12)
mixed_affine_annihilator_probe = (annihilator = x1*x2,
  rank = 15, missing = [], err ~ 1e-12)
mixed_affine_annihilator_probe = (annihilator = x1^2 + x2^2,
  rank = 15, missing = [], err ~ 1e-11)
mixed_affine_annihilator_probe = (annihilator = x1^2 - x2^2,
  rank = 15, missing = [], err ~ 1e-12)
mixed_affine_annihilator_probe = (annihilator = x1^2 + x1*x2,
  rank = 15, missing = [], err ~ 1e-10)
mixed_affine_annihilator_probe = (annihilator = generic (2,3,5),
  rank = 15, missing = [], err ~ 2e-10)
mixed_affine_annihilator_probe = (annihilator = generic_disc0 (1,2,1),
  rank = 15, missing = [], err ~ 1e-12)
\end{verbatim}

This corrects the earlier assumption that the five listed canonical models
already exhausted the independent rank-$4$ cases.

\section*{New Lean Strategy Adjustment}
The first representative theorem suggests a better formalization pattern than
generic coefficient reconstruction. For explicit normal forms, it is often
easier to prove either:
\begin{itemize}
\item full image surjectivity by reducing quartics to monomial cases; or
\item an image-plus-kernel decomposition for each quadratic square
  $q^2$, using a decomposition $q = r + t$ where $r$ already lies in the span
  of the factor coordinates and $t$ is the small missing piece.
\end{itemize}

This avoids the failed earlier attempt to build a large generic monomial-basis
library first.

The next candidate abstraction was smaller and more targeted than the abandoned
generic coefficient library: prove a \emph{quadratic normal-form} lemma for
two-variable quadratics by using the support restriction coming from
\texttt{IsQuadratic} and the fact that every support monomial of total degree
at most $2$ is one of exactly six finitely many \texttt{Fin 2 ->0 Nat} patterns.

This step is now formalized in
\texttt{TernaryQuarticProof/QuadraticNormalForm.lean}. The file introduces the
six quadratic monomial indices
\[
  m_{00},\ m_{10},\ m_{01},\ m_{20},\ m_{11},\ m_{02},
\]
proves a case split for all exponent vectors of total degree at most $2$,
proves that the support of any quadratic polynomial is contained in this six
element set, and derives the six-monomial sum formula
\[
  q = \sum_{d \in \{m_{00},m_{10},m_{01},m_{20},m_{11},m_{02}\}}
      \operatorname{monomial}(d,\operatorname{coeff}_d(q)).
\]

The new proof strategy shift is therefore no longer ``build a quadratic normal
form'' but ``use the formalized quadratic normal form to drive a codimension-one
or codimension-two image argument for the mixed-affine representatives.'' The
first target
\[
  u = (1,x_1,x_1x_2,x_2^2)
\]
is now complete: the codimension-one image theorem, the coefficient-of-square
lemma for the missing $x_1^4$ class, and the explicit kernel correction have
all been formalized, giving a full representative-level no-spurious-SOCP
theorem.

That second mixed-affine target is now complete as well. So the constant-
containing mixed-affine branch is no longer blocked on representative algebra:
the two nonsurjective representatives
\[
  (1,x_1,x_1^2,x_1x_2),\qquad
  (1,x_1,x_1x_2,x_2^2)
\]
are covered in Lean with explicit admissible kernel families, and the surjective
rank-$15$ branch is covered at the plane level for both real orbits
\[
  \operatorname{span}(x_1^2,x_2^2)
  \qquad\text{and}\qquad
  \operatorname{span}(x_1x_2,\ x_1^2-x_2^2).
\]

The low-affine representative analysis is now complete as well in
\texttt{TernaryQuarticProof/RepresentativeLowAffine.lean}. The coprime model
\[
  (x_1,x_2,x_1^2,x_2^2),
\]
the common-factor model
\[
  (x_1,x_2,x_1^2,x_1x_2),
\]
and the linear model
\[
  (x_1,x_1^2,x_1x_2,x_2^2)
\]
all now have representative-level no-spurious-SOCP theorems in Lean. In the
linear case, the image misses exactly the classes of $1$ and $x_2$, and the
kernel line
\[
  w_\ell = (-x_2\ell,0,\ell,0), \qquad \ell = c + d x_2,
\]
matches those missing classes via
\[
  \sigma(w_\ell) \equiv c^2 + 2cd\,x_2 \pmod{\operatorname{im}(A_u)}.
\]

The span-three file now goes beyond the representative
\[
  u = (1,x_1,x_2,h).
\]
It proves an abstract theorem: whenever an admissible rank-$4$ factor $u$
contains $1,x_1,x_2$ in its span, every SOCP at $u$ has zero residual. The raw
kernel direction attached to a quadratic factor $q=q_{\le 1}+q_{\mathrm{hom}}$
is built from the coefficient vectors for $1,x_1,x_2$:
\[
  w_q^{\mathrm{raw}}
  =
  \operatorname{rel}(-c_0,q_{\mathrm{hom}})
  + \operatorname{rel}(c_1,a)
  + \operatorname{rel}(c_2,b),
  \qquad q_{\mathrm{hom}} = x_1 a + x_2 b.
\]
After scaling by $\bigl(\sqrt{\sum_i c_{0,i}^2}\bigr)^{-1}$, the corrected
square
\[
  q^2 - \sigma(w_q)
\]
has total degree at most $3$, so it lies in $\operatorname{im}(A_u)$ via the
same span relations. Summing these cubic image terms over an SOS decomposition
gives a full no-spurious-SOCP theorem for the whole
$\dim(W \cap \mathbb{R}[x]_{\le 1}) = 3$ branch without any exact
normalization.

The mixed-affine plane layer has now been strengthened one more step.
\texttt{RepresentativeMixedAffinePlane.lean} no longer stops at orthonormal
in-plane coefficient vectors for the nonsurjective rank-$14$ and rank-$13$
branches. It now proves the same residual-vanishing statements for orthogonal
but not necessarily unit column data, and it includes the matching
algebra-equivalence transport wrappers. So the affine-normalization step no
longer needs to rescale those relation vectors back to unit length before
pulling the theorem back to the original SOCP point. This removes one more
artifact of the earlier representative-first approach: the remaining obstacle
is now exactly to extract orthogonal mixed-affine plane data from the constant-
containing $\dim=2$ branch, not to renormalize it afterward.

The nonsurjective mixed-affine planes have now both been strengthened beyond
the orthogonal-column formulation. For rank $14$, the same file introduces a
determinant-based kernel for an arbitrary basis
\[
  x_2(a x_1+b x_2),\qquad x_2(c x_1+d x_2)
\]
of \(\operatorname{span}(x_1x_2,x_2^2)\) and formalizes the theorem
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_x1Plane\_det}. The
coefficient-space relation vectors are still assumed orthonormal, but the
polynomial columns now only need
\[
  ad-bc \neq 0.
\]
Lean reconstructs the exact \(x_1x_2\) and \(x_2^2\) relations by the inverse
\(2\times2\) matrix and then uses the syzygy built from
\[
  -x_1(c x_1+d x_2)\cdot x_2(a x_1+b x_2)
  +x_1(a x_1+b x_2)\cdot x_2(c x_1+d x_2),
\]
whose square still matches the missing \(x_1^4\) coefficient up to the factor
\(a^2+c^2\neq 0\).

For rank $13$, the file now also formalizes the exact analogue
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0Plane\_det}. Starting from
an arbitrary basis
\[
  x_1(a x_1+b x_2),\qquad x_1(c x_1+d x_2)
\]
of \(\operatorname{span}(x_1^2,x_1x_2)\), Lean reconstructs the exact
\(x_1^2\) and \(x_1x_2\) relations by the inverse matrix and uses the kernel
\[
  \mathrm{rel}_{-c_2}\bigl((c x_1+d x_2)(\beta+\gamma x_2)\bigr)
  +\mathrm{rel}_{c_3}\bigl((a x_1+b x_2)(\beta+\gamma x_2)\bigr),
\]
whose square matches the missing \(x_2^3\) and \(x_2^4\) coefficients up to
the common factor \(b^2+d^2\neq 0\). Both determinant theorems now also have
affine-equivalence transport wrappers. So the remaining mixed-affine
obstruction is no longer an in-plane algebra issue at all: it is the
normalization/classification step that has to extract one of the canonical
plane types together with nonzero determinant data.

That low-affine simplification needed one correction. The exploration script
already recorded the additional rank-\(14\) no-constant plane
\[
  (x_1,x_2,x_1^2-x_2^2,x_1x_2),
\]
so the old claim that there were only two \(\dim=2\) low-affine plane types
was too optimistic. The same file now contains basis-invariant plane theorems
for the coprime and common-factor models, and in this round it also gained the
exact-relation theorem
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_x0x1\_diffsq}
together with the determinant plane wrapper
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_x0x1\_diffsqPlane}.
So the definite plane
\[
  \operatorname{span}(x_1x_2,\ x_1^2-x_2^2)
\]
is now formalized directly as well. The same round also added the fixed
transport theorem
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_x0x1\_sumsqPlane}, so the
split rank-\(14\) plane
\[
  \operatorname{span}(x_1x_2,\ x_1^2+x_2^2)
\]
is now discharged through the explicit linear map
\((x_1,x_2)\mapsto(x_1+x_2,x_1-x_2)\). So the final normalization step no
longer needs exact low-affine representative equalities in the \(\dim=2\)
no-constant branch.

The linear low-affine branch has now advanced part of the same way. The file
\texttt{RepresentativeLowAffine.lean} now proves both the exact-relation
theorem and the basis-invariant homogeneous-basis theorem for the model
\((x_1,x_1^2,x_1x_2,x_2^2)\). The new common helper in
\texttt{RepresentativeSurjective.lean},
\texttt{exact\_relations\_of\_homQuadBasis\_det}, reconstructs the exact
canonical relations \(x_1^2,x_1x_2,x_2^2\) from any independent homogeneous
quadratic triple. So the linear branch no longer needs separate representative
normalization either: one linear relation plus a nondegenerate \(3\times 3\)
homogeneous relation matrix now suffices, and the result is also transported
across affine equivalences. The next cleanup now matches it to the same final
transport surface as the mixed-affine branches: the new theorem
\texttt{residual\_eq\_zero\_of\_socp\_of\_eq\_mix\_affineEquiv\_x0\_homQuadBasis\_det}
consumes exact normalized coordinates
\[
  \operatorname{mix}(M^\top,e^{-1}\!\cdot u)=(x_1,q_1,q_2,q_3)
\]
with a nondegenerate homogeneous coefficient matrix and closes the branch
directly.

The constant \(\dim=1\) branch has been lifted in parallel. The same new
\(3\times 3\) reconstruction now upgrades
\((1,x_1^2,x_1x_2,x_2^2)\) from a representative-only surjectivity theorem to
a basis-invariant homogeneous-basis theorem with affine transport as well. In
the same round I added the top transport wrappers
\texttt{residual\_eq\_zero\_of\_socp\_of\_eq\_mix\_mapVec\_const\_homQuadBasis\_det}
and
\texttt{residual\_eq\_zero\_of\_socp\_of\_eq\_mix\_affineEquiv\_const\_homQuadBasis\_det},
so exact normalized coordinates
\[
  \operatorname{mix}(M^\top,e^{-1}\!\cdot u)=(1,q_1,q_2,q_3)
\]
with a nondegenerate homogeneous coefficient matrix now close that branch
directly too.

I also matched the generic \(\dim=3\) branch to that same transport endpoint.
The new theorems
\texttt{residual\_eq\_zero\_of\_socp\_of\_eq\_mix\_mapVec\_contains\_aff1} and
\texttt{residual\_eq\_zero\_of\_socp\_of\_eq\_mix\_affineEquiv\_contains\_aff1}
mean that exact normalized coordinates
\[
  \operatorname{mix}(M^\top,e^{-1}\!\cdot u)=(1,x_1,x_2,q)
\]
now close the whole \(\dim(W \cap \mathbb{R}[x]_{\le 1})=3\) branch without
returning to relation-level packaging.

The surjective mixed-affine layer is broader now as well. The same file now
contains a fixed split-diagonal transport lemma for planes of the form
\[
  \operatorname{span}(x_1x_2,\ x_1^2+x_2^2).
\]
It uses the explicit linear change of variables
\[
  (x_1,x_2)\mapsto(x_1+x_2,\ x_1-x_2),
\]
under which
\[
  x_1x_2 \mapsto x_1^2-x_2^2,
  \qquad
  x_1^2+x_2^2 \mapsto 2(x_1^2+x_2^2).
\]
So any invertible basis of that split-diagonal plane transports to an
invertible basis of \(\operatorname{span}(x_1^2,x_2^2)\), and the opposite-sign
diagonal annihilator case no longer needs a separate ad hoc surjectivity
argument. Once the normalization reaches \(x_1^2-x_2^2\), Lean can now close
that branch by a fixed transport step.

This branch now has one more formalized bridge layer. In
\texttt{QuadraticCoordinateForm.lean}, every homogeneous quadratic with one of
the canonical coefficient constraints
\[
  \operatorname{coeff}_{x_1x_2}=0,\qquad
  \operatorname{coeff}_{x_1^2}=0,\qquad
  \operatorname{coeff}_{x_2^2}=0,\qquad
  \operatorname{coeff}_{x_1^2}\pm \operatorname{coeff}_{x_2^2}=0
\]
is rewritten into the corresponding canonical plane basis. Then
\texttt{RepresentativeMixedAffinePlane.lean} turns those coefficient
decompositions into generic residual-zero wrappers for the rank-$13$,
rank-$14$, and rank-$15$ mixed-affine plane theorems. So the remaining
mixed-affine gap is now purely the extraction step: produce the normalized
homogeneous quadratics and prove which coefficient constraint they satisfy.

One correction came out of the next Lean round. The theorem
\texttt{mixedAffineAnnihilator\_normal\_form\_cases} acts covariantly on the
polynomial annihilator \(a x_1^2 + b x_1x_2 + c x_2^2\), but the mixed-affine
classification problem actually lives on the dual coefficient functionals of
the extracted homogeneous quadratics. So the right normal form layer is the
dual action on
\[
  (\operatorname{coeff}_{x_1^2},\ \operatorname{coeff}_{x_1x_2},\
  \operatorname{coeff}_{x_2^2}),
\]
not the direct transport of the annihilator polynomial itself. This is now
formalized in \texttt{QuadraticCoordinateForm.lean}. The new theorem
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_const\_homQuadratics\_crossAnnihilator}
shears a pure-cross dual relation to the already-solved
\(\operatorname{coeff}_{x_1x_2}=0\) plane, and
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_const\_x0\_homQuadratics\_annihilator\_disc\_zero}
shears the discriminant-zero dual relation to the already-solved
\(\operatorname{coeff}_{x_1^2}=0\) rank-$14$ plane.
That wrapper has now been strengthened once more: it no longer needs the last
two coefficient-space relation vectors to be given in orthonormal form. A
nonzero Gram determinant suffices, and the file now carries out the necessary
Gram--Schmidt normalization internally before calling the existing rank-$14$
plane theorem.
The same file now also contains the diagonal scaling wrappers
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_const\_homQuadratics\_diagAnnihilator\_sum}
and
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_const\_homQuadratics\_diagAnnihilator\_diff},
so once the dual relation has already been cross-killed, the positive and
negative diagonal cases reduce directly to the sum-square and split-diagonal
plane theorems by one explicit scaling step.
That bridge is now composed one step further as well. The theorems
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_const\_homQuadratics\_annihilator\_sum}
and
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_const\_homQuadratics\_annihilator\_diff}
first shear a general dual annihilator
\(
  a\,\operatorname{coeff}_{x_1^2} +
  b\,\operatorname{coeff}_{x_1x_2} +
  c\,\operatorname{coeff}_{x_2^2}
\)
to kill the middle coefficient, then call the diagonal scaling wrappers. So the
mixed-affine normal-form algebra is now essentially theorem-complete: the
remaining gap is no longer the annihilator algebra itself.
\texttt{RepresentativeMixedAffinePlane.lean} now also proves
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_homQuadratics\_plane\_nontrivial},
which constructs the canonical annihilator directly from the \(2\times 3\)
homogeneous coefficient matrix of \((q_2,q_3)\) by coefficient cross-products
and then dispatches automatically to the cross, discriminant-zero, sum-square,
split-diagonal, or \(\operatorname{coeff}_{x_2^2}=0\) branch as appropriate.
The next focused probe tested whether the unresolved branch generically already
contains such a full homogeneous quadratic plane. The new script
\texttt{julia/mixed\_affine\_tail\_dim\_2026\_04\_10.jl} perturbs the canonical
rank-$13$, rank-$14$, and rank-$15$ mixed-affine homogeneous bases by random
affine tails while keeping \(\dim(W \cap \mathbb{R}[x]_{\le 1})=2\). In all
$20/20$ samples for each base, the result still had affine dimension $2$ but
homogeneous-quadratic intersection dimension only $1$:
\begin{verbatim}
mixed_affine_tail_dim_probe = (case = rank14_base, samples = 20, all values = (2,1))
mixed_affine_tail_dim_probe = (case = rank13_base, samples = 20, all values = (2,1))
mixed_affine_tail_dim_probe = (case = rank15_base, samples = 20, all values = (2,1))
\end{verbatim}
So the remaining gap is now more specific. The generic mixed-affine branch
should be attacked as ``one homogeneous line plus one tailed quadratic,'' not
as ``a full homogeneous quadratic plane already sitting inside \(W\).'' To
support that bridge, \texttt{QuadraticCoordinateForm.lean} was extended in this
round with the missing explicit formulas for the constant and \(x_1\)
coefficients under the translation \(x_2 \mapsto x_2+t\).
The next Lean step sharpened that further: it now defines the canonical
homogeneous extractor
\[
  H(q_2,q_3)
  :=
  \operatorname{coeff}_{x_2}(q_3)\,q_2
  -
  \operatorname{coeff}_{x_2}(q_2)\,q_3,
\]
and proves that whenever \(q_2\) and \(q_3\) already have zero constant and
\(x_1\) coefficients, \(H(q_2,q_3)\) is automatically homogeneous quadratic.
Its three homogeneous coefficients are the obvious \(2\times2\) minors
\[
  \operatorname{coeff}_{x_2}(q_3)\operatorname{coeff}_{x_1^2}(q_2) -
  \operatorname{coeff}_{x_2}(q_2)\operatorname{coeff}_{x_1^2}(q_3),
\]
and similarly for \(x_1x_2\) and \(x_2^2\). So the remaining bridge is now even
more concrete: classify the companion tailed quadratic relative to this
explicit homogeneous line, rather than searching abstractly for a hidden
homogeneous subspace.
The next experiment then checked the strongest easy sub-branch of that bridge.
The new script
\texttt{julia/mixed\_affine\_common\_translation\_2026\_04\_10.jl} samples
random tailed quadratic pairs
\[
  q_i=a_i x_1^2+b_i x_1x_2+c_i x_2^2+e_i x_2
\]
with
\(
  b_2c_3-c_2b_3 \neq 0
\),
solves the common-translation system
\[
  e_i + u\,b_i + 2v\,c_i = 0,
\]
and measures the residual tail coefficients after
\((x_1,x_2)\mapsto(x_1+u,x_2+v)\). The output was
\begin{verbatim}
mixed_affine_common_translation_probe = (samples = 20,
  max_tail_error = 1.4210854715202004e-14,
  min_abs_det = 0.02515974405263005, ...)
\end{verbatim}
so the explicit common-translation formula is numerically stable at machine
precision on random nondegenerate samples. This sharpens the next Lean target:
the determinant-nondegenerate mixed-affine sub-branch should be reduced by a
formal common-translation lemma matching that explicit formula, after which the
existing homogeneous-plane theorem can take over.
That Lean reduction now exists in its minimal algebraic form. In
\texttt{QuadraticCoordinateForm.lean} I added the symmetric
\(x_1 \mapsto x_1+t\) translation API, proved the exact transformed
\(\operatorname{coeff}_{x_2}\) and \(\operatorname{coeff}_{x_2^2}\) formulas,
and then formalized the nested tail-killing theorem for the Cramér-rule choice
of \((u,v)\). The next Lean refinement then added the corresponding transformed
\(\operatorname{coeff}_{1}\) and \(\operatorname{coeff}_{x_1}\) formulas for the
nested translation as well. That integration step is now formalized too:
\texttt{RepresentativeMixedAffinePlane.lean} contains the new theorem
\texttt{residual\_eq\_zero\_of\_const\_x0\_tailedPair\_det}, which starts from
exact coordinates \(u=(1,x_1,q_2,q_3)\), applies the explicit common
translation, removes the induced \(1\)- and \(x_1\)-corrections, proves the
resulting Gram determinant is automatically nonzero, and then dispatches to the
existing homogeneous-plane theorem
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_homQuadratics\_plane\_nontrivial}.
The same file now also exports the affine-equivalence wrapper
\texttt{residual\_eq\_zero\_of\_equiv\_const\_x0\_tailedPair\_det}, so this
determinant-nondegenerate tail certificate is available directly after affine
normalization. In the next Lean step I also matched it to the standard final
assembly interface by adding
\texttt{residual\_eq\_zero\_of\_socp\_of\_eq\_mix\_affineEquiv\_const\_x0\_tailedPair\_det},
which consumes exactly the same kind of affine-normalization and orthogonal
factor-mixing data as the existing representative transport layer.
The next cleanup closed one more practical gap. The same file now proves
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_tailedPair\_det} and
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_const\_x0\_tailedPair\_det}.
Those theorems take orthonormal coefficient vectors
\(c_0,c_1,c_2,c_3 \in \mathbb{R}^4\) with explicit relations to
\(1,x_1,q_2,q_3\), build the corresponding orthogonal factor-mixing matrix
internally, and then dispatch to the transport-shaped tailed-pair certificate.
So the determinant-nondegenerate branch is no longer blocked on translation
algebra, on identifying the affine correction terms, or on integrating those
corrections with the plane certificate layer. It is now blocked upstream on
extracting the exact normalized coordinate picture \(u=(1,x_1,q_2,q_3)\), or
equivalently an orthonormal relation basis for that picture, from a general
constant-containing mixed-affine factor.
The next Lean cleanup removed the same interface mismatch for the homogeneous
mixed-affine subcase itself. \texttt{RepresentativeMixedAffinePlane.lean} now
contains
\texttt{residual\_eq\_zero\_of\_equiv\_const\_x0\_homQuadratics\_plane\_nontrivial}
and
\texttt{residual\_eq\_zero\_of\_socp\_of\_eq\_mix\_affineEquiv\_const\_x0\_homQuadratics\_plane\_nontrivial}.
So both constant-containing mixed-affine outputs are now available in the same
exact-coordinate transport form: after affine normalization and orthogonal
factor mixing, the remaining classification theorem can hand back
\((1,x_1,q_2,q_3)\) either with a tailed-pair determinant condition or with a
nontrivial homogeneous quadratic plane, and the top-level transport layer
handles either branch directly.
The next probe targeted the determinant-zero side directly. The new script
\texttt{julia/mixed\_affine\_det\_zero\_2026\_04\_10.jl} uses only coefficient
linear algebra, so it runs in the current workspace without the unavailable SDP
stack. It checks the canonical families
\[
  (1,x_1,\mu x_1^2 + D,\nu x_1^2 + x_2)
\]
for
\[
  D \in \{x_1x_2,\ x_2^2,\ x_1^2+x_2^2,\ x_1^2-x_2^2\}
\]
and \((\mu,\nu) \in
\{(-2,-1),(-1,1),(0,1),(1,1),(2,-1)\}\). The output is completely stable:
\begin{verbatim}
mixed_affine_det_zero_canonical_probe = (direction = "cross", ..., rank = 14, missing = [(0, 4)])
mixed_affine_det_zero_canonical_probe = (direction = "x1sq", ..., rank = 15, missing = [])
mixed_affine_det_zero_canonical_probe = (direction = "sumsq", ..., rank = 15, missing = [])
mixed_affine_det_zero_canonical_probe = (direction = "diffsq", ..., rank = 15, missing = [])
\end{verbatim}
for every sampled parameter pair. So the determinant-zero tailed branch now
looks much less amorphous than before: numerically the cross direction behaves
like the known rank-\(14\) mixed-affine obstruction with only the \(x_2^4\)
class missing, while the \(x_2^2\) and diagonal directions are already
surjective.
I then added the coefficient-only follow-up script
\texttt{julia/mixed\_affine\_det\_zero\_slice\_2026\_04\_10.jl}. It samples the
same tailed determinant-zero normal form with a generic homogeneous line
\[
  D = a x_1^2 + b x_1x_2 + c x_2^2
\]
and records the image ranks on several natural slices. The output is
\begin{verbatim}
mixed_affine_det_zero_slice_probe =
  (samples = 80, generic_ranks = [15], c0_ranks = [14],
   b0_ranks = [15], disc0_ranks = [15])
\end{verbatim}
So the rank-\(14\) behaviour now looks genuinely codimension one on the
determinant-zero side: generic lines are rank \(15\), the slice \(c=0\) with
nonzero cross term stays rank \(14\), and the \(b=0\) or discriminant-zero
slices are already rank \(15\). That makes the earlier canonical picture much
more believable: after the standard basis choice, only the pure cross-type
determinant-zero family appears to require a nonsurjective theorem.
I then added a small canonical tail-parameter sweep,
\texttt{julia/mixed\_affine\_det\_zero\_canonical\_a\_sweep\_2026\_04\_10.jl},
to check whether that picture depends on the choice of tail coefficient.
Its output is
\begin{verbatim}
mixed_affine_det_zero_canonical_a_sweep = [
  (family = "cross",  a = -1.0, rank = 14, missing = [(0, 4)]),
  (family = "cross",  a =  1.0, rank = 14, missing = [(0, 4)]),
  (family = "cross",  a =  2.0, rank = 14, missing = [(0, 4)]),
  (family = "x1sq",   a = -1.0, rank = 15, missing = []),
  (family = "x1sq",   a =  1.0, rank = 15, missing = []),
  (family = "x1sq",   a =  2.0, rank = 15, missing = []),
  (family = "sumsq",  a = -1.0, rank = 15, missing = []),
  (family = "sumsq",  a =  1.0, rank = 15, missing = []),
  (family = "sumsq",  a =  2.0, rank = 15, missing = []),
  (family = "diffsq", a = -1.0, rank = 15, missing = []),
  (family = "diffsq", a =  1.0, rank = 15, missing = []),
  (family = "diffsq", a =  2.0, rank = 15, missing = [])
]
\end{verbatim}
So the canonical determinant-zero picture is not just stable across one or two
ad hoc examples. The cross family keeps the same codimension-one obstruction
for all tested tails, while the \(x_2^2\), sum-square, and split-diagonal
families stay surjective for all tested tails.
The next Lean step removed one easy but real part of that determinant-zero
picture. In \texttt{RepresentativeMixedAffinePlane.lean} I added the coefficient
identity \texttt{relation\_x1\_of\_affineTail} and the SOCP consequences
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_affineTail} and
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_const\_x0\_affineTail}. These
show that if a coordinate relation already has the form
\[
  q = a + b x_1 + c x_2
\]
with \(b \neq 0\), then \(x_2\) is recoverable by an explicit linear
combination of the \(1\), \(x_1\), and \(q\) relations, so the whole branch
falls back to the existing \(\dim(W \cap \mathrm{Aff}_1)=3\) theorem. So the
determinant-zero mixed-affine branch is no longer blocked on the genuinely
affine subcase; only the case with a nonzero residual homogeneous quadratic
part in the tailed coordinate remains.

The next Lean step removed one full representative family from that residual
determinant-zero side. In
\texttt{RepresentativeMixedAffine.lean} I added the representative
\[
  (1,x_1,x_2^2,x_2 + a x_1^2), \qquad a \neq 0,
\]
as \texttt{mixedAffineTailX1SqRep a}, together with the direct image theorem
\texttt{quartic\_in\_image\_mixedAffineTailX1SqRep} and the SOCP consequence
\texttt{residual\_eq\_zero\_mixedAffineTailX1SqRep}. The proof is a short
explicit monomial generation argument: every quartic monomial is already
immediate except \(x_1^4\) and \(x_1^3x_2\), and each of those comes from the
tailed coordinate after one correction term using the \(x_1\) or \(x_2^2\)
coordinate. So, combined with the affine-tail theorem for the \(a=0\) case,
the entire \(x_2^2\)-type determinant-zero tailed family is now formalized at
representative level. I then lifted that result to exact relation data in the
same file via
\texttt{quartic\_in\_image\_of\_relations\_const\_x0\_x1sq\_x1PlusX0sq},
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_x1sq\_x1PlusX0sq}, and
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_const\_x0\_x1sq\_x1PlusX0sq}.
So the \(x_2^2\)-type determinant-zero slice is no longer only a concrete
representative theorem; it is now usable immediately after affine
normalization, without rebuilding the monomial image argument in the final
assembly.

I also tried to jump directly to a general diagonal determinant-zero Lean
theorem for \((1,x_1,x_2^2 + c x_1^2, x_2 + a x_1^2)\), but that broad version
was not stable enough to keep: the proof expanded into a large explicit
monomial calculation with several heartbeat-heavy normalization steps. I backed
that attempt out rather than leaving the theorem file dirty. So the strategy is
now sharper again: use the new tail-parameter sweep as evidence that the next
Lean target should be the cross-type branch or a smaller exact diagonal
representative, not another over-general unfinished statement.

I then pinned that diagonal evidence down in a dedicated script,
\texttt{julia/mixed\_affine\_det\_zero\_diag\_tail\_2026\_04\_10.jl}, for the
two-parameter family
\[
  (1,x_1,d x_1^2 + x_2^2,\ x_2 + a x_1^2),
  \qquad
  d \in \{-2,-1,0,1,2\},\ 
  a \in \{-2,-1,1,2\}.
\]
All \(20\) sampled points have image rank \(15\) and no missing quartic
monomials. So the failed direct Lean generalization was not pointing at a bad
numerical guess; it was only a proof-shape problem. Numerically, the
determinant-zero mixed-affine branch now looks even more rigid: once the
homogeneous line carries a nonzero \(x_2^2\)-coefficient in the canonical
basis, the tailed family appears uniformly surjective, and the only stable
nonsurjective slice still visible is the cross-type \(x_1 x_2\) direction.

The next Lean round then did exactly that. In
\texttt{RepresentativeMixedAffine.lean} I added the cross-type determinant-zero
representative
\[
  (1,x_1,x_1x_2,x_2 + a x_1^2), \qquad a \neq 0,
\]
as \texttt{mixedAffineTailCrossRep a}. The new theorems
\texttt{quartic\_in\_image\_mixedAffineTailCrossRep\_of\_coeff\_m04\_zero},
\texttt{mixedAffineTailCrossKerBase}, and
\texttt{residual\_eq\_zero\_mixedAffineTailCrossRep}
show that this branch is a direct rank-\(14\) image-plus-kernel certificate:
every quartic with zero \(x_2^4\) coefficient lies in the image, while the
explicit kernel direction contributes exactly the missing \(x_2^4\) class.
The next Lean round then lifted that cross slice through the orthonormal
relation transport layer in \texttt{RepresentativeMixedAffinePlane.lean}. The
new theorems
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_x0x1\_x1PlusX0sq\_orthonormal}
and
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_const\_x0\_x0x1\_x1PlusX0sq\_orthonormal}
show that the cross-type determinant-zero family is no longer trapped at the
bare representative level: once affine normalization produces an orthonormal
relation basis with polynomial images \(1\), \(x_1\), \(x_1x_2\), and
\(x_2 + a x_1^2\), the SOCP conclusion follows directly. So the canonical
determinant-zero mixed-affine branch is now theorem-complete at the
representative-or-orthonormal-relation level; the remaining issue is no longer
``find the missing cross theorem,'' but only ``extract exact normalized cross
data from a general admissible factor.''

The next Lean round removed even that orthonormality bottleneck for the
determinant-zero cross slice. In \texttt{RepresentativeMixedAffine.lean} the
new theorems
\texttt{quartic\_in\_image\_of\_relations\_const\_x0\_x0x1\_x1PlusX0sq\_of\_coeff\_m04\_zero},
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_x0x1\_x1PlusX0sq}, and
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_const\_x0\_x0x1\_x1PlusX0sq}
lift the same branch to plain exact relation data. The proof keeps the same
rank-\(14\) shape as the concrete representative, but the kernel direction is
now built directly from the relation vectors, and the missing
\(\operatorname{coeff}_{x_2^4}\) depends only on the constant relation column.
So the usable cross-type theorem layer no longer needs an orthonormal relation
basis at all. The remaining mixed-affine extraction gap is therefore smaller
again: it is now genuinely ``produce exact normalized determinant-zero cross
relations after affine normalization,'' not ``orthonormalize them first.''

The next Lean round tightened that cross branch once more at the coefficient
level. In \texttt{RepresentativeMixedAffinePlane.lean} the new theorems
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_crossTail} and
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_const\_x0\_crossTail} start
from exact relations to \(1\), \(x_1\), a pure cross quadratic \(q_2\), and a
second quadratic \(q_3\) whose \(x_2^2\)-coefficient vanishes but whose
\(x_2\)- and \(x_1^2\)-coefficients are both nonzero. The proof then performs
the remaining normalization internally: it rescales \(q_2\) to \(x_1x_2\),
subtracts the cross component from \(q_3\), rescales the remainder to
\(x_2 + a x_1^2\), and invokes the exact determinant-zero cross theorem from
\texttt{RepresentativeMixedAffine.lean}. So the cross-type determinant-zero
slice is no longer waiting for the classification step to output fully
normalized polynomial images; a coefficient-level pure-cross-plus-tail
description now already closes the SOCP branch.

The next Lean round then added the symmetric coefficient-level wrapper for the
\(x_2^2\)-type determinant-zero slice. The new theorems
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_x1sqTail} and
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_const\_x0\_x1sqTail} start
from exact relations to \(1\), \(x_1\), a pure \(x_2^2\) homogeneous line, and
a second quadratic whose \(x_1x_2\)-coefficient vanishes but whose \(x_2\)- and
\(x_1^2\)-coefficients are both nonzero. The proof rescales the \(x_2^2\) line,
subtracts the \(x_2^2\)-component from the tailed quadratic, rescales the
remainder to \(x_2 + a x_1^2\), and then invokes the already-formalized exact
\(x_2^2\)-tail theorem from \texttt{RepresentativeMixedAffine.lean}. So both
nonsurjective determinant-zero mixed-affine slices now admit coefficient-level
wrappers above the exact normalized relation layer.

The next Lean round finally turned the simplest diagonal determinant-zero slice
from a numerical expectation into a theorem. In
\texttt{RepresentativeMixedAffine.lean} I added the exact diagonal theorems
\texttt{quartic\_in\_image\_of\_relations\_const\_x0\_diagX1sq\_x1PlusX0sq},
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_diagX1sq\_x1PlusX0sq},
and
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_const\_x0\_diagX1sq\_x1PlusX0sq}.
These prove directly that exact relations to
\[
  (1,\;x_1,\; d x_1^2 + x_2^2,\; x_2 + a x_1^2), \qquad a \neq 0,
\]
already force zero residual. The proof is by direct monomial image generation:
after the easy degree-\(\le 3\) cases, the quartic monomials are produced in
the chain \(x_1^4\), \(x_1^3x_2\), \(x_1^2x_2^2\), \(x_1x_2^3\), \(x_2^4\),
each corrected using only the diagonal homogeneous line and the tailed
coordinate. I then lifted that exact theorem one level higher in
\texttt{RepresentativeMixedAffinePlane.lean} via
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_diagX1sqTail} and
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_const\_x0\_diagX1sqTail},
which start only from the coefficient conditions on a diagonal homogeneous
line and a compatible tail. So the diagonal determinant-zero slice is no longer
just a numerical observation; it is now part of the checked theorem layer.

The next Lean round removed one more piece of explicit normalization from that
same determinant-zero branch. In
\texttt{RepresentativeMixedAffinePlane.lean} I added the new theorems
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_mixedAffineTailHomLine\_crossTail},
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_mixedAffineTailHomLine\_x1sqTail},
and
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_mixedAffineTailHomLine\_diagX1sqTail},
plus their affine-equivalence wrappers. These start only from exact relations
to \(1\), \(x_1\), and a tailed quadratic pair \(q_2,q_3\) with vanishing
constant and \(x_1\)-linear coefficients. Lean now forms the canonical
homogeneous extractor
\[
  H(q_2,q_3)
  =
  \operatorname{coeff}_{x_2}(q_3)\,q_2
  -
  \operatorname{coeff}_{x_2}(q_2)\,q_3,
\]
proves the resulting relation internally, and then dispatches to the existing
cross, \(x_2^2\)-type, or diagonal determinant-zero wrapper depending on the
coefficient pattern of \(H(q_2,q_3)\). So the determinant-zero mixed-affine
classification no longer has to output a separate explicit homogeneous-line
relation in those three cases; the line is now extracted inside Lean from the
tailed pair itself.

I then weakened the diagonal slice one more step. The new theorems
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_diagX1sqTail\_det} and
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_const\_x0\_diagX1sqTail\_det}
remove the old hypothesis that the tailed quadratic already have zero
\(x_2^2\)-coefficient. Instead, Lean now subtracts the explicit multiple of the
diagonal homogeneous line
\[
  t = \frac{\operatorname{coeff}_{x_2^2}(q_3)}
           {\operatorname{coeff}_{x_2^2}(q_2)}
\]
internally, and the only extra condition is the nonvanishing determinant
\[
  \operatorname{coeff}_{x_2^2}(q_2)\operatorname{coeff}_{x_1^2}(q_3)
  -
  \operatorname{coeff}_{x_1^2}(q_2)\operatorname{coeff}_{x_2^2}(q_3) \neq 0,
\]
which guarantees that the corrected tail still has nonzero \(x_1^2\)-part. So
even the diagonal determinant-zero branch is now one normalization step closer
to the natural coefficient data expected from the remaining classification.

The next Lean round lifted the same weakening through the extracted-line
interface. In \texttt{RepresentativeMixedAffinePlane.lean} I added
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_mixedAffineTailHomLine\_crossTail\_det},
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_mixedAffineTailHomLine\_x1sqTail\_det},
and
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_mixedAffineTailHomLine\_diagX1sqTail\_det},
with the matching equivariant wrappers. These start from the canonical
extracted homogeneous line \(H(q_2,q_3)\) and a tailed quadratic \(q_3\), but
they no longer require the classification step to separately prove the residual
tail-coefficient vanishing in the determinant-zero branch. In the cross case,
the determinant-zero identity
\[
  \operatorname{coeff}_{x_1x_2}(H)\operatorname{coeff}_{x_2^2}(q_3)
  -
  \operatorname{coeff}_{x_2^2}(H)\operatorname{coeff}_{x_1x_2}(q_3)=0
\]
reduces immediately to \(\operatorname{coeff}_{x_2^2}(q_3)=0\) because
\(\operatorname{coeff}_{x_2^2}(H)=0\) and
\(\operatorname{coeff}_{x_1x_2}(H)\neq 0\). The \(x_2^2\)-type case is the
same with roles of \(x_1x_2\) and \(x_2^2\) reversed. In the diagonal case,
that same determinant-zero identity now automatically gives
\(\operatorname{coeff}_{x_1x_2}(q_3)=0\), and the second diagonal determinant
is passed straight into the already-formalized diagonal det wrapper. So the
remaining mixed-affine extraction step now only has to identify the coefficient
type of \(H(q_2,q_3)\) and the relevant determinant data, not the leftover
tail cancellations themselves.

The next cleanup removed one more artificial conversion from that same branch.
I added the coefficient identities
\texttt{det\_m11\_m02\_mixedAffineTailHomLine} and
\texttt{det\_m02\_m20\_mixedAffineTailHomLine} in
\texttt{QuadraticCoordinateForm.lean}. They express the extracted-line
determinants as the original tailed-pair determinants multiplied by the tail
coefficient \(\operatorname{coeff}_{x_2}(q_3)\). With those in place,
\texttt{RepresentativeMixedAffinePlane.lean} now proves the stronger wrappers
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_mixedAffineTailHomLine\_crossTail\_origDetZero},
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_mixedAffineTailHomLine\_x1sqTail\_origDetZero},
and
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_mixedAffineTailHomLine\_diagX1sqTail\_origDetZero},
again with equivariant wrappers. So the determinant-zero mixed-affine
classification no longer has to hand-convert the original pair determinant
\[
  \operatorname{coeff}_{x_1x_2}(q_2)\operatorname{coeff}_{x_2^2}(q_3)
  -
  \operatorname{coeff}_{x_2^2}(q_2)\operatorname{coeff}_{x_1x_2}(q_3)=0
\]
into the corresponding determinant identity for the extracted line
\(H(q_2,q_3)\); Lean now performs that conversion internally.

The next Lean round then compressed the diagonal determinant-zero slice itself.
In \texttt{RepresentativeMixedAffinePlane.lean} I added
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_mixedAffineTailHomLine\_diagX1sqTail\_origDetZero\_cases}
and its equivariant wrapper. These no longer require the second diagonal
determinant
\[
  \operatorname{coeff}_{x_2^2}(q_2)\operatorname{coeff}_{x_1^2}(q_3)
  -
  \operatorname{coeff}_{x_1^2}(q_2)\operatorname{coeff}_{x_2^2}(q_3)
\]
as an input hypothesis. Lean now splits internally:
\begin{itemize}
\item if that determinant is nonzero, it calls the already-formalized diagonal
tail theorem;
\item if it vanishes, it subtracts the corresponding multiple of the extracted
homogeneous line \(H(q_2,q_3)\) from \(q_3\), checks that the quadratic part
disappears completely, and falls back to
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_affineTail}.
\end{itemize}
So even the diagonal determinant-zero extracted-line branch is now expressed in
terms of the natural original pair data plus an internal Lean case split,
rather than another external subclassification.

The next Lean round showed that the same cleanup applies to the other two
determinant-zero extracted-line slices. I added
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_mixedAffineTailHomLine\_crossTail\_origDetZero\_cases}
and
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_mixedAffineTailHomLine\_x1sqTail\_origDetZero\_cases},
again with equivariant wrappers. In both cases Lean now splits on whether the
remaining \(x_1^2\)-tail coefficient vanishes:
\begin{itemize}
\item if it is nonzero, the branch dispatches to the already-formalized cross
or \(x_2^2\)-type determinant-zero theorem;
\item if it vanishes, Lean subtracts the corresponding multiple of the
extracted homogeneous line and reduces the coordinate to an affine tail, so the
branch falls back to
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_affineTail}.
\end{itemize}
So the determinant-zero mixed-affine extracted-line layer is now uniform across
all three coefficient types: the external classification only has to identify
the extracted-line type and the original determinant-zero relation, while the
remaining affine-vs-quadratic tail split is handled internally in Lean.

The next Lean round lifted those three determinant-zero extracted-line slices
to the top transport interface. For the cross, \(x_2^2\)-type, and diagonal
extracted-line branches, \texttt{RepresentativeMixedAffinePlane.lean} now has
coordinate-level equivariant theorems and matching
\texttt{residual\_eq\_zero\_of\_socp\_of\_eq\_mix\_affineEquiv\_\dots} wrappers.
So the remaining mixed-affine classification no longer needs to output
relation-level data on the transformed point. It can work at the cleaner
coordinate level
\[
  \operatorname{mix}(M^\top, e^{-1}\!\cdot u) = (1,x_0,q_2,q_3),
\]
plus the extracted-line coefficient pattern and the original determinant-zero
relation, and then Lean handles the residual-zero conclusion from there.

The next Lean round weakened the cross-type determinant-zero slice one step
further. I added
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_mixedAffineTailHomLine\_crossShear\_origDetZero\_cases},
together with its equivariant wrapper and the top transport theorem
\texttt{residual\_eq\_zero\_of\_socp\_of\_eq\_mix\_affineEquiv\_const\_x0\_mixedAffineTailHomLine\_crossShear\_origDetZero\_cases}.
This means the external classification no longer has to hand back a line
already normalized to the pure-cross form. It is now enough to produce
coordinates \((1,x_0,q_2,q_3)\) with zero constant and \(x_0\)-linear parts in
\(q_2,q_3\), original determinant-zero relation for the pair, nonzero
\(x_1\)-tail in \(q_3\), and extracted line \(H(q_2,q_3)\) satisfying only
\[
  \operatorname{coeff}_{x_1^2} H(q_2,q_3)=0,\qquad
  \operatorname{coeff}_{x_0x_1} H(q_2,q_3)\neq 0.
\]
Lean now performs the remaining \(x_1\)-shear internally, subtracts the
induced \(x_0\)-tail from the transformed pair, and falls through to the
already-formalized pure-cross determinant-zero theorem. So the genuinely new
content of this round is that the last cross-type shear normalization has been
absorbed into the theorem interface itself rather than left as an external
classification obligation.

The next Lean round did the same for the complementary determinant-zero side.
I added
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_mixedAffineTailHomLine\_diagShear\_origDetZero\_cases},
its equivariant wrapper, and the corresponding top transport theorem. So the
external classification no longer has to hand back a homogeneous line already
sheared into pure \(x_1^2\)-type or diagonal form. It is now enough to produce
coordinates \((1,x_0,q_2,q_3)\) with zero constant and \(x_0\)-linear parts in
\(q_2,q_3\), the original determinant-zero relation for the pair, a nonzero
\(x_1\)-tail in \(q_3\), and only the coarse condition
\[
  \operatorname{coeff}_{x_1^2} H(q_2,q_3)\neq 0,
  \qquad H(q_2,q_3)=\texttt{mixedAffineTailHomLine}(q_2,q_3).
\]
Lean now performs the remaining \(x_1\)-shear internally, subtracts the
induced \(x_0\)-tails, re-extracts the homogeneous line, and falls through to
the already-formalized diagonal/\(x_1^2\)-type determinant-zero theorem. So
the determinant-zero extracted-line layer is now genuinely two-way at theorem
level for the two theorem families already formalized, but one residual
pure-\(x_0^2\) slice still has to be classified separately. To check whether
that residual slice is a real new obstruction, I added the script
\texttt{julia/mixed\_affine\_det\_zero\_x0sq\_tail\_2026\_04\_10.jl}, which
probes the family
\[
  q_2=b_2x_1+a_2x_0^2+c x_0x_1+d x_1^2,\qquad
  q_3=b_3x_1+a_3x_0^2+\frac{b_3}{b_2}c x_0x_1+\frac{b_3}{b_2}d x_1^2,
\]
so that \(H(q_2,q_3)\) is always pure \(x_0^2\). It prints
\[
  \texttt{mixed\_affine\_det\_zero\_x0sq\_tail\_probe}
  =
  (\texttt{samples}=20,\ \texttt{generic\_ranks}=[15],\
   \texttt{low\_affine\_ranks}=[13],\
   \texttt{low\_affine\_missing}=[(1,3),(0,4)]).
\]
So the pure-\(x_0^2\) extracted-line family is not uniformly rank \(13\): the
generic slice looks surjective, while the special low-affine slice
\(c=d=0\) still matches the known rank-\(13\) behavior.

The next Lean round internalized the generic side of that picture. In
\texttt{RepresentativeMixedAffinePlane.lean} I added theorems
\texttt{residual\_eq\_zero\_of\_const\_x0\_x0sqTail\_m02\_nonzero},
\texttt{residual\_eq\_zero\_of\_equiv\_const\_x0\_x0sqTail\_m02\_nonzero}, and
their transport wrappers. They cover exact normalized coordinates
\((1,x_0,q_2,q_3)\) where \(q_2\) is pure \(x_0^2\) and
\(\operatorname{coeff}_{x_1^2}(q_3)\neq 0\). Lean now performs the same
\(x_1\)-translation suggested by the numerical probe, kills the \(x_1\)-tail
of \(q_3\), subtracts the induced constant and \(x_0\) pieces by explicit
coordinate relations, and reduces automatically to the existing
homogeneous-plane theorem. So the pure-\(x_0^2\) determinant-zero branch is no
longer open generically; the remaining pure-\(x_0^2\) difficulty is now the
complementary \(\operatorname{coeff}_{x_1^2}(q_3)=0\) slice, which is exactly
where the low-affine rank-\(13\) behavior lives numerically.

The next Lean round removed another generic part of that complementary slice.
In the same file I added
\texttt{residual\_eq\_zero\_of\_const\_x0\_x0sqTail\_m11\_nonzero},
\texttt{residual\_eq\_zero\_of\_equiv\_const\_x0\_x0sqTail\_m11\_nonzero}, and
their transport wrappers. These cover exact normalized coordinates
\((1,x_0,q_2,q_3)\) with \(q_2\) pure \(x_0^2\),
\(\operatorname{coeff}_{x_1^2}(q_3)=0\), but
\(\operatorname{coeff}_{x_0x_1}(q_3)\neq 0\). Lean now performs the explicit
\(x_0\)-translation killing the \(x_1\)-tail of \(q_3\), subtracts the induced
constant and \(x_0\) pieces from both coordinates, and reduces automatically to
the existing homogeneous-plane theorem. So the only remaining pure-\(x_0^2\)
determinant-zero slice is the special low-affine-looking one with both
\(\operatorname{coeff}_{x_1^2}(q_3)=0\) and
\(\operatorname{coeff}_{x_0x_1}(q_3)=0\).

The next Lean round removed even that remaining exact-coordinate slice. I added
\texttt{residual\_eq\_zero\_of\_const\_x0\_x0sqTail\_m11\_m02\_zero} together
with the coarser wrapper
\texttt{residual\_eq\_zero\_of\_const\_x0\_x0sqTail\_m02\_zero} and their
transport versions. So if exact normalized coordinates
\((1,x_0,q_2,q_3)\) are available with \(q_2\) a pure \(x_0^2\)-type line and
\(\operatorname{coeff}_{x_1^2}(q_3)=0\), Lean now handles the whole branch
internally: if \(\operatorname{coeff}_{x_0x_1}(q_3)\neq 0\), it invokes the
translated homogeneous-plane theorem; if
\(\operatorname{coeff}_{x_0x_1}(q_3)=0\), it falls either to the affine-tail
case or to an explicit constant relation. So the pure-\(x_0^2\)
determinant-zero family is no longer blocked by one more ad hoc representative
argument; what remains is the upstream extraction/normalization statement.

The next Lean round lifts that same pure-\(x_0^2\) slice to the extracted-line
transport surface. I added
\texttt{residual\_eq\_zero\_of\_const\_x0\_mixedAffineTailHomLine\_x0sqTail\_cases}
and its equivariant/\(\operatorname{mix}(M^\top,e^{-1}\!\cdot u)\)-transport
wrappers. So the classification layer no longer has to produce a specially
chosen coordinate \(q_2\) which is already the pure \(x_0^2\) line. It is now
enough to produce normalized coordinates \((1,x_0,q_2,q_3)\) such that the
extracted homogeneous line
\[
  H(q_2,q_3)=\texttt{mixedAffineTailHomLine}(q_2,q_3)
\]
is pure \(x_0^2\) with nonzero coefficient. Lean then performs an internal
orthogonal \(2\times 2\) rotation of the last two factor coordinates, turning
\((q_2,q_3)\) into \((q_H,q_P)\) with \(q_H\) exactly the normalized pure
\(x_0^2\) line and dispatching automatically to the already-formalized
\texttt{x0sqTail\_m02\_nonzero} or \texttt{x0sqTail\_m02\_zero} theorem
according to \(\operatorname{coeff}_{x_1^2}(q_P)\).

The next Lean round compresses the determinant-zero extracted-line branch one
step further. I added
\texttt{residual\_eq\_zero\_of\_const\_x0\_mixedAffineTailHomLine\_origDetZero\_cases}
and its equivariant/\(\operatorname{mix}(M^\top,e^{-1}\!\cdot u)\)-transport
wrappers. So the classification layer no longer has to pre-classify the
determinant-zero branch into cross, \(x_2^2\)-type, diagonal, or pure
\(x_0^2\) outputs. It is now enough to produce exact normalized coordinates
\((1,x_0,q_2,q_3)\) with zero constant/\(x_0\) coefficients, the original
determinant-zero relation for \((q_2,q_3)\), and a nonzero \(x_1\)-tail in
\(q_3\). Lean then forms the extracted line
\[
  H(q_2,q_3)=\texttt{mixedAffineTailHomLine}(q_2,q_3)
\]
internally and performs the whole coarse split itself: if \(H=0\), it closes
by a direct constant relation; if \(H\) is cross-like, diagonal/\(x_2^2\)-like,
or pure \(x_0^2\), it dispatches to the already-formalized theorem for that
case. So the determinant-zero mixed-affine side now really asks only for one
exact normalized tailed pair, not for a separately normalized extracted-line
output.

The next Lean round compresses the branch again, now at the level of the whole
constant-containing mixed-affine exact coordinate picture. I added
\texttt{residual\_eq\_zero\_of\_const\_x0\_pair\_coeff\_m00\_m10\_zero} and
its equivariant/\(\operatorname{mix}(M^\top,e^{-1}\!\cdot u)\)-transport
wrappers. So once exact normalized coordinates
\((1,x_0,q_2,q_3)\) are available with \(q_2,q_3\) quadratic and with zero
constant and \(x_0\) coefficients, the classification layer no longer has to
hand back determinant data, extracted-line data, or a preferred choice of
which tail carries the nonzero \(x_1\)-coefficient. Lean now performs the
entire internal split on its own: determinant-nondegenerate pairs fall to the
tailed-pair theorem; determinant-zero pairs with a nonzero \(x_1\)-tail in one
coordinate fall to the coarse determinant-zero theorem after swapping if
needed; purely homogeneous nontrivial pairs fall to the homogeneous-plane
theorem; and the homogeneous dependent case closes by a direct constant
relation. So the constant-containing mixed-affine branch now really only asks
for exact normalized coordinates of the form \((1,x_0,q_2,q_3)\) with
\(\operatorname{coeff}_{1}(q_i)=\operatorname{coeff}_{x_0}(q_i)=0\).

The next Lean round removed one more artificial hypothesis from the
determinant-nondegenerate sub-branch. I added
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_tailedPair\_det\_general}
and
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_const\_x0\_tailedPair\_det\_general}.
These replay the old exact-coordinate tailed-pair proof directly from exact
relation data, not from an orthonormal factor basis. The key new ingredient is
an internal Gram-degeneracy contradiction for the corrected relation pair:
if the corrected coefficient vectors had zero Gram determinant, the equality
case of Cauchy--Schwarz would force a nontrivial linear relation between them,
but the nonzero
\(
\operatorname{coeff}_{x_0x_1}/\operatorname{coeff}_{x_1^2}
\)
minor of the corrected quadratics rules that out. So the determinant-
nondegenerate mixed-affine branch now accepts the exact relation data that
classification naturally produces, without any separate orthonormalization
step.

The next Lean round stabilized and packaged the remaining purely homogeneous
sub-branch of the constant-containing mixed-affine exact dispatcher. I added
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_homQuadratics\_cases},
along with the supporting linear-algebra lemmas showing that a nontrivial
homogeneous quadratic plane forces nonzero Gram determinant for the relation
pair, while the complementary dependent homogeneous-line case yields an
explicit constant relation and hence closes directly. I then rewired
\texttt{residual\_eq\_zero\_of\_const\_x0\_pair\_coeff\_m00\_m10\_zero} to use
that packaged theorem instead of carrying a long inlined proof of the
``both \(x_1\)-tails vanish'' case. No new Julia experiments were needed for
this round; the change is purely formal consolidation of the homogeneous
mixed-affine branch.

The next Lean round lifted two more pieces of the pure-\(x_0^2\)
determinant-zero branch to relation level. I added
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_x0sqTail\_m02\_nonzero}
and
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_x0sqTail\_m11\_nonzero}.
Both are direct replays of the earlier exact-coordinate proofs, but they now
start from exact relation data rather than from coordinates already equal to
\((1,x_0,q_2,q_3)\). In each case the translation and affine-tail correction
produce a nontrivial homogeneous quadratic plane, and the new
\texttt{\dots\_plane\_nontrivial\_noGram} wrapper closes that plane without any
separate Gram hypothesis. No new Julia experiments were needed for this round.
The pure-\(x_0^2\) determinant-zero branch is therefore now missing only its
residual affine/dependent relation-level subcase, namely the case where the
corrected tail has both \(x_0x_1\)- and \(x_1^2\)-coefficients zero.

The next Lean round removed even that residual gap in the form actually needed
by the extracted-line branch. I added
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_x0sqTail\_m11\_m02\_zero\_of\_m01\_nonzero}
and
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_x0sqTail\_m02\_zero\_of\_m01\_nonzero}.
Because the determinant-zero extracted-line dispatcher always keeps a nonzero
\(x_1\)-tail in the second quadratic, the affine/dependent special slice falls
directly to the relation-level affine-tail theorem, so no extra constant-
relation construction is needed at that interface.

I then lifted the whole pure-\(x_0^2\) extracted-line branch itself to
relation level by adding
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_mixedAffineTailHomLine\_x0sqTail\_cases}
and its equivariant wrapper. This means the classification layer no longer has
to produce exact coordinates even for that slice; exact relation data for
\((1,x_0,q_2,q_3)\) together with the extracted-line coefficient conditions now
suffice.

The next compression packages the whole determinant-zero tailed mixed-affine
branch at relation level with a single natural independence hypothesis. I
added
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_mixedAffineTailHomLine\_origDetZero\_cases\_gram},
plus its equivariant wrapper. The new hypothesis is the nonzero Gram
determinant of the two quadratic relation vectors \(c_2,c_3\). That is exactly
what makes the vanishing extracted-line case honest: if the extracted-line
relation vector collapsed as well, Cauchy--Schwarz would force zero Gram,
contradicting the hypothesis. So Lean can now handle the entire determinant-
zero mixed-affine split internally from relation data.

The same round then packaged the whole constant-containing mixed-affine coarse
surface at relation level too. I added
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0\_pair\_coeff\_m00\_m10\_zero\_gram}
and its equivariant wrapper. So the theorem-level target for that branch is no
longer an exact normalized coordinate picture \((1,x_0,q_2,q_3)\). It is now
enough to extract exact relation data to \(1,x_0,q_2,q_3\), with \(q_2,q_3\)
quadratic and zero constant/\(x_0\) coefficients, together with nonzero Gram
determinant of the last two relation vectors. From there Lean performs the
determinant split, tail swap, extracted-line normalization, homogeneous-plane
fallback, and dependent constant-relation closure internally. No new Julia
experiments were needed for these rounds.

The next round stabilized the affine normalization layer that the final
classification will need anyway. In
\texttt{MixedAffineNormalization.lean} I added the explicit affine
equivalences \texttt{affineLineEquiv} and \texttt{linearPairEquiv}, together
with the exact image lemmas showing that they send a chosen affine line to
\(x_0\) and an independent linear pair to \((x_0,x_1)\). I then used those in
two theorem-level wrappers:
\begin{itemize}
\item \texttt{RepresentativeSpanThree.lean} now proves
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_contains\_aff1} and the
concrete affine-basis wrapper
\texttt{residual\_eq\_zero\_of\_relations\_const\_affinePair\_det}.
\item \texttt{RepresentativeMixedAffinePlane.lean} now proves
\texttt{residual\_eq\_zero\_of\_relations\_const\_affineLine\_pair\_coeff\_m00\_m10\_zero\_gram}.
\end{itemize}
So the remaining classification no longer has to hand back relations already
normalized to \(x_0\) or \((x_0,x_1)\) for those branches. It can work with an
arbitrary affine basis and let Lean perform the explicit affine change of
variables internally. No new Julia experiments were needed for this round.

The next Lean round removed the last external quadratic-normalization burden
from the constant-containing \(\dim=2\) branch itself. In
\texttt{RepresentativeMixedAffinePlane.lean} I added
\texttt{residual\_eq\_zero\_of\_relations\_const\_x0} and
\texttt{residual\_eq\_zero\_of\_relations\_const\_affineLine}. Instead of
asking the classification step to return a quadratic pair
\((q_2,q_3)\) with zero constant/\(x_0\) coefficients and nonzero Gram
determinant, Lean now constructs that pair internally: after normalizing the
chosen affine line to \(x_0\), it forms the coefficient map
\[
  c \longmapsto
  \bigl(\operatorname{coeff}_1,\operatorname{coeff}_{x_0}\bigr)
  \bigl(\sum_i c_i u_i\bigr),
\]
takes a basis of its kernel inside \(\mathbf{R}^4\), and applies the existing
coarse mixed-affine relation theorem to the resulting quadratics. So the
constant-containing mixed-affine \(\dim=2\) branch is no longer waiting on any
external \(q_2,q_3\) extraction at all: theorem-level input has shrunk to
``there is a constant and one nonconstant affine line in the factor span.''
No new Julia experiments were needed for this round.

I then mirrored that same normalization surface on the low-affine side. In
\texttt{RepresentativeLowAffine.lean} I added
\texttt{residual\_eq\_zero\_of\_relations\_linearForm\_homQuadBasis\_det} for
the linear \(\dim=1\) branch, and the four arbitrary-linear-basis plane
wrappers
\texttt{residual\_eq\_zero\_of\_relations\_linearPair\_x0sq\_x0x1Plane},
\texttt{residual\_eq\_zero\_of\_relations\_linearPair\_x0sq\_x1sqPlane},
\texttt{residual\_eq\_zero\_of\_relations\_linearPair\_x0x1\_diffsqPlane}, and
\texttt{residual\_eq\_zero\_of\_relations\_linearPair\_x0x1\_sumsqPlane}. These
use \texttt{linearPairEquiv} or \texttt{affineLineEquiv 0} internally, so the
remaining classification does not need to normalize low-affine branches to
exact \(x_0\) or \((x_0,x_1)\) coordinates either. It can stay at the level of
arbitrary linear bases and transformed homogeneous quadratic relations. No new
Julia experiments were needed for this round.

I then pushed the linear \(\dim=1\) low-affine branch up to the same transport
surface. The new theorems
\texttt{residual\_eq\_zero\_of\_socp\_of\_eq\_mix\_mapVec\_linearForm\_homQuadBasis\_det}
and
\texttt{residual\_eq\_zero\_of\_socp\_of\_eq\_mix\_affineEquiv\_linearForm\_homQuadBasis\_det}
mean that the remaining classifier no longer has to first turn an arbitrary
nonzero affine line into exact \(x_0\) coordinates before invoking the linear
low-affine certificate. No new Julia experiments were needed for this round.

The next Lean rounds removed the analogous external homogeneous-basis burden
from both rank-one affine-image branches. I added the new file
\texttt{RepresentativeAffineRankOne.lean}, imported it through
\texttt{Basic.lean}, and proved
\texttt{residual\_eq\_zero\_of\_relations\_const\_affineRankOne} and then
\texttt{residual\_eq\_zero\_of\_relations\_x0\_affineRankOne}. These theorems
ask for a relation to \(1\) or \(x_0\) together with the rank-one statement
that the affine coefficient map
\[
  c \longmapsto
  \bigl(\operatorname{coeff}_1,\operatorname{coeff}_{x_0},
    \operatorname{coeff}_{x_1}\bigr)\Bigl(\sum_i c_i u_i\Bigr)
\]
has image exactly the corresponding constant or \(x_0\) line. Lean then splits
internally:
\begin{itemize}
\item if the scalar relation map has nontrivial kernel, it closes by the
existing constant-relation certificate;
\item otherwise it extracts a basis of the affine-kernel in \(\mathbf{R}^4\),
proves the induced three quadratic relations are homogeneous and linearly
independent, reconstructs exact \((x_0^2,x_0x_1,x_1^2)\) relations, and
applies the constant or linear representative theorem.
\end{itemize}
So neither rank-one affine-image branch is waiting on an externally supplied
homogeneous quadratic basis either. No new Julia experiments were needed for
these rounds.

I then ran a focused Julia sweep on the unresolved low-affine homogeneous plane
classifier. The new script
\texttt{julia/low\_affine\_hom\_plane\_2026\_04\_10.jl} works directly in the
coefficient model
\[
  (x_0,x_1,q_2,q_3),
  \qquad
  q_j \in \operatorname{span}(x_0^2,x_0x_1,x_1^2),
\]
using only the image matrix \(A_u\). Its recorded outputs are:
\begin{itemize}
\item \texttt{low\_affine\_hom\_plane\_random\_probe = (samples = 20, ranks =
  [14], missing = [(0, 0)])}
\item \texttt{low\_affine\_hom\_plane\_control\_probe = (common\_factor =
  (rank = 13, missing = [(0, 0), (0, 4)]), diagonal = (rank = 14, missing =
  [(0, 0)]), split = (rank = 14, missing = [(0, 0)]))}
\end{itemize}
So the numerical picture is now sharper than before. Random full-rank
homogeneous planes always behaved like the same codimension-one branch, with
only the constant quartic class missing. The only stable rank-\(13\) slice in
that normalized low-affine family is still the common-factor plane
\(\operatorname{span}(x_0^2,x_0x_1)\). This does not prove the missing Lean
classifier, but it makes the likely final theorem surface much narrower: a
generic low-affine homogeneous plane theorem plus the already-formalized
common-factor exception should be enough.

I then tightened the low-affine common-factor interface in Lean without adding
new numerics. In \texttt{RepresentativeLowAffine.lean} I added the swapped
common-factor plane theorem
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_x0x1\_x1sqPlane} and its
arbitrary-linear-basis wrapper
\texttt{residual\_eq\_zero\_of\_relations\_linearPair\_x0x1\_x1sqPlane}. So
the repeated-line low-affine branch no longer depends on an external variable
swap before entering the theorem layer: both
\(\operatorname{span}(x_0^2,x_0x_1)\) and
\(\operatorname{span}(x_0x_1,x_1^2)\) are now directly available at the same
\texttt{linearPair} surface.

I then tried to formalize the generic non-common-factor low-affine homogeneous
plane theorem directly in \texttt{RepresentativeLowAffine.lean}. The retained
part of that attempt is the new coefficient-space helper layer
\texttt{lowHomQuadPlaneA/B/C} together with the exact relation lemmas they
satisfy on both quadratics in the plane. Those are the cross-product scalars of
the \(2 \times 3\) homogeneous coefficient matrix, and they encode the
canonical annihilator line of the plane. The larger dispatcher theorem itself
was backed out after repeated heartbeat blowups, so the repository stays
build-clean.

I then rebuilt a first stable piece of that missing classifier on top of the
helper layer instead of retrying the whole dispatcher. The file now proves
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_homQuadratics\_crossDet\_zero},
together with the matching equivariant and \texttt{linearPair} wrappers. This
covers the homogeneous low-affine chart where the
\((m_{11},m_{02})\)-minor vanishes and the \((m_{20},m_{02})\)-minor is
nonzero: Lean performs the needed \(x_1\)-shear internally, kills the cross
coefficients on both quadratics, and reduces the branch to the already solved
\(\operatorname{span}(x_0^2,x_1^2)\) plane. No new Julia experiments were
needed for this round.

I then closed the next low-affine homogeneous chart in the same style instead
of reopening the generic dispatcher. The file now also proves
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_homQuadratics\_commonFactorChart},
again with the equivariant and \texttt{linearPair} wrappers. This covers the
repeated-line chart where the canonical annihilator has nonzero
\((m_{11},m_{02})\)-minor but zero post-shear diagonal term
\[
  \texttt{lowHomQuadPlaneC} - \texttt{lowHomQuadPlaneB}^2 /
  \texttt{lowHomQuadPlaneA} = 0.
\]
Lean performs the same internal \(x_1\)-shear, proves that both transformed
quadratics have vanishing \(x_0^2\) coefficient, and dispatches to the already
formalized common-factor plane
\(\operatorname{span}(x_0x_1,x_1^2)\). No new Julia experiments were needed for
this round either.

I then filled in the scale-only theorem surface for the last diagonal
low-affine homogeneous branch before reintroducing another outer chart wrapper.
The file now proves
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_homQuadratics\_diag\_sum\_zero}
and
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_homQuadratics\_diag\_diff\_zero},
together with the equivariant and \texttt{linearPair} wrappers for each. These
assume the quadratics already satisfy a diagonal relation
\[
  a\,\operatorname{coeff}_{x_0^2}(q)+d\,\operatorname{coeff}_{x_1^2}(q)=0
\]
with the relevant sign on \(d/a\), then perform the internal \(x_1\)-scaling
and reduce to the already solved
\(\operatorname{span}(x_0x_1,x_0^2-x_1^2)\) or
\(\operatorname{span}(x_0x_1,x_0^2+x_1^2)\) plane. No new Julia experiments
were needed for this round either.

I then added the missing outer diagonal chart wrappers on top of that scale-only
surface instead of trying to jump straight to the root theorem. The file now
proves
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_homQuadratics\_diagSumChart}
and
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_homQuadratics\_diagDiffChart}.
These start from the raw cross-product scalars
\texttt{lowHomQuadPlaneA/B/C}, perform the internal \(x_1\)-shear killing the
cross term in the dual annihilator, and then dispatch to the already-built
diagonal scale theorems according to the sign of the post-shear scalar
\[
  \texttt{lowHomQuadPlaneC} - \texttt{lowHomQuadPlaneB}^2 /
  \texttt{lowHomQuadPlaneA}.
\]

On top of those new chart theorems, the same file now proves the coarse
homogeneous low-affine dispatcher
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_homQuadratics\_plane\_nontrivial},
together with the matching equivariant and \texttt{linearPair} wrappers. So the
low-affine homogeneous side is no longer an external classification problem at
the theorem surface: if the normalized homogeneous quadratic plane is nontrivial,
Lean now internally distinguishes the cross chart, the two common-factor
charts, and the two diagonal rank-\(14\) charts. No new Julia experiments were
needed for this round.

I then cleaned the same theorem surface one step further. The raw coefficient
minors \texttt{lowHomQuadPlaneA/B/C} are now public, and
\texttt{RepresentativeLowAffine.lean} proves the reusable criterion
\texttt{lowHomQuadPlane\_nontrivial\_of\_independent\_pair}: for homogeneous
quadratics, linear independence already forces one of those three minors to be
nonzero. On top of that, the new wrapper
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_homQuadratics\_independent}
lets later classifier code feed the low-affine branch with linear independence
instead of carrying its own separate nontrivial-plane proof. No new Julia
experiments were needed for this round either.

I then pushed that same low-affine homogeneous theorem all the way up to the
transport interface. The file now also proves
\texttt{residual\_eq\_zero\_of\_socp\_of\_eq\_mix\_mapVec\_x0\_x1\_homQuadratics\_plane\_nontrivial}
and
\texttt{residual\_eq\_zero\_of\_socp\_of\_eq\_mix\_affineEquiv\_x0\_x1\_homQuadratics\_plane\_nontrivial}.
So the low-affine homogeneous branch now matches the mixed-affine and rank-one
branches at the theorem surface: after affine normalization and orthogonal
mixing, an exact coordinate picture \((x_0,x_1,q_2,q_3)\) with nontrivial
homogeneous plane data is enough to close the SOCP point directly. Again, no
new Julia experiments were needed for this round.

I then started moving the global classifier itself into Lean, rather than only
the representative and plane endpoints. The new file
\texttt{ExactAffineClassification.lean} defines the exact affine relation space
as the kernel of the homogeneous coefficient map
\[
  c \mapsto
  \bigl(\operatorname{coeff}_{x_0^2},
    \operatorname{coeff}_{x_0x_1},
    \operatorname{coeff}_{x_1^2}\bigr)\Bigl(\sum_i c_i u_i\Bigr).
\]
The first theorem there,
\texttt{residual\_eq\_zero\_of\_exactAffineDimThree}, shows that if this kernel
has dimension \(3\) and the scalar relation map is injective, Lean reconstructs
exact relations to \(1\), \(x_0\), and \(x_1\) internally and closes by the
already-formalized \texttt{contains\_aff1} theorem. The second theorem,
\texttt{residual\_eq\_zero\_of\_exactAffineDimTwo\_contains\_one}, handles the
constant-containing exact-affine \(\dim=2\) branch in the same language: from a
single exact constant relation and the dimension-\(2\) kernel hypothesis, Lean
extracts a second independent exact affine line and closes by the coarse
mixed-affine theorem \texttt{residual\_eq\_zero\_of\_relations\_const\_affineLine}.
So the final extraction step is no longer purely external: two global outputs
are now packaged directly in terms of the kernel of the homogeneous coefficient
map. No new Julia experiments were needed for this round either.

I then added the next no-constant extraction lemma in the same file instead of
jumping prematurely to a larger residual theorem. The new theorem
\texttt{exists\_exactAffine\_affinePair\_of\_dimTwo\_noConst} proves that if the
exact affine kernel has dimension \(2\) and contains no exact constant
relation, then the projection to the linear coefficients
\((\operatorname{coeff}_{x_0}, \operatorname{coeff}_{x_1})\) is automatically an
isomorphism. Lean therefore reconstructs exact relations to a normalized affine
pair \(x_0 + r_0\), \(x_1 + r_1\) internally. So the remaining no-constant
\(\dim=2\) gap is no longer the affine-pair extraction itself; it is the next
quadratic-package theorem above that pair. No new Julia experiments were needed
for this round either.

I then formalized exactly that next package surface on the low-affine side. The
new theorem
\texttt{residual\_eq\_zero\_of\_relations\_affinePair\_homQuadratics\_independent}
in \texttt{RepresentativeLowAffine.lean} starts from exact relations to
\(x_0 + r_0\), \(x_1 + r_1\), and a quadratic pair \(q_2,q_3\). If the common
translation by \((-r_0,-r_1)\) makes \(q_2,q_3\) homogeneous and linearly
independent, Lean now performs that translation internally and closes through
the existing low-affine homogeneous plane theorem. So the remaining
no-constant \(\dim=2\) gap is now only the extraction of the right quadratic
pair above the affine pair, not the affine-pair normalization itself. No new
Julia experiments were needed for this round either.

I then tightened the linear-algebra side of that same no-constant branch in
\texttt{ExactAffineClassification.lean}. The new theorems
\texttt{translatedLinearCoeffMap\_surjective\_of\_affinePair} and
\texttt{translatedLinearCoeffKernel\_finrank\_two\_of\_affinePair} prove that
once an affine pair \(x_0 + r_0\), \(x_1 + r_1\) is known, the common
translation by \((-r_0,-r_1)\) automatically produces a surjective linear
coefficient map onto the normalized linear basis \((x_0,x_1)\), and hence a
two-dimensional kernel of translated zero-linear-tail relations. So the
remaining no-constant \(\dim=2\) gap is now even sharper: the open work is not
to recover that kernel, but to analyze the constant-term functional on it and
separate the mixed-affine and low-affine outcomes. No new Julia experiments
were needed for this round either.

I then formalized the zero-constant side of that split instead of guessing at
the complementary mixed-affine branch. The new theorem
\texttt{residual\_eq\_zero\_of\_relations\_affinePair\_translatedKernel\_constZero}
in \texttt{ExactAffineClassification.lean} starts from an exact affine pair
\(x_0+r_0\), \(x_1+r_1\) and the hypothesis that every relation in the common
translated zero-linear-tail kernel already has zero translated constant term.
Lean chooses a basis of that two-dimensional kernel, proves the translated
basis images are homogeneous and linearly independent, and dispatches directly
to the previously-built low-affine theorem
\texttt{residual\_eq\_zero\_of\_relations\_affinePair\_homQuadratics\_independent}.
So the no-constant \(\dim=2\) classifier gap is now split cleanly: the
constant-zero half is done, and only the complementary nonzero-constant half
remains. No new Julia experiments were needed for this round either.

I then formalized the linear-algebra core of that remaining half. The new
theorem \texttt{exists\_translatedKernel\_constSplit\_of\_affinePair} in
\texttt{ExactAffineClassification.lean} assumes the same exact affine pair and
the existence of some translated zero-linear-tail relation with nonzero
constant term. Lean then constructs a normalized pair inside that translated
kernel: one relation has constant coefficient \(1\), and a second nonzero
relation has constant coefficient \(0\). So the unresolved no-constant
\(\dim=2\) branch is now reduced to a concrete constant-perturbed low-affine
input \((1+H_1, H_2)\), not an abstract functional split.

I then probed that complementary nonzero-constant half directly in normalized
coordinates rather than guessing its theorem shape. The new script
\texttt{julia/low\_affine\_const\_perturbation\_2026\_04\_10.jl} tests models
\((x_0, x_1, a + H_1, b + H_2)\) where \((H_1,H_2)\) spans a low-affine
homogeneous plane. The recorded output is
\[
\texttt{low\_affine\_const\_perturbation\_probe}
  = (\texttt{generic\_ranks}=[15],\ \dots).
\]
So all \(20\) random constant perturbations were rank \(15\). The controls show
the same pattern for the coprime, diagonal, and split planes, while the
constant-perturbed common-factor slice remains rank \(14\), missing only
\((0,4)\). This sharply changes the expected final split: the nonzero-constant
branch now looks generically surjective, with only the repeated-line/common-factor
subcase still behaving like a genuine low-rank obstruction.

I then converted the easiest of those generic-looking slices into Lean. In
\texttt{TernaryQuarticProof/RepresentativeLowAffine.lean} I added
\texttt{quartic\_in\_image\_of\_relations\_x0\_x1\_onePlusX0sq\_x1sq} and
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_onePlusX0sq\_x1sq}. So the
coprime constant-perturbed model \((x_0, x_1, 1 + x_0^2, x_1^2)\) is now
formally proved quartic-surjective at the exact-relation level, matching the
rank-\(15\) numerical probe. I then lifted that slice to the affine-pair
surface in the same file via
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_x0\_x1\_onePlusX0sq\_x1sq}
and
\texttt{residual\_eq\_zero\_of\_relations\_affinePair\_onePlusX0sq\_x1sq}, so
later classifier code will not need to normalize that case all the way down to
literal \((x_0, x_1, 1 + x_0^2, x_1^2)\) coordinates by hand.

I then formalized the split constant-perturbed slice in the same file. The new
theorems
\texttt{quartic\_in\_image\_of\_relations\_x0\_x1\_onePlusX0x1\_diffsq},
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_onePlusX0x1\_diffsq},
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_x0\_x1\_onePlusX0x1\_diffsq},
and
\texttt{residual\_eq\_zero\_of\_relations\_affinePair\_onePlusX0x1\_diffsq}
show that
\[
  (x_0,\ x_1,\ 1 + x_0x_1,\ x_0^2 - x_1^2)
\]
is quartic-surjective and transports across the same affine-pair surface. No
new Julia experiments were needed for this round. This narrows the remaining
nonzero-constant exact-affine \(\dim=2\) work again: the rank-\(15\) coprime
and split deformations are now theorem-level endpoints, so the only visible
special low-rank constant-perturbed slice is the repeated-line/common-factor
case.

I then formalized that repeated-line/common-factor slice in the same file. The
new theorems
\texttt{quartic\_in\_image\_of\_relations\_x0\_x1\_onePlusX0sq\_x0x1\_of\_coeff\_m04\_zero},
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_onePlusX0sq\_x0x1},
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_x0\_x1\_onePlusX0sq\_x0x1},
and
\texttt{residual\_eq\_zero\_of\_relations\_affinePair\_onePlusX0sq\_x0x1}
show that
\[
  (x_0,\ x_1,\ 1 + x_0^2,\ x_0x_1)
\]
has the expected theorem-level rank-\(14\) structure: Lean proves quartic image
surjectivity modulo the single missing \(x_1^4\) class, then removes that class
with the already formalized common-factor quartic kernel. No new Julia probe was
needed for this round either. This means the nonzero-constant exact-affine
\(\dim=2\) branch is no longer waiting on another representative theorem; the
remaining work is the classifier that routes the translated-kernel output into
the coprime, split, or repeated-line surface.

I then added the symmetric repeated-line/common-factor orientation in the same
file. The new theorems
\texttt{quartic\_in\_image\_of\_relations\_x0\_x1\_onePlusX0x1\_x1sq\_of\_coeff\_m40\_zero},
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_onePlusX0x1\_x1sq},
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_x0\_x1\_onePlusX0x1\_x1sq},
and
\texttt{residual\_eq\_zero\_of\_relations\_affinePair\_onePlusX0x1\_x1sq}
show that
\[
  (x_0,\ x_1,\ 1 + x_0x_1,\ x_1^2)
\]
has the symmetric rank-\(14\) structure, with only the \(x_0^4\) class missing
before the internal quartic-kernel correction. No new Julia probe was needed.
This is useful because the already-formalized homogeneous common-factor chart
naturally lands in the \((x_0x_1, x_1^2)\) orientation after shear, so the
remaining no-constant exact-affine \(\dim=2\) gap is now even more clearly a
classifier/transport problem rather than a missing endpoint theorem.

I then formalized the dependent half of the translated-kernel constant-split
branch. In \texttt{TernaryQuarticProof/RepresentativeLowAffine.lean}, the new
theorem
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_onePlus\_homQuadratics\_dependent}
shows that if the normalized constant-perturbed pair has the form
\((1 + H,\ tH)\) with zero linear coefficients and nonzero second relation,
then Lean can recover an exact constant relation from those two quadratics and
close through \(\operatorname{span}(1,x_0,x_1)\). I then lifted that directly
into the classifier file as
\texttt{residual\_eq\_zero\_of\_relations\_affinePair\_translatedKernel\_constSplit\_dependent}
in \texttt{ExactAffineClassification.lean}. So the no-constant exact-affine
\(\dim=2\) branch is no longer open on the whole nonzero-constant side: the
translated-kernel constant-split case now only remains unresolved when its
homogeneous quadratic part is genuinely independent. No new Julia experiments
were needed for this round.

I then tightened that dependent branch to the natural linear-algebra input. In
\texttt{RepresentativeLowAffine.lean}, the theorem
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_onePlus\_homQuadratics\_not\_independent}
shows that Lean no longer needs the three explicit equations
\texttt{lowHomQuadPlaneA = lowHomQuadPlaneB = lowHomQuadPlaneC = 0} to enter
the dependent constant-perturbed branch. It is enough to assume
\(\neg\,\operatorname{LinearIndependent}(q_2 - 1, q_3)\) together with
\(q_3 \neq 0\); Lean now extracts the scalar-multiple relation internally and
reuses the existing dependent theorem. I lifted that immediately into
\texttt{ExactAffineClassification.lean} as
\texttt{residual\_eq\_zero\_of\_relations\_affinePair\_translatedKernel\_constSplit\_not\_independent}.
So the classifier can now branch directly on linear independence of the
homogeneous pair. No new Julia experiment was needed for this round.

I then widened the exact coprime constant-perturbed endpoint from the
normalized model \(1 + x_0^2\) to the whole nonzero scalar family
\[
  (x_0,\ x_1,\ 1 + \alpha x_0^2,\ x_1^2), \qquad \alpha \neq 0.
\]
In \texttt{RepresentativeLowAffine.lean}, the new theorems
\texttt{quartic\_in\_image\_of\_relations\_x0\_x1\_onePlusAX0sq\_x1sq},
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_onePlusAX0sq\_x1sq},
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_x0\_x1\_onePlusAX0sq\_x1sq},
and
\texttt{residual\_eq\_zero\_of\_relations\_affinePair\_onePlusAX0sq\_x1sq}
show that Lean no longer needs the square-line coefficient normalized to
\(1\) before entering that exact theorem surface. The earlier
\texttt{onePlusX0sq\_x1sq} theorems are now just the \(\alpha = 1\) wrappers.
No new Julia probe was needed for this round; this was a formalization-only
generalization aimed at the still-open independent translated-kernel branch.

I then ran the previously prepared script
\texttt{julia/low\_affine\_const\_perturbation\_2026\_04\_10.jl} and recorded:
\[
\texttt{low\_affine\_const\_perturbation\_probe}
= (\texttt{samples} = 20,\ \texttt{generic\_ranks} = [15],\
\texttt{controls} = (\texttt{coprime\_const} = 15,\ \texttt{common\_factor\_const} = 14,\ \texttt{diagonal\_const} = 15,\ \texttt{split\_const} = 15)).
\]
So the generic constant-perturbed low-affine plane is numerically surjective,
and the only stable nonsurjective control among those canonical slices is the
common-factor model \((x_0,\ x_1,\ 1 + x_0^2,\ x_0x_1)\). This makes the
remaining independent translated-kernel branch look less like a missing family
of rank-\(14\) endpoints and more like a classifier problem whose exceptional
output is specifically the repeated-line/common-factor surface.

I then widened that repeated-line/common-factor endpoint on both canonical
scalar families in \texttt{RepresentativeLowAffine.lean}. The file now proves
the nonzero-\(\alpha\) family
\[
  (x_0,\ x_1,\ 1 + \alpha x_0^2,\ x_0x_1),\qquad \alpha \neq 0,
\]
via the new theorems
\texttt{quartic\_in\_image\_of\_relations\_x0\_x1\_onePlusAX0sq\_x0x1\_of\_coeff\_m04\_zero},
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_onePlusAX0sq\_x0x1}, and
\texttt{residual\_eq\_zero\_of\_relations\_affinePair\_onePlusAX0sq\_x0x1}.
It also proves the symmetric nonzero-\(\beta\) family
\[
  (x_0,\ x_1,\ 1 + \beta x_0x_1,\ x_1^2),\qquad \beta \neq 0,
\]
via
\texttt{quartic\_in\_image\_of\_relations\_x0\_x1\_onePlusBX0x1\_x1sq\_of\_coeff\_m40\_zero},
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_onePlusBX0x1\_x1sq}, and
\texttt{residual\_eq\_zero\_of\_relations\_affinePair\_onePlusBX0x1\_x1sq}.
So the repeated-line branch no longer needs those coefficients normalized to
\(1\) before entering the theorem surface.

I then recorded the matching scalar sweep in the new script
\texttt{julia/low\_affine\_common\_factor\_scalars\_2026\_04\_10.jl}. Its
output is
\[
\texttt{low\_affine\_common\_factor\_scalar\_probe}
= (\texttt{x0sq\_tail} = [\dots],\ \texttt{x1sq\_tail} = [\dots]).
\]
For every tested
\(\alpha,\beta \in \{-3,-2,-1,-\tfrac12,\tfrac12,1,2,3\}\), the family
\((x_0,\ x_1,\ 1+\alpha x_0^2,\ x_0x_1)\) had rank \(14\) with only
\((0,4)\) missing, while the family
\((x_0,\ x_1,\ 1+\beta x_0x_1,\ x_1^2)\) had rank \(14\) with only
\((4,0)\) missing. So the repeated-line/common-factor output now looks stable
across its whole nonzero scalar family, both numerically and in Lean.

I then lifted that repeated-line/common-factor branch one level higher in
\texttt{TernaryQuarticProof/RepresentativeLowAffine.lean}. The new theorem
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_onePlusAX0sq\_x0x1Plane}
takes exact relations to
\[
  x_0,\quad x_1,\quad 1 + a x_0^2 + b x_0x_1,\quad c x_0^2 + d x_0x_1
\]
with \(ad-bc \neq 0\), and internally reduces them to one of the already solved
scalar families. The \(d=0\) branch cancels the \(x_0^2\) term, rescales to
\((1+\beta x_0x_1,\ x_1^2)\), and swaps coordinates; the \(d \neq 0\) branch
uses an internal \(x_1\)-shear to kill the \(x_0^2\) term in the second
quadratic and then subtracts off the cross term in the first, reducing to
\((1+\alpha x_0^2,\ x_0x_1)\). So the classifier will only need determinant-nondegenerate
common-factor plane data, not already-normalized scalar representatives.

I then filled the last orientation gap in that same common-factor plane layer.
In \texttt{RepresentativeLowAffine.lean}, the new theorem
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_onePlusBX0x1\_x1sqPlane}
proves the swapped determinant-nondegenerate model
\[
  (x_0,\ x_1,\ 1 + a x_0x_1 + b x_1^2,\ c x_0x_1 + d x_1^2)
\]
directly by transporting it through the fixed coordinate swap. I also added the
missing equivariant wrapper
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_x0\_x1\_onePlusAX0sq\_x0x1Plane},
so both common-factor plane orientations now sit on the same transport-ready
surface. No new Julia experiments were needed for this round.

I then lifted the same repeated-line/common-factor idea one step upward inside
\texttt{RepresentativeLowAffine.lean}. The new theorem
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_onePlus\_homQuadratics\_commonFactorChart}
starts from exact relations to \(x_0\), \(x_1\), a quadratic \(q_2\) with
constant coefficient \(1\) and zero linear terms, and a quadratic \(q_3\) with
zero constant and linear terms, under the same cross-product chart conditions
\[
  \texttt{lowHomQuadPlaneA}(q_2,q_3) \neq 0,\qquad
  \texttt{lowHomQuadPlaneC}(q_2,q_3) -
    \texttt{lowHomQuadPlaneB}(q_2,q_3)^2 /
    \texttt{lowHomQuadPlaneA}(q_2,q_3) = 0.
\]
Lean then performs the internal \(x_1\)-shear, proves that both transformed
quadratics have zero \(x_0^2\) coefficient, reconstructs the first as
\(1 + \beta x_0x_1 + \gamma x_1^2\), and dispatches to the already solved
\texttt{...onePlusBX0x1\_x1sqPlane} theorem. No new Julia experiment was needed
for this round either.

I then lifted that chart theorem into
\texttt{ExactAffineClassification.lean} as
\texttt{residual\_eq\_zero\_of\_relations\_affinePair\_translatedKernel\_constSplit\_commonFactorChart}.
So the exact-affine \(\dim = 2\), no-constant branch now closes automatically
on the independent translated-kernel subcase whenever the normalized
constant-split pair lands in the repeated-line/common-factor chart. That means
the still-open independent translated-kernel work is no longer the whole
constant-perturbed low-affine layer; it is only the complementary charts.

I then cleaned that same chart up to the standard transport surface in
\texttt{RepresentativeLowAffine.lean}. The file now also proves
\texttt{residual\_eq\_zero\_of\_equiv\_relations\_x0\_x1\_onePlus\_homQuadratics\_commonFactorChart}
and
\texttt{residual\_eq\_zero\_of\_relations\_affinePair\_onePlus\_homQuadratics\_commonFactorChart}.
So the repeated-line/common-factor constant-perturbed chart is no longer only a
normalized \((x_0,x_1)\)-theorem. It now sits on the same
\texttt{equiv}/\texttt{affinePair} interface as the other low-affine theorem
surfaces. No new Julia experiment was needed for this round.

I then added the analogous coprime cross-chart theorem surface.
\texttt{RepresentativeLowAffine.lean} now proves
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_onePlus\_homQuadratics\_crossDet\_zero},
its \texttt{equiv} wrapper, and
\texttt{residual\_eq\_zero\_of\_relations\_affinePair\_onePlus\_homQuadratics\_crossDet\_zero}.
The proof reuses the existing homogeneous \(x_1\)-shear algebra: when the
translated homogeneous pair has \(\texttt{lowHomQuadPlaneA}=0\) and
\(\texttt{lowHomQuadPlaneB}\neq 0\), Lean shears internally, proves that the
mixed term disappears, reconstructs a constant-perturbed diagonal pair, and
dispatches to \texttt{...onePlusAX0sq\_x1sqPlane}. No new Julia experiment was
needed for this round either.

I then lifted that coprime cross-chart theorem into
\texttt{ExactAffineClassification.lean} as
\texttt{residual\_eq\_zero\_of\_relations\_affinePair\_translatedKernel\_constSplit\_crossDet\_zero}.
So the exact-affine \(\dim = 2\), no-constant, independent translated-kernel
branch is no longer open on the coprime cross chart either. The still-open
independent translated-kernel work is now narrower again: it remains in the
complementary constant-perturbed diagonal/split outputs, plus the final global
assembly.

I then closed the constant-perturbed sum-square side of that complementary
output. In \texttt{RepresentativeLowAffine.lean} I added
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_onePlusX0x1\_sumsqPlane},
which sends the \((x_0x_1,\ x_0^2+x_1^2)\)-plane through the fixed
\((x_0,x_1) \mapsto (x_0+x_1,\ x_0-x_1)\) transform and reduces it to the
already solved \texttt{...onePlusAX0sq\_x1sqPlane} theorem. Using that, the
same file now proves
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_onePlus\_homQuadratics\_diagDiffChart}.
So when the normalized constant-perturbed low-affine pair lands in the
positive-sign diagonal-difference chart, Lean performs the internal
\(x_1\)-shear and \(x_1\)-scaling and closes it automatically.

I then lifted that chart into \texttt{ExactAffineClassification.lean} as
\texttt{residual\_eq\_zero\_of\_relations\_affinePair\_translatedKernel\_constSplit\_diagDiffChart}.
So the exact-affine \(\dim = 2\), no-constant, independent translated-kernel
branch is no longer open on the sum-square side either. After this round, the
remaining independent translated-kernel low-affine work is concentrated in the
opposite-sign trace-free branch, i.e.\ the \texttt{diagSumChart} output. No
new Julia experiment was needed for this round.

I then stopped treating the independent constant-perturbed low-affine branch as
a bag of chart wrappers and packaged it into a single theorem surface in
\texttt{RepresentativeLowAffine.lean}. The new theorem
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1\_onePlus\_homQuadratics\_independent}
starts from exact relations to \(x_0\), \(x_1\), \(q_2\), \(q_3\), the usual
constant/linear coefficient conditions, and linear independence of
\((q_2-1,q_3)\). Lean now performs the whole internal branch split itself:
repeated-line \((x_0^2,x_0x_1)\), coprime cross chart, common-factor chart,
and the two diagonal charts. I also added the affine-pair transport wrapper
\texttt{residual\_eq\_zero\_of\_relations\_affinePair\_onePlus\_homQuadratics\_independent}.
No new Julia experiment was needed for this round.

With that in place, I added the matching exact-affine wrapper
\texttt{residual\_eq\_zero\_of\_relations\_affinePair\_translatedKernel\_constSplit\_independent}
and then the global theorem
\texttt{residual\_eq\_zero\_of\_exactAffineDimTwo\_noConst} in
\texttt{ExactAffineClassification.lean}. So the exact-affine \(\dim = 2\),
no-constant branch is no longer open at all: Lean now extracts the affine pair,
splits the translated zero-linear-tail kernel into the all-zero-constant or
constant-split cases, and closes both branches internally. After this round,
the main global blocker is no longer the exact-affine \(\dim = 2\) analysis. It
is the remaining affine-rank-one extraction/classification layer above the
already-formalized rank-one endpoint theorems, followed by the final assembly
in \texttt{TernaryQuarticProof.lean}.

The next rank-one probe clarified one more endpoint. I added
\texttt{julia/affine\_rank\_one\_nonzero\_const\_2026\_04\_11.jl} and checked
the normalized family
\[
  (\alpha + x_0,\ q_1,\ q_2,\ q_3),
  \qquad \operatorname{span}(q_1,q_2,q_3)=
    \operatorname{span}(x_0^2,x_0x_1,x_1^2).
\]
The recorded result is
\[
  \texttt{affine\_rank\_one\_nonzero\_const\_probe}
  = (\texttt{generic\_ranks}=[15],\dots),
\]
with the canonical scalar sweep showing rank \(15\) for every tested
\(\alpha \in \{-3,-2,-1,-\tfrac12,\tfrac12,1,2,3\}\), while the limiting
\(\alpha=0\) model drops back to the old linear rank-\(13\) pattern missing
\((0,0)\) and \((0,1)\). So the nonzero-constant affine-rank-one endpoint now
looks surjective rather than ``image plus kernel.'' I then formalized exactly
that in Lean:
\texttt{RepresentativeSurjective.lean} now proves
\texttt{quartic\_in\_image\_of\_relations\_constX0\_x0sq\_x0x1\_x1sq},
\texttt{residual\_eq\_zero\_of\_relations\_constX0\_x0sq\_x0x1\_x1sq}, and
\texttt{residual\_eq\_zero\_of\_relations\_constX0\_homQuadBasis\_det}, and
\texttt{RepresentativeAffineRankOne.lean} lifts that to the rank-one surface
\texttt{residual\_eq\_zero\_of\_relations\_constX0\_affineRankOne}. So the
rank-one layer now has a third endpoint beside the pure constant and pure
linear cases.

That did not actually collapse the whole remaining \(\dim=1\) branch. The new
probe \texttt{julia/affine\_dim\_one\_tailed\_2026\_04\_11.jl} now records both
image ranks and exact-affine dimensions. It first tests the three normalized
tailed families
\[
  (x_0,\ x_1+\lambda x_0x_1,\ x_0^2,\ x_1^2),\qquad
  (x_0,\ x_1+\lambda x_0x_1,\ x_0^2,\ x_0x_1),\qquad
  (x_0,\ x_1+\lambda x_0^2,\ x_0^2,\ x_1^2),
\]
over \(\lambda \in \{-3,-2,-1,-\tfrac12,\tfrac12,1,2,3\}\). The recorded
output is
\[
  \texttt{affine\_dim\_one\_tailed\_probe}
  = (\texttt{cross\_tail\_ranks}=[14],\ 
     \texttt{common\_factor\_tail\_ranks}=[13],\
     \texttt{x0sq\_tail\_ranks}=[14],\
     \texttt{cross\_tail\_exact\_affine\_dims}=[1],\
     \texttt{common\_factor\_tail\_exact\_affine\_dims}=[2],\
     \texttt{x0sq\_tail\_exact\_affine\_dims}=[2],\dots),
\]
with stable missing sets
\[
  \texttt{cross\_tail\_missing}=[[(0,0)]],\qquad
  \texttt{common\_factor\_tail\_missing}=[[(0,0),(0,4)]],\qquad
  \texttt{x0sq\_tail\_missing}=[[(0,0)]].
\]
So only the cross slice is genuinely exact-affine \(\dim=1\) among those
three; the common-factor and \(x_0^2\)-tail slices already belong to the
exact-affine \(\dim=2\) branch because the homogeneous plane already contains
the tailed monomial.

I then formalized that genuine cross slice in
\texttt{RepresentativeAffineRankOne.lean}. The new theorem
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1PlusAX0x1\_x0sq\_x1sq}
proves zero residual from exact relations to
\[
  (x_0,\ x_1 + a x_0x_1,\ x_0^2,\ x_1^2),\qquad a \ne 0.
\]
The image proof uses the explicit identity
\[
  x_1 = (x_1 + a x_0x_1)(1-a x_0) + a^2 x_0^2 x_1,
\]
and the kernel proof matches the missing constant coefficient by the admissible
direction built from the exact \(x_0\) and \(x_0^2\) relations.

The same updated Julia script then checked two further genuinely
exact-affine-\(\dim=1\) slices:
\[
  (x_0,\ x_1+\lambda x_1^2,\ x_0^2,\ x_0x_1),\qquad
  (x_0,\ x_1+\lambda x_0^2,\ x_0x_1,\ x_1^2).
\]
For every tested nonzero \(\lambda\), the first has exact-affine dimension
\(1\), image rank \(13\), and missing set
\((0,0),(0,1),(0,2),(0,3),(0,4)\), while the second has exact-affine
dimension \(1\), image rank \(14\), and only the constant monomial missing.

I then formalized that second slice too in
\texttt{RepresentativeAffineRankOne.lean}. The new theorem
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1PlusAX0sq\_x0x1\_x1sq}
proves zero residual from exact relations to
\[
  (x_0,\ x_1 + a x_0^2,\ x_0x_1,\ x_1^2),\qquad a \ne 0.
\]
The proof is simpler than the cross slice: Lean first recovers \(x_1\) in the
admissible image by subtracting the exact \(x_0^2\)-image from
\(x_1 + a x_0^2\), then reaches every nonconstant quartic monomial using the
exact relations to \(x_0\), \(x_0x_1\), and \(x_1^2\). The missing constant is
handled by the admissible kernel direction built from the exact \(x_0\) and
\(x_0x_1\) relations.

I then formalized the repeated-line slice too. The new theorem
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1PlusAX1sq\_x0sq\_x0x1}
proves zero residual from exact relations to
\[
  (x_0,\ x_1 + a x_1^2,\ x_0^2,\ x_0x_1),\qquad a \ne 0.
\]
Here the image proof splits quartics into a pure-\(x_1\) piece, handled by the
divisibility identity through \(x_1 + a x_1^2\), and an \(x_0\)-divisible
remainder, handled directly from the exact \(x_0\), \(x_0^2\), and \(x_0x_1\)
relations. The kernel part uses a two-parameter \(x_0/x_0^2\)-based admissible
family, matching both the missing constant coefficient and the extracted
\texttt{x1sqTailFunctional} square-by-square in the SOS decomposition.

So the surviving \(\dim=1\) work is now purely classifier-level: all three
visible normalized tailed endpoints are closed in Lean, and the remaining gap
is to extract one of those solved surfaces, or the pure affine-rank-one
surface, from a general exact-affine-\(\dim=1\) factor.

I then added the matching affine-equivalence transport wrappers in
\texttt{RepresentativeAffineRankOne.lean} for all three normalized tailed
\(\dim=1\) endpoints:
\texttt{...equiv\_relations\_x0\_x1PlusAX0x1\_x0sq\_x1sq},
\texttt{...equiv\_relations\_x0\_x1PlusAX0sq\_x0x1\_x1sq}, and
\texttt{...equiv\_relations\_x0\_x1PlusAX1sq\_x0sq\_x0x1}. No new numerical
claim was needed for that round. The effect is only on the theorem surface:
the remaining classifier no longer has to carry raw transported relation data
for these three tailed branches. It can instead target exact normalized
coordinates after affine normalization, in the same style as the mixed-affine
and low-affine transport layers.

I then added the first retained extraction lemma for that remaining branch in
\texttt{ExactAffineClassification.lean}:
\texttt{exists\_exactAffine\_affineLine\_of\_dimOne\_noConst}. This is a small
but real classifier-facing gain. Under exact-affine dimension \(1\), injectivity
of the scalar relation map, and absence of an exact constant relation, Lean now
extracts a nonconstant exact affine line
\[
  r + a x_0 + b x_1,\qquad a^2+b^2 \ne 0,
\]
by contradiction: if the linear part vanished, the unique exact-affine
direction would scale to an exact constant relation. No new Julia evidence was
needed for this round.

I then pushed that same branch one layer deeper without touching the remaining
tailed endpoint algebra. The same file now proves
\texttt{exists\_relations\_x0\_homQuadBasis\_of\_exactAffineDimOne}: once an
exact-affine-\(\dim=1\) point has been normalized so that one exact relation is
\(x_0\), Lean constructs three more exact relations
\[
  \alpha_{20} + \beta_{20}x_1 + x_0^2,\qquad
  \alpha_{11} + \beta_{11}x_1 + x_0x_1,\qquad
  \alpha_{02} + \beta_{02}x_1 + x_1^2.
\]
So the classifier no longer has to re-extract the homogeneous quadratic basis
inside the rank-\(4\) span; it only has to analyze the six residual scalar tail
coefficients.

I also packaged the easy subcase sitting above that structure. The theorem
\texttt{residual\_eq\_zero\_of\_relations\_x0\_affineRankOne\_of\_finrank\_range\_one}
shows that if the affine coefficient image is already one-dimensional after
normalizing the exact affine line to \(x_0\), then Lean closes the branch
immediately through the existing affine-rank-one theorem. So the surviving
exact-affine-\(\dim=1\) gap is now narrower again: the genuinely remaining work
is the affine-image rank-\(2\) tailed classifier, not the pure affine-rank-one
surface.

I then corrected one missing affine-\(\dim=1\) endpoint and recorded the
matching numerical split. In
\texttt{RepresentativeAffineRankOne.lean}, the new theorem
\texttt{residual\_eq\_zero\_of\_relations\_x0\_onePlusAX0sqBX1sq\_x0x1\_x1sq}
shows that exact relations
\[
  x_0,\qquad 1 + a x_0^2 + b x_1^2,\qquad x_0x_1,\qquad x_1^2,
  \qquad a \ne 0,
\]
already force zero residual. So the constant \(x_0^2/x_1^2\) affine-\(\dim=1\)
family is no longer just a numerical guess.

I also extended the Julia file
\texttt{julia/affine\_dim\_one\_tailed\_2026\_04\_11.jl} with the new output
\texttt{affine\_dim\_one\_constant\_probe}. The retained summary is:
\[
  \texttt{const\_x0sq\_diag\_ranks}=[15],\qquad
  \texttt{const\_cross\_ranks}=[15],\qquad
  \texttt{const\_x1sq\_common\_ranks}=[13].
\]
So the pure constant-tailed affine-\(\dim=1\) models now split numerically into
three stable types:
\begin{itemize}
\item \((x_0,\ 1 + a x_0^2 + b x_1^2,\ x_0x_1,\ x_1^2)\) is always rank \(15\)
with no missing quartic monomials in the tested scalar sweep;
\item \((x_0,\ 1 + a x_0x_1,\ x_0^2,\ x_1^2)\) also stays rank \(15\);
\item \((x_0,\ 1 + a x_1^2,\ x_0^2,\ x_0x_1)\) stays rank \(13\), with the full
\(x_1\)-axis monomial strip \((0,0),(0,1),(0,2),(0,3),(0,4)\) missing.
\end{itemize}
I then formalized the second rank-\(15\) constant-tail model in
\texttt{RepresentativeAffineRankOne.lean}. The new theorem
\texttt{residual\_eq\_zero\_of\_relations\_x0\_onePlusAX0x1\_x0sq\_x1sq}
closes
\[
  (x_0,\ 1 + a x_0x_1,\ x_0^2,\ x_1^2)
\]
directly by admissible-image surjectivity: Lean reconstructs \(x_1\) from the
constant-cross relation using the exact \(x_0\) and \(x_1^2\) relations, then
reconstructs \(1\), and then reaches every quartic monomial from
\((1,x_0,x_1,x_0^2,x_1^2)\). So the affine-\(\dim=1\) classifier is now
visibly narrower: both rank-\(15\) constant-tail families have moved into the
proved column, while the repeated-line \(1 + a x_1^2\) slice remains the
obvious nonsurjective constant endpoint.

I then formalized that last repeated-line constant endpoint too. The new
theorem
\texttt{residual\_eq\_zero\_of\_relations\_x0\_onePlusAX1sq\_x0sq\_x0x1},
again in \texttt{RepresentativeAffineRankOne.lean}, closes
\[
  (x_0,\ 1 + a x_1^2,\ x_0^2,\ x_0x_1).
\]
The proof uses a new admissible-kernel family built from the exact relations to
\[
  x_0^2,\qquad x_0x_1,
\]
so that \(\sigma\) contributes exactly the pure \(x_1^2/x_1^3/x_1^4\) strip.
That lets Lean correct each SOS summand into the divisibility conditions for
\((1 + a x_1^2)\), and then finish via the indexed-family certificate. So the
constant-tail affine-\(\dim=1\) side is no longer missing a representative or
endpoint theorem. The remaining work is now purely classifier-level: package a
general exact-affine \(\dim=1\), affine-image-rank-\(2\) branch into one of the
already solved constant-tail or \(x_1\)-tail surfaces.

I then stabilized the same repeated-line constant branch under an additional
linear \(x_1\)-tail instead of opening a new classifier abstraction too early.
The new theorem
\texttt{residual\_eq\_zero\_of\_relations\_x0\_onePlusBX1PlusAX1sq\_x0sq\_x0x1},
again in \texttt{RepresentativeAffineRankOne.lean}, closes
\[
  (x_0,\ 1 + b x_1 + a x_1^2,\ x_0^2,\ x_0x_1),\qquad a \ne 0.
\]
Lean performs the internal translation \(x_1 \mapsto x_1 - b/(2a)\), so the
second relation becomes either a pure \(x_1^2\) relation or a
\((1 + a' x_1^2)\) relation, and the translated \(x_0x_1\) relation is then
repaired by subtracting the induced \(x_0\) term. That reduces the whole mixed
constant-linear repeated-line slice to the already solved pure repeated-line
endpoint.

I also extended the Julia script
\texttt{julia/affine\_dim\_one\_tailed\_2026\_04\_11.jl} with the retained
probe \texttt{affine\_dim\_one\_constant\_linear\_probe}. For all tested
nonzero \(a,b \in \{-2,-1,1,2\}\) it reports
\[
  \texttt{const\_linear\_x1sq\_common\_ranks}=[13],\qquad
  \texttt{const\_linear\_x1sq\_common\_missing}
    =[ \{(0,0),(0,1),(0,2),(0,3),(0,4)\} ],\qquad
  \texttt{const\_linear\_x1sq\_common\_exact\_affine\_dims}=[1].
\]
So numerically the mixed repeated-line family behaves exactly like the pure
\(1 + a x_1^2\) slice, which matches the translation proof.

I then lifted that same repeated-line constant branch to determinant-nondegenerate
plane data instead of keeping it tied to exact \((x_0^2,x_0x_1)\) relations.
The retained theorem
\texttt{residual\_eq\_zero\_of\_relations\_x0\_onePlusBX1PlusAX1sq\_x0sq\_x0x1Plane}
in \texttt{RepresentativeAffineRankOne.lean} assumes exact relations to
\[
  x_0,\qquad 1 + b x_1 + a x_1^2,\qquad
  r x_0^2 + s x_0x_1,\qquad t x_0^2 + w x_0x_1
\]
with \(a \ne 0\) and \(rw-st \ne 0\). Lean then reconstructs exact \(x_0^2\)
and \(x_0x_1\) by the same \(2 \times 2\) Cramer's-rule linear combination used
elsewhere in the low-affine plane layer, and finishes through the already
proved translated repeated-line theorem. So that slice now sits one level
closer to the eventual classifier surface.

I then did the same on the diagonal side. In
\texttt{RepresentativeAffineRankOne.lean}, the new theorems
\texttt{residual\_eq\_zero\_of\_relations\_x0\_onePlusAX0x1\_x0sq\_x1sqPlane}
and
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1PlusAX0x1\_x0sq\_x1sqPlane},
together with their equivariant wrappers, lift the constant-cross and
linear-cross affine-rank-one endpoints from exact \((x_0^2,x_1^2)\) relations
to arbitrary determinant-nondegenerate bases of
\(\operatorname{span}(x_0^2,x_1^2)\). So the classifier no longer needs to
hand those cross slices a pre-normalized diagonal basis either.

I then lifted the remaining linear repeated-line endpoints to the same
determinant-nondegenerate plane surface. The new theorems
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1PlusAX0sq\_x0x1\_x1sqPlane}
and
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1PlusAX1sq\_x0sq\_x0x1Plane},
again with their equivariant wrappers, reconstruct exact
\((x_0x_1,x_1^2)\) or \((x_0^2,x_0x_1)\) internally by the same \(2\times2\)
Cramer's-rule step and then dispatch to the already proved linear repeated-line
endpoint theorems. So the exact-affine \(\dim=1\) classifier no longer has to
hand those branches an exact repeated-line basis either.

I then pushed one retained exact-affine \(\dim=1\) classifier case into Lean.
In \texttt{ExactAffineClassification.lean}, the new theorem
\texttt{residual\_eq\_zero\_of\_exactAffineDimOne\_tailRangeZero} closes the
tail-rank \(0\) output after fixing an exact \(x_0\) relation: if the
\((1,x_1)\)-tail map on the \(x_0\)-kernel has zero image, Lean proves the
three residual quadratic relations are already exactly
\[
  x_0^2,\qquad x_0x_1,\qquad x_1^2
\]
and dispatches to the existing exact theorem
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x0sq\_x0x1\_x1sq}. So the
exact-affine \(\dim=1\) branch no longer leaves the pure homogeneous-tail case
to the external classifier.

In the same round I retained the structural extractor
\texttt{exists\_x0\_tail\_const\_x1\_hom\_of\_exactAffineDimOne\_rangeTwo}.
That theorem does not close the tail-rank \(2\) branch by itself, but it shows
that once the tail map has full rank \(2\), Lean can normalize the
\(x_0\)-kernel to exact relations with tails \(1\), \(x_1\), and \(0\), plus a
single nonzero pure homogeneous quadratic relation.

I then strengthened that extractor in
\texttt{ExactAffineClassification.lean} to
\texttt{exists\_x0\_tail\_const\_x1\_hom\_independent\_of\_exactAffineDimOne\_rangeTwo}.
This packages the first two contradictions that the later classifier would
otherwise have to rediscover. Lean now proves that the normalized constant
relation minus \(1\) and the normalized \(x_1\)-tail relation minus \(x_1\)
are each linearly independent from the unique pure homogeneous line. If either
pair collapsed, one would recover an exact \(1\) or exact \(x_1\) relation,
contradicting \(\dim(\mathrm{ExactAff})=1\). So the tail-rank \(2\) output is
now a cleaner classifier-ready package rather than only a basis-normalization
statement.

I then strengthened the same extractor once more, to
\texttt{exists\_x0\_tail\_const\_x1\_hom\_basis\_of\_exactAffineDimOne\_rangeTwo}.
This closes the remaining linear-algebra ambiguity on the homogeneous side:
Lean now proves that the three homogeneous quadratics
\[
  q_0-1,\qquad q_1-x_1,\qquad q_2
\]
are linearly independent. The proof is the expected exact-affine contradiction:
any nontrivial linear relation among those three would produce an exact affine
relation \(r+b x_1\), but in a one-dimensional exact-affine space already
containing \(x_0\), every exact affine relation must be a scalar multiple of
\(x_0\), forcing \(r=b=0\) and then the last coefficient to vanish too. So the
tail-rank \(2\) output is now a full homogeneous-basis package rather than only
two pairwise independence statements.

I then added the matching rank-\(1\) extractor
\texttt{exists\_x0\_tail\_nonzero\_homPair\_of\_finrank\_range\_one}. This
packages the remaining intermediate exact-affine \(\dim=1\) case much more
cleanly: once the \((1,x_1)\)-tail map on the \(x_0\)-kernel has rank \(1\),
Lean now extracts one exact relation with nonzero tail and two further exact
relations whose constant and \(x_1\) coefficients both vanish, and proves those
two homogeneous quadratics are linearly independent. So the rank-\(1\) tail
case is no longer blocked at a raw kernel-dimension statement; the remaining
work is the classifier that routes one tailed quadratic over a homogeneous
plane into the already solved affine-rank-one endpoint surfaces.

I then strengthened that same rank-\(1\) extractor to
\texttt{exists\_x0\_tail\_nonzero\_hom\_basis\_of\_exactAffineDimOne\_rangeOne}.
This removes the last purely linear-algebra ambiguity on the homogeneous side:
after subtracting the explicit affine tail \(r+b x_1\) from the unique tailed
relation, Lean now proves that the three homogeneous quadratics
\[
  q_{\mathrm{tail}}-(r+b x_1),\qquad q_1,\qquad q_2
\]
are linearly independent. The proof is the same exact-affine contradiction as
in the tail-rank \(2\) branch: any nontrivial dependence would create an exact
affine relation of the form \(r'+b' x_1\), but in a one-dimensional
exact-affine space already containing \(x_0\), every such relation must vanish.
So the rank-\(1\) output is now not just one tailed relation plus a
homogeneous plane, but a full homogeneous-basis package behind that picture.

I also recorded the new rank-\(1\) numerical split in
\texttt{julia/affine\_dim\_one\_tail\_rank\_one\_2026\_04\_11.jl}. Its retained
output is
\[
  \texttt{affine\_dim\_one\_tail\_rank\_one\_probe}
  =
  (\texttt{samples}=20,\ 
   \texttt{mixed\_tail\_ranks}=[1],\ 
   \texttt{mixed\_exact\_affine\_dims}=[1],\ 
   \texttt{mixed\_image\_ranks}=[15],\ 
   \texttt{mixed\_missing}=[\varnothing],\ 
   \texttt{const\_image\_ranks}=[15],\ 
   \texttt{const\_missing}=[\varnothing],\ 
   \texttt{linear\_image\_ranks}=[14],\ 
   \texttt{linear\_missing}=[\{(0,0)\}]).
\]
So the generic rank-\(1\) tail family looks surjective in the mixed and
constant directions, while the pure \(x_1\)-tail direction still shows the
same codimension-one missing constant monomial pattern as the retained
normalized affine-rank-one linear endpoint.

For the tail-rank \(2\) side, I had already recorded the corresponding probe in
the Julia file \texttt{julia/affine\_dim\_one\_mixed\_tails\_2026\_04\_11.jl}.
Its retained output is:
\[
  \texttt{affine\_dim\_one\_mixed\_tail\_probe} =
  (\texttt{samples}=20,\ 
   \texttt{tail\_ranks}=[2],\ 
   \texttt{exact\_affine\_dims}=[1],\ 
   \texttt{image\_ranks}=[15],\ 
   \texttt{missing}=[\varnothing]).
\]
So the exact-affine \(\dim=1\) picture is now sharper on both numerical and
formal sides: both the tail-rank \(1\) and tail-rank \(2\) cases already have
extractor theorems, and numerically the only stable rank-\(1\) obstruction
still visible is the pure \(x_1\)-tail direction. The remaining work is the
formal classifier that routes those outputs into the correct already-solved
theorem surfaces.

I then added one more affine-rank-one plane wrapper in
\texttt{RepresentativeAffineRankOne.lean}:
\texttt{residual\_eq\_zero\_of\_relations\_x0\_onePlusAX0sqBX1sq\_x0x1\_x1sqPlane},
with the matching equivariant version. This lifts the mixed constant repeated-line
endpoint
\[
  (x_0,\ 1 + a x_0^2 + b x_1^2,\ x_0x_1,\ x_1^2)
\]
from an exact repeated-line basis to arbitrary determinant-nondegenerate plane
data in \(\operatorname{span}(x_0x_1,x_1^2)\). The proof is the same internal
\(2\times 2\) Cramer's-rule reconstruction already used for the other plane
wrappers. So the remaining exact-affine \(\dim=1\) classifier no longer needs
to hand that constant repeated-line slice an exact \((x_0x_1,x_1^2)\) basis
either.

I then added a direct exact-normalized mixed-tail endpoint in the same file:
\texttt{quartic\_in\_image\_of\_relations\_x0\_onePlusAX0sq\_x1PlusBX0x1\_x1sq}
and
\texttt{residual\_eq\_zero\_of\_relations\_x0\_onePlusAX0sq\_x1PlusBX0x1\_x1sq}.
This closes exact relations to
\[
  x_0,\qquad 1 + a x_0^2,\qquad x_1 + b x_0x_1,\qquad x_1^2
\]
with \(a \ne 0\) by a direct monomial image argument. Lean first reconstructs
\(1\), \(x_1\), \(x_0^2\), and \(x_0x_1\), then produces
\(x_0^2x_1\), \(x_0^3x_1\), and \(x_0^4\) using quadratic multipliers against
the \(x_0\)- and \(1+a x_0^2\)-relations. No new Julia experiment was needed
for this round. The practical effect is that one more tailed exact-affine
\(\dim=1\) output now closes directly at representative level instead of
needing an extra plane reduction.

I then returned to the exact-affine \(\dim=1\), tail-rank \(1\) classifier
surface and strengthened the affine-rank-one plane layer in
\texttt{RepresentativeAffineRankOne.lean}. The new theorems
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1Plus\_homQuadratics\_x0sq\_x1sqPlane},
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1Plus\_homQuadratics\_x0x1\_x1sqPlane},
and
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1Plus\_homQuadratics\_x0sq\_x0x1Plane},
with their equivariant wrappers, no longer require the classifier to hand back
an already normalized tailed coordinate. Instead, once the affine part is
normalized to \(x_1\) and the homogeneous part is known to lie in one of the
three canonical determinant-nondegenerate planes, Lean reconstructs the exact
plane basis internally by Cramer's rule, subtracts the homogeneous part of the
tailed relation, and dispatches to the previously solved exact linear-tail
endpoint. No new Julia computation was needed for this round. The remaining
exact-affine \(\dim=1\) work is therefore pushed further upward into the final
classifier, not the local affine-rank-one plane algebra.

I then added the constant-tail companions in the same file:
\texttt{residual\_eq\_zero\_of\_relations\_x0\_onePlus\_homQuadratics\_x0x1\_x1sqPlane},
\texttt{residual\_eq\_zero\_of\_relations\_x0\_onePlus\_homQuadratics\_x0sq\_x1sqPlane},
and
\texttt{residual\_eq\_zero\_of\_relations\_x0\_onePlus\_homQuadratics\_x0sq\_x0x1Plane}.
These are the exact same Cramer's-rule upgrades on the constant side: once the
classifier can produce \(x_0\), one quadratic whose affine part is exactly
\(1\), and a determinant-nondegenerate basis of one canonical homogeneous
plane, Lean rebuilds the exact plane basis internally and reduces to the
already solved constant affine-rank-one endpoints. So both the pure \(x_1\)
tail and pure \(1\) tail branches now have coarse plane wrappers above the
exact normalized representative layer. No new Julia experiment was needed for
this round either.

I then pushed one small but real piece of the exact-affine \(\dim=1\),
tail-rank \(1\) classifier itself into Lean. In
\texttt{ExactAffineClassification.lean} I added
\texttt{residual\_eq\_zero\_of\_relations\_x0\_homQuadBasis\_singleConstTail}
and
\texttt{residual\_eq\_zero\_of\_relations\_x0\_homQuadBasis\_singleX1Tail}.
These sit directly above the canonical hom-basis extraction theorem
\texttt{exists\_relations\_x0\_homQuadBasis\_of\_exactAffineDimOne}. Once the
extracted tails are supported on exactly one canonical homogeneous direction,
Lean now rescales that single tailed relation and dispatches immediately to the
already-solved affine-rank-one endpoint or coarse-plane theorem. Concretely,
all six support-one patterns are now closed:
\[
  (1+x_0^2,\ x_0x_1,\ x_1^2),\quad
  (1+x_0x_1,\ x_0^2,\ x_1^2),\quad
  (1+x_1^2,\ x_0^2,\ x_0x_1),
\]
and their pure \(x_1\)-tail analogues
\[
  (x_1+x_0^2,\ x_0x_1,\ x_1^2),\quad
  (x_1+x_0x_1,\ x_0^2,\ x_1^2),\quad
  (x_1+x_1^2,\ x_0^2,\ x_0x_1).
\]
I then added one more exact-affine \(\dim=1\), tail-rank \(1\) closure that is
still classifier-facing but no longer only pure support. The new theorem
\texttt{residual\_eq\_zero\_of\_relations\_x0\_homQuadBasis\_singleMixedTail\_x1sq}
handles the canonical subfamily in which only the \(x_1^2\)-direction carries
tails, but it carries both a nonzero constant term and a nonzero \(x_1\)-term:
\[
  x_0,\qquad x_0^2,\qquad x_0x_1,\qquad a + b x_1 + x_1^2.
\]
The Lean proof is short because the algebra is already in place. It rescales
the tailed \(x_1^2\)-relation to \(1 + b' x_1 + a' x_1^2\), leaves the exact
\((x_0^2,x_0x_1)\) pair alone, and dispatches directly to the translated
repeated-line theorem in \texttt{RepresentativeAffineRankOne.lean}. So the
support-one repeated-line mixed slice is no longer open either. No new Julia
computation was needed for this round.
I then added the wrapper
\texttt{residual\_eq\_zero\_of\_relations\_x0\_homQuadBasis\_tailOnX1sq}. This
collapses the three \(x_1^2\)-supported support-one possibilities into one
classifier-facing theorem:
\[
  x_0,\qquad x_0^2,\qquad x_0x_1,\qquad a + b x_1 + x_1^2,
\]
with the only assumption \(a \neq 0 \vee b \neq 0\). Lean now does the
remaining split internally:
\[
  a = 0,\ b \neq 0,\qquad a \neq 0,\ b = 0,\qquad a \neq 0,\ b \neq 0,
\]
and dispatches to the already proved pure \(x_1\)-tail, pure constant-tail,
and mixed repeated-line theorems respectively. So the exact-affine
\(\dim=1\), tail-rank \(1\) classifier is one support test simpler on that
canonical slice.

I then closed the remaining support-one mixed slices too. In
\texttt{RepresentativeAffineRankOne.lean} I added the exact endpoints
\texttt{residual\_eq\_zero\_of\_relations\_x0\_onePlusBX1PlusAX0sq\_x0x1\_x1sq}
and
\texttt{residual\_eq\_zero\_of\_relations\_x0\_onePlusBX1PlusAX0x1\_x0sq\_x1sq}.
These cover the two missing mixed constant-plus-\(x_1\) support-one families:
\[
  x_0,\qquad 1 + b x_1 + a x_0^2,\qquad x_0x_1,\qquad x_1^2,
\]
and
\[
  x_0,\qquad 1 + b x_1 + a x_0x_1,\qquad x_0^2,\qquad x_1^2.
\]
On top of those, \texttt{ExactAffineClassification.lean} now proves
\texttt{residual\_eq\_zero\_of\_relations\_x0\_homQuadBasis\_singleMixedTail\_x0sq},
\texttt{residual\_eq\_zero\_of\_relations\_x0\_homQuadBasis\_singleMixedTail\_x0x1},
\texttt{residual\_eq\_zero\_of\_relations\_x0\_homQuadBasis\_tailOnX0sq},
and
\texttt{residual\_eq\_zero\_of\_relations\_x0\_homQuadBasis\_tailOnX0x1}. So
the whole support-one mixed tail family is now formalized, not just the
\(x_1^2\) repeated-line slice.

I also recorded the matching scalar sweep in
\texttt{julia/affine\_dim\_one\_tailed\_2026\_04\_11.jl}. Its new retained
output is
\[
  \texttt{affine\_dim\_one\_support\_one\_mixed\_probe} =
  (\texttt{vals}=[-2,-1,1,2],\ 
   \texttt{mixed\_x0sq\_ranks}=[15],\ 
   \texttt{mixed\_x0sq\_missing}=[\varnothing],\ 
   \texttt{mixed\_x0sq\_exact\_affine\_dims}=[1],\ 
   \texttt{mixed\_x0x1\_ranks}=[15],\ 
   \texttt{mixed\_x0x1\_missing}=[\varnothing],\ 
   \texttt{mixed\_x0x1\_exact\_affine\_dims}=[1]).
\]
So the two newly formalized support-one mixed slices are uniformly surjective
on that retained test grid, exactly as the new Lean endpoints assert.

I then widened the affine-rank-one plane surface in the same mixed direction.
In \texttt{RepresentativeAffineRankOne.lean} I added
\texttt{residual\_eq\_zero\_of\_relations\_x0\_onePlusBX1Plus\_homQuadratics\_x0x1\_x1sqPlane},
\texttt{residual\_eq\_zero\_of\_relations\_x0\_onePlusBX1Plus\_homQuadratics\_x0sq\_x1sqPlane},
and
\texttt{residual\_eq\_zero\_of\_relations\_x0\_onePlusBX1Plus\_homQuadratics\_x0sq\_x0x1Plane}.
These are the mixed constant-plus-\(x_1\) analogues of the older
\texttt{onePlus\_homQuadratics\_\dots Plane} wrappers: once the classifier has
exact \(x_0\), a quadratic with affine part \(1 + b x_1\), and a
determinant-nondegenerate basis of one canonical homogeneous plane, Lean now
subtracts the in-plane homogeneous directions internally and reduces to the
already formalized mixed endpoints. No new Julia data was needed for this
round; the change is purely on the theorem-surface side.

I then lifted that mixed surface one chart higher. The new theorem
\texttt{residual\_eq\_zero\_of\_relations\_x0\_onePlusBX1Plus\_homQuadratics\_crossDet\_zero}
handles the coprime cross chart for a mixed constant-plus-\(x_1\) tailed
relation. It uses the same internal \(x_1\)-shear as the older constant-tail
cross theorem, then subtracts the induced \(x_0\)-linear term from the tailed
relation before applying the existing
\texttt{onePlusBX1Plus\_homQuadratics\_x0sq\_x1sqPlane} wrapper. No new Julia
data was needed for this step either.

I then added the pure \(x_1\)-tail companion chart theorem
\texttt{residual\_eq\_zero\_of\_relations\_x0\_x1Plus\_homQuadratics\_crossDet\_zero}.
The proof is the same shear-and-repair pattern as above, but because the
endpoint is now \texttt{x1Plus\_homQuadratics\_x0sq\_x1sqPlane}, the theorem
records one real structural condition explicitly: after the internal
\(x_1\)-shear, the tailed relation must still have nonzero \(x_0x_1\)
coefficient. No new Julia data was needed for this step.

I then added the constant-tail coprime cross-chart analogue. In
\texttt{RepresentativeAffineRankOne.lean} the new theorem
\texttt{residual\_eq\_zero\_of\_relations\_x0\_onePlus\_homQuadratics\_crossDet\_zero}
shows that once the pure homogeneous pair satisfies
\(\texttt{lowHomQuadPlaneA}=0\) and \(\texttt{lowHomQuadPlaneB}\neq 0\), Lean
can do the necessary internal \(x_1\)-shear itself and reduce directly to the
already solved \((x_0^2,x_1^2)\)-plane wrapper. So the exact-affine
\(\dim=1\) classifier no longer needs to hand back an explicitly diagonalized
homogeneous pair for that constant-tail chart; chart data alone is enough.

\section*{Open Subgoals}
\begin{itemize}
\item Extract from a general admissible factor the remaining exact-affine
\(\dim=1\) data. The constant-containing \(\dim=2\) mixed-affine branch no
longer needs external quadratic-pair extraction, and the exact-affine
\(\dim=2\) low-affine branch is now closed. What remains is the upstream
classification step that either excludes or normalizes the newly observed
tailed \(\dim=1\) families, not merely the pure affine-rank-one image.
\item After that branch is settled, assemble the final theorem in
\texttt{TernaryQuarticProof.lean} and rerun the verification harness.
\end{itemize}

\end{document}
